### 2024-03-21 data_processing/Readme.Rmd
holiday break

---
  title: "RNAseq aCRY processing"
author: "Kelterborn"
date: "2024-03-15"
output:
  md_document:
  variant: gfm
toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

# Download data (AWS CLI)

The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
Index in AWS: bmkdatarelease-37/delivery_20240313120107785
Access key ID: AKIA45O7KHZWJCKW62VE
Secret Access Key: 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
Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
Username:delivery_20240313120107785
Password:Pwd2024@x4S416

```{bash, eval=FALSE}

# install aws
curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
unzip awscliv2.zip
sudo ./aws/install

aws --version

aws configure
AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
AWS Secret Access Key [None]: 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
Default region name [None]: eu-north-1
Default output format [None]: table

fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
cd $fastq

# test
aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt

# all files
aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive

```

# 0. Prepare System
BiocManager::install()

## -R_libraries
```{r library}
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library("ggpubr")
library(vsn)

library(PCAtools)

# library(wget)

```

## -R_folders
```{r r_folders}

```

## -Linux folders
```{bash linux_folders, eval=FALSE}
fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;

dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY"

quantdir="$dir/quants"

indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1"

```

# 1. Mapping
```{bash mapping, eval=FALSE}
mamba activate salmon;

fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);

# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
fastqfiles_all=$fastqfiles;

echo "${fastqfiles:0:100}";
echo "${fastqfiles2:0:100}";
echo "${fastqfiles_all:0:100}";

# files: Unknown_BU327-002T0001_1.fq.gz;
for fn in $fastqfiles_all;
do
bfn=$(echo "`basename ${fn}`");
samp=${bfn%"_1.fq.gz"};
R1=$fn;
R2=$(echo "$R1" | sed 's/_1/_2/');
echo "Processing Sample: $samp";
test -f $R1 && echo "--> File: $R1 exists"
test -f $R2 && echo "--> File: $R2 exists"
outquant=${quantdir}/${samp}_quant
echo $outquant
echo $indexdir

salmon quant -i $indexdir -l A \
  -1 $R1 \
  -2 $R2 \
  -p 25 --validateMappings --gcBias -o $outquant

done

```

### 2024-04-09 1_data_processing/1_data_processing.Rmd
deseq2

---
  title: "RNAseq aCRY processing"
author: "Kelterborn"
date: "2024-03-15"
output:
  github_document:
    html_preview: false
    toc: true
always_allow_html: true
editor_options: 
  chunk_output_type: console
knit: (function(input_file, encoding) {
    })
    rmarkdown::render(input_file,output_file= 'README.md')
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)
```

# 0. Prepare System
BiocManager::install()

## -R_libraries
```{r r_libraries}
library(stringr)
library(R.utils)
library(RColorBrewer)
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)
library(AnnotationHub)
library(DESeq2)
library(EnhancedVolcano)
library(pheatmap)
library(writexl)
library(biomaRt)
library(ape)
library(kableExtra)
library(knitr)

library(stringr)
library(R.utils)
library(RColorBrewer)

library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

library(SummarizedExperiment)
library(GenomicRanges)
library(ape)

library(viridis)
library(patchwork)
library("ggpubr")
library(vsn)

library(PCAtools)

# library(wget)

```

## -R_folders
```{r r_folders}
ifelse(Sys.info()["sysname"]== "Linux",
       s <- "/mnt/s",
       s <- "S:")
dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
list.files(dir) %>% head()
gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
list.files(gitdir) %>% head()

fastqdir <- paste(dir,"fastq",sep="/")
list.files(fastqdir) %>% tail()
fastqdir2 <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/",sep="/")
list.files(fastqdir2) %>% head()

indexDir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")

fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
head(readLines(fastaPath,n=50))

gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
readLines(gtfPath,n=20)

quantdir <- paste(dir,"quants",sep="/")

```

## -Linux folders
```{bash linux_folders, eval=FALSE}
fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;

dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY";

quantdir="$dir/quants";

indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";

```

# 1. Raw data processing
## 1a. Download data (AWS CLI)

The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
Index in AWS: bmkdatarelease-37/delivery_20240313120107785
Access key ID: AKIA45O7KHZWJCKW62VE
Secret Access Key: 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
Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
Username:delivery_20240313120107785
Password:Pwd2024@x4S416

```{bash download data, eval=FALSE}

# install aws
curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
unzip awscliv2.zip
sudo ./aws/install

aws --version

aws configure
AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
AWS Secret Access Key [None]: 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
Default region name [None]: eu-north-1
Default output format [None]: table

fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
cd $fastq

# test
aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt

# all files
aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive

```

### check md5
```{bash md5, eval=FALSE}
cd $fastqdir
ls
# 1 file
md5sum Unknown_BU327-002T0008_1.fq.gz

# Check downloads

# Path to the text file containing the MD5 sums
md5_file="data_md5.txt"

# Function to check the MD5 sum of a file
check_md5() {
    file_name=$1
    expected_sum=$2

    md5_sum=$(md5sum "$file_name" | awk '{print $1}')
    if [[ $md5_sum == $expected_sum ]]; then
        echo "$file_name: MD5 sum is correct."
    else
        echo "$file_name: MD5 sum is incorrect."
    fi
}

# Read the MD5 sums from the text file and check them
while IFS= read -r line; do
    file_name=$(echo "$line" | awk '{print $2}')
    expected_sum=$(echo "$line" | awk '{print $1}')
    if [[ -f $file_name ]]; then
        check_md5 "$file_name" "$expected_sum"
    else
        echo "$file_name: File not found."
    fi
done < "$md5_file" | tee -a $log

```

## 1b. Fastqc

## 1c. Mapping
```{bash mapping, eval=FALSE}
mamba activate salmon;

fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);

# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
fastqfiles_all=$fastqfiles;

echo "${fastqfiles:0:100}";
echo "${fastqfiles2:0:100}";
echo "${fastqfiles_all:0:100}";

# files: Unknown_BU327-002T0001_1.fq.gz;
for fn in $fastqfiles_all;
do
bfn=$(echo "`basename ${fn}`");
samp=${bfn%"_1.fq.gz"};
R1=$fn;
R2=$(echo "$R1" | sed 's/_1/_2/');
echo "Processing Sample: $samp";
test -f $R1 && echo "--> File: $R1 exists"
test -f $R2 && echo "--> File: $R2 exists"
outquant=${quantdir}/${samp}_quant
echo $outquant
echo $indexdir

salmon quant -i $indexdir -l A \
  -1 $R1 \
  -2 $R2 \
  -p 25 --validateMappings --gcBias -o $outquant

done

```

# 2. Process Data
## 2a. Sample names
```{r sample_table}
# Get Sample table
# Filenames

quantdir <- paste(dir,"quants",sep="/")

f <- list.files(path = quantdir)
files <- as.factor(str_remove(f,pattern ="_quant"))
# short form
sample <- str_remove(files,pattern ="Unknown_")
sample
# condition
library(readxl)
Samplepath <- filePath(path=dir, file="sample_table.xlsx")
Samplefile <- read_xlsx(Samplepath)
samplename <- Samplefile$clientName
Samplefile$treatment
Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor() %>% relevel(ref="WT_dark")
# levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red")

# Make table
sample.table <- data.frame(Samplefile[,-3])
sample.table

sample.table <- data.frame(lapply(sample.table, factor))

# Check the updated structure of the data frame
str(sample.table)

```

## 2b. Mapping Rates

```{r mapping_rates}
library(stringr)
library(R.utils)
library(RColorBrewer)

samplelist <- {}
mappingrates <- {}
for (i in list.files(path = quantdir)){
print(i)
si <- paste("",str_sub(i,11,-7),sep = "")
si
samplelist <- c(samplelist,si)
f <- readLines(paste(quantdir,i,"logs/salmon_quant.log", sep="/"))
line <- grep("Mapping rate = ",f,value=TRUE)
sl <- str_length(line)
sl
notime <- substring(line,30,sl)
notime
manual <- substring(line,sl-7,sl-1)
val <- as.numeric(str_extract(notime,"[0-9.]+"))
val
valr<-round(val, digits=2)
print(paste("Mapping rate of ",si," is: ",valr," %"))
mappingrates <- c(mappingrates,valr)
}
sample.table$mappingrates <- mappingrates
# Make table

m.table <- data.frame(sample.table,mappingrates)
```

### Plot mapping rates
```{r plot_mappingrates}

# Colours

# Plot
par(mar=c(2,4,0,0)+.1)
# plot(sample_table_all$mappingrates)
# -> boring

# increase margin for longer names
par(mar=c(2,5,2,2)+.1)
# barplot(height=mappingrates, names=samplename, horiz=T, las=1)

xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename, horiz=T, las=1, xlim=c(0,100))
text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.5, col = "red")

par(mar=c(4,4,4,4)+.1)

```

## 2c. Tximeta

```{r tximeta}
library(sessioninfo)
library(data.table)
library(plyr)
library(tidyverse)
library(tximeta)
library(tximport)
library(curl)

getwd()
example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")

list.files(quantdir)[1]
file.exists(example.quant)

# first file:
example.table <- read.table(example.quant, header=T)
head(example.table)
basename(example.quant)
dirname(example.quant)
basename(dirname(example.quant))

# generate file list & prepare variables
files={}
for (i in list.dirs(path = dirname(dirname(example.quant)), full.names = TRUE, recursive = FALSE)) {
files <- c(files,paste(i,"/quant.sf",sep=""))
print(basename(i))
# print(head(read.table(files[length(files)], header=T)))
print (files[length(files)])
}
head(files)

dirfiles <- dirname(files)
head(dirfiles)
head(basename(dirfiles))
run <- basename(dirfiles)
dirdir <- dirname(dirfiles)
head(dirdir)

tximeta_files <- file.path(dirdir, run, "quant.sf") 
file.exists(tximeta_files)
file.exists(files)

# names as working sample names
# see 'm.table' at mapping rates
coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
coldata <- data.frame(coldata,sample.table)
coldata %>% head()

# load tximeta
# with linked Transcriptome
## Chlamy
file.exists(indexDir, fastaPath, gtfPath)

makeLinkedTxome(indexDir=indexDir,
                source="Phytozome",
                organism="Chlamydomonas reinhardtii",
                release="v6.1",
                genome="CreinhardtiiCC_4532_707",
                fasta=fastaPath,
                gtf=gtfPath,
                write=FALSE)

gtfdata <- readLines(gtfPath)
head(gtfdata, n=20)

# gene_name info sind in gff3 abder nicht in se?
coldata_Cre <- filter(coldata, species == "Cre")
se <- tximeta(coldata_Cre, useHub=T)
se

library(SummarizedExperiment)

colData(se)
# meta infos sind da (genotype, treatment,...)

# rename Samples
rownames(colData(se)) <- colData(se)$clientName

rowRanges(se)
# genome info
seqinfo(se)
# ?
head(assays(se)[["counts"]])
# counts. THE DATA

# Mapping infos:
names(metadata(se)[["quantInfo"]])
str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
metadata(se)[["quantInfo"]]$percent_mapped
metadata(se)[["quantInfo"]]$num_processed

par(mfrow=c(1,2))
barplot(metadata(se)[["quantInfo"]]$percent_mapped, main="Mapping Rate")
barplot(metadata(se)[["quantInfo"]]$num_processed/1000000, main="Mio. Reads")
par(mfrow=c(1,1))

edb <- retrieveDb(se)
class(edb)
genes(edb)
columns(edb)

se.exons <- addExons(se)
rowRanges(se.exons)[[1]]
gse <- summarizeToGene(se)
rowRanges(gse)

head(assays(gse)[["counts"]])[1:5,1:5]
head(assays(gse)[["abundance"]])[1:5,1:5]
head(assays(gse)[["length"]])[1:5,1:5]

#### add gene symbol
columns(retrieveDb(se))
edb
metadata(se)$txomeInfo$source

TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="first"))
# select()' returned 1:many mapping between keys and columns = 1 gene has many transcripts...
TXNAME.list <- (mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="list"))
head(TXNAME.list)
# info is already there
CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
head(CDSID)
# ?
CDSNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSNAME", keytype = "GENEID", multiVals="first"))
head(CDSNAME)
# same as gene_id
EXONNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "EXONNAME", keytype = "GENEID", multiVals="first"))
head(EXONNAME)
TXTYPE <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXTYPE", keytype = "GENEID", multiVals="first"))
head(TXTYPE)
# no useful info... no gene Symbol!!

# add CDSID
mcols(gse)$CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))

colnames(mcols(gse))
head(rownames(gse))

getwd()
# save(gse,file=paste(dir,"tximeta.txm", sep="/"))
# gse <- 1
# load(file=paste(outdir,"tximeta.txm", sep="/"))

```

### -Add gene info from Master annotation

Use other files from Phytozome database: CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv

```{r anno}
library(data.table)
library(plyr)
library(tidyverse)
library(ape)

# more gene info from master_annotation_table
annopath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")

anno <- read.delim2(annopath, header = T, sep = "\t",)
head(anno)

# search for genes
anno[str_detect(anno[["locusName_4532"]],"Cre01.g000050"),]
anno[str_detect(anno[["geneSymbol"]],"HKR"),]
anno[str_detect(anno[["previousIdentifiers"]],"HKR"),]
anno[str_detect(anno[["geneSymbol"]],"COP"),]
anno[str_detect(anno[["previousIdentifiers"]],"COP1"),]
anno[str_detect(anno[["geneSymbol"]],"cop"),]
anno[str_detect(anno[["geneSymbol"]],"PLAP6"),]
anno[str_detect(anno[["geneSymbol"]],"COQ3"),]
anno[str_detect(anno[["geneSymbol"]],"PCRY"),]
anno[str_detect(anno[["geneSymbol"]],"CRY"),]
anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH"),]

dim(anno)

# Remove "_4532"
anno$gene_id <- str_remove(anno$locusName_4532,pattern = "_4532")
rownames(anno) <- anno$gene_id

# convert previousIdentifiers to list
anno$previousIdentifiers_list <- strsplit(anno$previousIdentifiers, "#")
head(anno)
anno$previousIdentifiers_list[1]

# polish "previousIdentifiers"
# remove Cre & g-numbers
head(anno$previousIdentifiers)
tail(anno$previousIdentifiers)
prev.symbols <-  
  str_replace_all(anno$previousIdentifiers,pattern="Cre\\d+.g\\d+.t\\d+.\\d+", replacement="?") %>% 
  str_replace_all(pattern="Cre...g\\d+", replacement="?") %>% 
  str_replace_all(pattern="g\\d+.t\\d+", replacement="?") %>% 
  str_remove_all(pattern="\\?") %>%
  str_remove_all("^#|#$| ") %>% str_remove_all("^#|#$") %>% str_remove_all("^#|#$|^0") %>% 
  str_replace_all(pattern="##","#") %>% str_replace_all(pattern="##","#")
prev.symbols[str_detect(prev.symbols,"#")] %>% head(n=20)
prev.symbols[str_detect(prev.symbols,"0")] %>% head(n=20)
prev.symbols[str_detect(prev.symbols,"Cre")] %>% head(n=20)
tail(anno$previousIdentifiers[str_detect(anno$previousIdentifiers,"Cre")])

anno$prev.symbols <- prev.symbols
head(anno)

# Make colum with gene symbol or prev.symbol or Gene id
anno$combname <- anno$geneSymbol

summary(anno$combname)
summary(anno$combname=="")
summary(anno$combname==" ")
summary(is.na(anno$combname))
summary(anno$combname=="0")

# replace "" with prev.symbol
summary(anno$combname=="")
  length(anno$combname[anno$combname==""])
  length(str_split(
  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1])
anno$combname[anno$combname==""] <- str_split(
  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1]
summary(anno$combname=="")

anno$combname[anno$combname==""] <- anno$gene_id[anno$combname==""]
summary(anno$combname=="")
subset(anno,anno$combname=="")

summary(duplicated(anno$combname))
# some duplicates

colnames(anno)[which(names(anno) == "combname")] <- "id.symbol"

summary(anno$geneSymbol!="")
# 4408 gene Symbols
summary(anno$previousIdentifiers!="")
# all (!) 17712 genes have "previousIdentifiers"
summary(anno$prev.symbols!="")
# 7236 have alternative/old symbols
summary(anno$id.symbol)

dim(anno)

anno %>% head()

# saveRDS(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
# anno <- readRDS(paste(outdir,"anno.RDS",sep="/"))

```

### -add anno data into gse ####
```{r anno_gse}

# load(file=paste(outdir,"tximeta.txm", sep="/"))
gse
anno <- readRDS(file=paste(dir,"/anno.RDS", sep="/"))
anno %>% head()

# Remove "_4532"
mcols(gse)$gene_id2 <- mcols(gse)[,"gene_id"]
mcols(gse)$gene_id <- str_remove(mcols(gse)[,"gene_id2"], pattern = "_4532")
rownames(gse) <- mcols(gse)$gene_id

summary(anno$gene_id %in% mcols(gse)$gene_id)
summary(mcols(gse)$gene_id %in% anno$gene_id)

# choose info
gsetable <- as.data.frame(mcols(gse)[,c("gene_id","gene_id2")])

anno_join <- anno[,c("gene_id", "geneSymbol", "id.symbol","prev.symbols", "previousIdentifiers", "previousIdentifiers_list", "Description", "Comments", "TargetP", "Predalgo")]
anno_join

class(anno_join)
dim(anno_join)

# any duplicated:
summary(duplicated(anno$gene_id))

new <- plyr::join(gsetable,anno_join, by = "gene_id", type = "left")
dim(new)
head(new)
new2 <- left_join(gsetable,anno_join,by = "gene_id")
dim(new2)
head(new2)

mcols(gse) <- left_join(as.data.frame(mcols(gse)),anno_join,by = "gene_id")
mcols(gse)

getwd()

# save(gse,file=paste(dir,"tximeta.txm",sep="/"))
```

## 2c. DESeq2 Analysis

```{r deseq2}
getwd()
# load(file=paste(outdir,"tximeta.txm", sep="/"))

library(DESeq2)
colData(gse)
mcols(gse)
gse$condition

design <- ~condition
# design <- ~genotype+treatment+genotype:treatment

dds <- DESeqDataSet(gse, design=design)
colData(dds) %>% head()

##########################################\n
## filter all rows with rowsum = 0 ####
par(mfrow=c(1,2))
hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
# -> dip at log(counts(dds)=2-3)
par(mfrow=c(1,1))

# min counts
# at least 5 counts in 3 samples
keep.sn <- rowSums(counts(dds) >= 5) >= 3
keep.sn %>% summary()
dds <- dds[keep.sn,]

dds <- estimateSizeFactors(dds)
dds
head(assays(dds)[["counts"]])[1:5,1:5]
length(rownames(counts(dds)))

colData(dds)
# add metadata
# colData(dds)$strain <- factor(colData(dds)$strain)

# ## combine technical replicates #####
# colData(dds)
# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
# colData(dds)
# dds$runsCollapsed
# colnames(dds)
# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))

###########
# run DESeq
dds <- DESeq(dds)
DESeq2::plotMA(dds)

plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition")

head(mcols(dds)$geneSymbol)

anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
# no hit
g <- anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH1"),"gene_id"]
g
plotCounts(dds, gene = g, intgroup = "condition")
g <- anno[str_detect(anno[["geneSymbol"]],"PCRY"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
g <- anno[str_detect(anno[["geneSymbol"]],"ACRY"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
g <- anno[str_detect(anno[["geneSymbol"]],"ROC15"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
g <- anno[str_detect(anno[["geneSymbol"]],"ROC40"),"gene_id"]
plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])

# search genes
anno[str_detect(anno[["geneSymbol"]],"CRY"),]
anno[str_detect(anno[["geneSymbol"]],"ROC"),]
anno[str_detect(anno[["previousIdentifiers"]],"CRY"),]

###############
# save dds #####

# save(dds, file = paste(dir,"dds_23design.RDS",sep="/"), compress = FALSE)
# save(dds, file = paste(dir,"dds.RDS",sep="/"), compress = FALSE)
# dds <- readRDS(paste(outdir,"dds.RDS",sep="/"))
# dds

```

### 2024-04-09 1_data_processing/1_data_processing2.Rmd
deseq2

--- 2024-04-09 1_data_processing/1_data_processing.Rmd
@@
 knit: (function(input_file, encoding) {
+    rmarkdown::render(input_file,output_file= 'README.md')
     })
-    rmarkdown::render(input_file,output_file= 'README.md')
 ---
@@
 ## -R_libraries
-```{r r_libraries}
+```{r library}
 library(stringr)
@@
 
-indexDir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")
+indexdir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")
 
@@
 head(readLines(fastaPath,n=50))
-
-gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
-readLines(gtfPath,n=20)
 
@@
 
-```{bash download data, eval=FALSE}
+```{bash, eval=FALSE}
 
@@
 ### check md5
-```{bash md5, eval=FALSE}
+```{bash}
 cd $fastqdir
@@
 ## 2a. Sample names
-```{r sample_table}
+```{r}
 # Get Sample table
@@
 Samplefile <- read_xlsx(Samplepath)
-samplename <- Samplefile$clientName
-Samplefile$treatment
-Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor() %>% relevel(ref="WT_dark")
-# levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red")
+Samplefile
+samplename <- Samplefile$'Sample name [user defined]'
+samplename
+genotype <- str_split(samplename, pattern="-", simplify = TRUE)[,1] %>% as.factor() %>% relevel(ref="WT")
+genotype
+
+light <- str_split(samplename, pattern="-", simplify = TRUE)[,2] %>% str_remove(pattern="\\d")
+# \\d = digits
+light <- str_split(samplename, pattern="-", simplify = TRUE)[,2] %>% str_remove_all(pattern="[0-9]") %>% as.factor() %>% relevel(ref="D")
+# [0-9]$ number at the end
+light
+
+repetition <- str_split(samplename, pattern="-", simplify = TRUE)[,2] %>% str_remove_all(pattern="[A-Z]") %>% as.factor()
+repetition
+
+RNA_conc <- Samplefile$'RNA conc. [ng/µl]'
+RNA_conc
+
+lane <- str_split(files,pattern ="_",simplify = T)[,5]
+lane
+
+condition <- paste(genotype, light, sep="_") %>% factor(levels = c("WT_D","pcry_D","WT_BL","pcry_BL","WT_R","pcry_R"))
+condition
 
 # Make table
-sample.table <- data.frame(Samplefile[,-3])
+sample.table <- data.frame(sample.id = sample, filenames = files, sample = samplename, genotype = genotype, treatment = light, condition = condition, repetition = repetition, lane = lane,RNA_conc=RNA_conc)
 sample.table
-
-sample.table <- data.frame(lapply(sample.table, factor))
-
-# Check the updated structure of the data frame
-str(sample.table)
-
 ```
@@
 
-```{r mapping_rates}
+```{r}
 library(stringr)
@@
 library(RColorBrewer)
+
+outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
+datadir <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044"
+quantdir <- paste(datadir,"quants",sep="/")
+
+list.files(path = outdir)
+list.files(path = datadir)
+list.files(path = quantdir)
 
@@
 }
-sample.table$mappingrates <- mappingrates
+
 # Make table
@@
 ### Plot mapping rates
-```{r plot_mappingrates}
+```{r}
 
@@
 # Plot
-par(mar=c(2,4,0,0)+.1)
-# plot(sample_table_all$mappingrates)
+plot(m.table$mappingrates)
 # -> boring
@@
 # increase margin for longer names
-par(mar=c(2,5,2,2)+.1)
-# barplot(height=mappingrates, names=samplename, horiz=T, las=1)
+par(mar=c(4,15,4,4)+.1)
+barplot(height=mappingrates, names=samplename, horiz=T, las=1)
+mkdirs("graphs")
 
-xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename, horiz=T, las=1, xlim=c(0,100))
-text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.5, col = "red")
-
-par(mar=c(4,4,4,4)+.1)
+pdf(file="graphs/Mapping_Rates.pdf", width=10, height=10)
+par(mar=c(4,12,4,4)+.1)
+xx <- barplot(height=mappingrates, names=samplename, horiz=T, las=1, xlim=c(0,100)) #col=col
+text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.8, col = "red")
+dev.off()
 
@@
 
-```{r tximeta}
+```{r}
 library(sessioninfo)
@@
 getwd()
+outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
+datadir <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044"
+quantdir <- paste(datadir,"quants",sep="/")
 example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")
@@
 
-dirfiles <- dirname(files)
-head(dirfiles)
-head(basename(dirfiles))
-run <- basename(dirfiles)
-dirdir <- dirname(dirfiles)
+dir <- dirname(files)
+head(dir)
+head(basename(dir))
+run <- basename(dir)
+dirdir <- dirname(dir)
 head(dirdir)
@@
 coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
-coldata <- data.frame(coldata,sample.table)
-coldata %>% head()
+coldata <- data.frame(coldata,m.table)
+coldata
+colnames(coldata)
+head(coldata$genotype)
+condition
 
@@
 ## Chlamy
+dir1 <- "S:/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy"
+
+indexDir <- file.path(dir1,"chlamy_index_v6.1")
+fastaPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
+head(readLines(fastaPath,n=50))
+gtfPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.repeatmasked_assembly_v6.0.gff3.gz")
+readLines(gtfPath,n=20)
+gtfPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene_exons.gff3.gz")
+readLines(gtfPath,n=20)
+gtfPath <- file.path(dir1,"Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
+readLines(gtfPath,n=20)
+
 file.exists(indexDir, fastaPath, gtfPath)
@@
 # gene_name info sind in gff3 abder nicht in se?
-coldata_Cre <- filter(coldata, species == "Cre")
-se <- tximeta(coldata_Cre, useHub=T)
+
+se <- tximeta(coldata, useHub=F)
 se
@@
 # rename Samples
-rownames(colData(se)) <- colData(se)$clientName
+rownames(colData(se)) <- colData(se)$sample.id
 
@@
 barplot(metadata(se)[["quantInfo"]]$num_processed/1000000, main="Mio. Reads")
-par(mfrow=c(1,1))
 
@@
 getwd()
-# save(gse,file=paste(dir,"tximeta.txm", sep="/"))
-# gse <- 1
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
+save(gse,file=paste(outdir,"tximeta.txm", sep="/"))
+gse <- 1
+load(file=paste(outdir,"tximeta.txm", sep="/"))
 
 ```
-
-### -Add gene info from Master annotation
-
-Use other files from Phytozome database: CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv
-
-```{r anno}
-library(data.table)
-library(plyr)
-library(tidyverse)
-library(ape)
-
-# more gene info from master_annotation_table
-annopath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")
-
-anno <- read.delim2(annopath, header = T, sep = "\t",)
-head(anno)
-
-# search for genes
-anno[str_detect(anno[["locusName_4532"]],"Cre01.g000050"),]
-anno[str_detect(anno[["geneSymbol"]],"HKR"),]
-anno[str_detect(anno[["previousIdentifiers"]],"HKR"),]
-anno[str_detect(anno[["geneSymbol"]],"COP"),]
-anno[str_detect(anno[["previousIdentifiers"]],"COP1"),]
-anno[str_detect(anno[["geneSymbol"]],"cop"),]
-anno[str_detect(anno[["geneSymbol"]],"PLAP6"),]
-anno[str_detect(anno[["geneSymbol"]],"COQ3"),]
-anno[str_detect(anno[["geneSymbol"]],"PCRY"),]
-anno[str_detect(anno[["geneSymbol"]],"CRY"),]
-anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
-anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH"),]
-
-dim(anno)
-
-# Remove "_4532"
-anno$gene_id <- str_remove(anno$locusName_4532,pattern = "_4532")
-rownames(anno) <- anno$gene_id
-
-# convert previousIdentifiers to list
-anno$previousIdentifiers_list <- strsplit(anno$previousIdentifiers, "#")
-head(anno)
-anno$previousIdentifiers_list[1]
-
-# polish "previousIdentifiers"
-# remove Cre & g-numbers
-head(anno$previousIdentifiers)
-tail(anno$previousIdentifiers)
-prev.symbols <-  
-  str_replace_all(anno$previousIdentifiers,pattern="Cre\\d+.g\\d+.t\\d+.\\d+", replacement="?") %>% 
-  str_replace_all(pattern="Cre...g\\d+", replacement="?") %>% 
-  str_replace_all(pattern="g\\d+.t\\d+", replacement="?") %>% 
-  str_remove_all(pattern="\\?") %>%
-  str_remove_all("^#|#$| ") %>% str_remove_all("^#|#$") %>% str_remove_all("^#|#$|^0") %>% 
-  str_replace_all(pattern="##","#") %>% str_replace_all(pattern="##","#")
-prev.symbols[str_detect(prev.symbols,"#")] %>% head(n=20)
-prev.symbols[str_detect(prev.symbols,"0")] %>% head(n=20)
-prev.symbols[str_detect(prev.symbols,"Cre")] %>% head(n=20)
-tail(anno$previousIdentifiers[str_detect(anno$previousIdentifiers,"Cre")])
-
-anno$prev.symbols <- prev.symbols
-head(anno)
-
-# Make colum with gene symbol or prev.symbol or Gene id
-anno$combname <- anno$geneSymbol
-
-summary(anno$combname)
-summary(anno$combname=="")
-summary(anno$combname==" ")
-summary(is.na(anno$combname))
-summary(anno$combname=="0")
-
-# replace "" with prev.symbol
-summary(anno$combname=="")
-  length(anno$combname[anno$combname==""])
-  length(str_split(
-  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1])
-anno$combname[anno$combname==""] <- str_split(
-  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1]
-summary(anno$combname=="")
-
-anno$combname[anno$combname==""] <- anno$gene_id[anno$combname==""]
-summary(anno$combname=="")
-subset(anno,anno$combname=="")
-
-summary(duplicated(anno$combname))
-# some duplicates
-
-colnames(anno)[which(names(anno) == "combname")] <- "id.symbol"
-
-summary(anno$geneSymbol!="")
-# 4408 gene Symbols
-summary(anno$previousIdentifiers!="")
-# all (!) 17712 genes have "previousIdentifiers"
-summary(anno$prev.symbols!="")
-# 7236 have alternative/old symbols
-summary(anno$id.symbol)
-
-dim(anno)
-
-anno %>% head()
-
-# saveRDS(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
-# anno <- readRDS(paste(outdir,"anno.RDS",sep="/"))
-
-```
-
-### -add anno data into gse ####
-```{r anno_gse}
-
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-gse
-anno <- readRDS(file=paste(dir,"/anno.RDS", sep="/"))
-anno %>% head()
-
-# Remove "_4532"
-mcols(gse)$gene_id2 <- mcols(gse)[,"gene_id"]
-mcols(gse)$gene_id <- str_remove(mcols(gse)[,"gene_id2"], pattern = "_4532")
-rownames(gse) <- mcols(gse)$gene_id
-
-summary(anno$gene_id %in% mcols(gse)$gene_id)
-summary(mcols(gse)$gene_id %in% anno$gene_id)
-
-# choose info
-gsetable <- as.data.frame(mcols(gse)[,c("gene_id","gene_id2")])
-
-anno_join <- anno[,c("gene_id", "geneSymbol", "id.symbol","prev.symbols", "previousIdentifiers", "previousIdentifiers_list", "Description", "Comments", "TargetP", "Predalgo")]
-anno_join
-
-class(anno_join)
-dim(anno_join)
-
-# any duplicated:
-summary(duplicated(anno$gene_id))
-
-new <- plyr::join(gsetable,anno_join, by = "gene_id", type = "left")
-dim(new)
-head(new)
-new2 <- left_join(gsetable,anno_join,by = "gene_id")
-dim(new2)
-head(new2)
-
-mcols(gse) <- left_join(as.data.frame(mcols(gse)),anno_join,by = "gene_id")
-mcols(gse)
-
-getwd()
-
-# save(gse,file=paste(dir,"tximeta.txm",sep="/"))
-```
-
-## 2c. DESeq2 Analysis
-
-```{r deseq2}
-getwd()
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-
-library(DESeq2)
-colData(gse)
-mcols(gse)
-gse$condition
-
-design <- ~condition
-# design <- ~genotype+treatment+genotype:treatment
-
-dds <- DESeqDataSet(gse, design=design)
-colData(dds) %>% head()
-
-##########################################\n
-## filter all rows with rowsum = 0 ####
-par(mfrow=c(1,2))
-hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
-hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
-# -> dip at log(counts(dds)=2-3)
-par(mfrow=c(1,1))
-
-# min counts
-# at least 5 counts in 3 samples
-keep.sn <- rowSums(counts(dds) >= 5) >= 3
-keep.sn %>% summary()
-dds <- dds[keep.sn,]
-
-dds <- estimateSizeFactors(dds)
-dds
-head(assays(dds)[["counts"]])[1:5,1:5]
-length(rownames(counts(dds)))
-
-colData(dds)
-# add metadata
-# colData(dds)$strain <- factor(colData(dds)$strain)
-
-# ## combine technical replicates #####
-# colData(dds)
-# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
-# colData(dds)
-# dds$runsCollapsed
-# colnames(dds)
-# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
-# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))
-
-###########
-# run DESeq
-dds <- DESeq(dds)
-DESeq2::plotMA(dds)
-
-plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition")
-
-head(mcols(dds)$geneSymbol)
-
-anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
-# no hit
-g <- anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH1"),"gene_id"]
-g
-plotCounts(dds, gene = g, intgroup = "condition")
-g <- anno[str_detect(anno[["geneSymbol"]],"PCRY"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ACRY"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ROC15"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ROC40"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-
-# search genes
-anno[str_detect(anno[["geneSymbol"]],"CRY"),]
-anno[str_detect(anno[["geneSymbol"]],"ROC"),]
-anno[str_detect(anno[["previousIdentifiers"]],"CRY"),]
-
-###############
-# save dds #####
-
-# save(dds, file = paste(dir,"dds_23design.RDS",sep="/"), compress = FALSE)
-# save(dds, file = paste(dir,"dds.RDS",sep="/"), compress = FALSE)
-# dds <- readRDS(paste(outdir,"dds.RDS",sep="/"))
-# dds
-
-```

### 2024-04-09 1_data_processing/1_data_processing.Rmd
acry basic finished

@@
 Samplefile$treatment
-Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor() %>% relevel(ref="WT_dark")
-# levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red")
-
+Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor(levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red","WT_3HL","WT_16HL","cia5_16HL"))
+Samplefile$treatment <- Samplefile$treatment %>% factor(levels = c("dark","blue","red","light","3HL","16HL"))
 # Make table
@@
 
-anno %>% head()
+# anno[1:10,1:10]
 
@@
 anno <- readRDS(file=paste(dir,"/anno.RDS", sep="/"))
-anno %>% head()
+# anno[1:10,1:10]
 
@@
 ```
+
+# 3. Pre-Analysis
+### - Data transformations
+```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
+# load(file=paste(data,"deseq2.dds", sep="/"))
+vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
+ntd <- normTransform(dds)
+rld <- rlog(dds, blind=FALSE) #regularized logarithm
+```
+
+```{r pre_trans_fig, figures-side, fig.show="hold", out.width="33%"}
+# load(file=paste(data,"rlog.rld", sep="/"))
+meanSdPlot(assay(ntd))
+meanSdPlot(assay(vsd))
+meanSdPlot(assay(rld))
+
+```
+
+### - Check sample distance
+```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
+sampleDists <- dist(t(assay(vsd)))
+
+sampleDistMatrix <- as.matrix(sampleDists)
+rownames(sampleDistMatrix) <- vsd$clientName
+# colnames(sampleDistMatrix) <- NULL
+colData(vsd)$treatment
+colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype")])
+anno_colors <- list(treatment = c("grey20","skyblue1","lightcoral","ivory","yellow1","gold"),
+                    genotype = c("grey","seagreen3","tan2"))
+
+names(anno_colors$treatment) <- levels(anno_col$treatment)
+names(anno_colors$genotype) <- levels(anno_col$genotype)
+
+pheatmap(sampleDistMatrix,
+         clustering_distance_rows=sampleDists,
+         clustering_distance_cols=sampleDists,
+         annotation_col=anno_col,
+         annotation_colors = anno_colors,
+         show_colnames     = FALSE,
+         col=viridis(20),
+         cutree_rows = 8,
+         cutree_cols = 8,
+         fontsize_row = 5)
+
+```
+
+### - Perform principal component analysis
+```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
+# transform data
+# load(file=paste(data,"deseq2.dds", sep="/"))
+vst_dat <- assay(vst(dds))
+
+### PCA with top 500 genes with highest row variance 
+pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE)
+# levels(pcaData$genotype)
+percentVar <- round(100 * attr(pcaData, "percentVar"))
+g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "treatment") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
+  scale_color_manual(values = anno_colors$treatment)
+
+g2 <- ggplot(pcaData, aes(PC1, PC2, color=genotype, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "genotype") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  scale_color_manual(values = anno_colors$genotype)
+
+g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "condition") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  # coord_fixed()
+  scale_color_viridis_d(option ="viridis")
+
+((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")
+
+```
+
+###### -- Advanced PCA
+```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
+# transform data
+# calculate PCA (all data)
+## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots
+
+# p <- pca(vst_dat, metadata = colData(dds))
+# vst <- assay(vst(dds))
+p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)
+
+# check different PCAs
+p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment
+
+p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype
+
+p4 <- pairsplot(p,colby = 'condition', colkey = viridis(9),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition
+
+(p1+p2) / (p4)
+
+# Determine optimum number of PCs to retain
+elbow <- findElbowPoint(p$variance)
+elbow
+
+# plot Scree
+# find explaining PCs
+# horn <- parallelPCA(vst_dat)
+# horn$n = 4
+# warnings()
+
+horn <- list()
+horn$n <- 4
+
+screeplot(p,
+    components = getComponents(p, 1:20),
+    vline = c(horn$n, elbow)) +
+    geom_label(aes(x = horn$n + 1, y = 30,
+      label = 'Horn\'s=4', vjust = -1, size = 8)) +
+    geom_label(aes(x = elbow + 1, y = 50,
+      label = 'Elbow=5', vjust = -1, size = 8))
+
+bi <- biplot(p,x="PC3",y="PC1",
+    lab = p$metadata$condition,
+    colby = 'condition',colkey = viridis(9),
+    hline = 0, vline = 0,
+    encircle = TRUE, encircleFill = TRUE,
+    labSize = 3, legendIconSize = 4.0,
+    legendPosition = 'bottom',
+    sizeLoadingsNames = 3,
+    axisLabSize = 10,
+    captionLabSize = 1)
+
+bi <- bi +theme(panel.background = element_rect(fill='transparent'))
+
+pairs <- pairsplot(p,
+    components = getComponents(p, c(1:7)),
+    triangle = TRUE, trianglelabSize = 6,
+    hline = 0, vline = 0,
+    pointSize = 2,
+    gridlines.major = FALSE, gridlines.minor = FALSE,
+    colby = 'condition', colkey = viridis(9),
+    plotaxes = FALSE,
+    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))
+
+layout <- c(
+  area(t = 1, l = 1, b = 6, r = 6),
+  area(t = 5, l = 1, b = 7, r = 3)
+)
+pairs + bi + 
+  plot_layout(design = layout)
+
+# pairs + inset_element(bi, left = 0, bottom = 0, right = 0.5, top = 0.5)
+
+plotloadings(p,
+             components = getComponents(p)[1:7],
+             col = viridis(3),
+             labSize = 3)
+
+```
+
+### - Plot example counts
+```{r example_counts, fig.show="hold", out.width="50%", eval=TRUE, include=TRUE, echo=FALSE}
+goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "DCRY2", "PHOT1", "UVR8", "HKR1") #"HKR2"
+
+anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
+rownames(anno_phot) <- anno_phot$geneSymbol
+anno_phot <- anno_phot[goi_phot,]
+
+rownames(anno_phot) <- anno_phot$geneSymbol
+anno_phot <- anno_phot[goi_phot,]
+
+# Get Count data
+i <- 1
+l <- length(goi_phot)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=anno_phot[i,"gene_id"], intgroup=c("condition","genotype","treatment"), main=anno_phot[i,"id.symbol"],returnData=TRUE)
+  d$Gene <- rep(anno_phot[i,"id.symbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  # rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+# TF genes
+
+goi_tf <- c("CCM1","LCR1", "HY5", "PHOT1", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
+anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
+anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
+anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
+summary(anno_tf$gene_id %in% rownames(dds))
+anno_tf[!anno_tf$gene_id %in% rownames(dds),]
+anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]
+
+# Combined counts table TFs
+
+goi <- anno_tf
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","treatment","genotype"),col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+all_counts$Gene <- factor(all_counts$Gene, levels = anno_tf$geneSymbol[c(9,6,8,7,15,2,1,4,3,5,13,11,14,10,12)])
+levels(all_counts$Gene)
+
+# colours
+# group.colors <- rep(anno_colors$treatment, each = 2)
+group.colors <- rep(anno_colors$treatment, each = 2)[c(1:6,7,9,11)]
+names(group.colors) <- colData(dds)$condition %>% levels()
+anno_colors$condition <- group.colors
+
+max_val <- 1.0*max(all_counts$count)
+
+# Plot
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  scale_y_continuous(trans = "log2")
+gcounts
+
+cols = brewer.pal(n=9,name = 'Paired')
+
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition, color=condition)) +
+  geom_boxplot(color="black") +
+  geom_point(position=position_dodge(width=0.75), alpha=1) +
+  scale_fill_manual(values=cols) +
+  scale_color_manual(values=cols) +
+  scale_y_continuous(trans = "log2")
+gcounts
+
+gcounts2 <- ggplot(all_counts, aes(x = condition, y = count, fill=Gene)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_viridis_d() +
+  scale_y_continuous(trans = "log2")
+gcounts2
+
+gcounts3 <- ggplot(all_counts[all_counts$Gene=="ROC40",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
+  geom_boxplot(color="black") +
+  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
+  scale_fill_manual(values=anno_colors$treatment) +
+  scale_color_manual(values=anno_colors$treatment) +
+  scale_y_continuous(trans = "log2") +
+  labs(title = "ROC40") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+gcounts3
+
+gcounts4 <- ggplot(all_counts[all_counts$Gene=="ROC114",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
+  geom_boxplot(color="black") +
+  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
+  scale_fill_manual(values=anno_colors$treatment) +
+  scale_color_manual(values=anno_colors$treatment) +
+  scale_y_continuous(trans = "log2") +
+  labs(title = "ROC114") +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
+gcounts4
+
+```
+
+```{r}
+sessionInfo()
+```

### 2024-04-09 1_data_processing/1_data_processing2.Rmd
acry basic finished

--- 2024-04-09 1_data_processing/1_data_processing.Rmd
@@
 ```
+
+# 3. Pre-Analysis
+### - Data transformations
+```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
+# load(file=paste(data,"deseq2.dds", sep="/"))
+vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
+ntd <- normTransform(dds)
+```
+
+#### -#rlog
+```{r pre_trans_rlog, eval=FALSE, include=FALSE, echo=FALSE}
+rld <- rlog(dds, blind=FALSE) #regularized logarithm
+# save(rld,file=paste(dir,"rlog.rld", sep="/"))
+# rld <- 1
+# load(file=paste(data,"rlog.rld", sep="/"))
+rld
+```
+
+```{r pre_trans_fig, figures-side, fig.show="hold", out.width="33%"}
+# load(file=paste(data,"rlog.rld", sep="/"))
+meanSdPlot(assay(ntd))
+meanSdPlot(assay(vsd))
+meanSdPlot(assay(rld))
+
+```
+
+### - Check sample distance
+```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
+sampleDists <- dist(t(assay(vsd)))
+
+sampleDistMatrix <- as.matrix(sampleDists)
+rownames(sampleDistMatrix) <- vsd$names
+# colnames(sampleDistMatrix) <- NULL
+colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
+anno_col <- as.data.frame(colData(vsd)[,c("experiment","treatment","genotype")])
+anno_colors <- list(treatment = c("lightcoral","skyblue1"),
+                    genotype = c("grey","seagreen3","turquoise3","tan2"),
+                    experiment = viridis(4, option="plasma"))
+
+names(anno_colors$treatment) <- levels(anno_col$treatment)
+names(anno_colors$genotype) <- levels(anno_col$genotype)
+names(anno_colors$experiment) <- levels(anno_col$experiment)
+
+pheatmap(sampleDistMatrix,
+         clustering_distance_rows=sampleDists,
+         clustering_distance_cols=sampleDists,
+         annotation_col=anno_col,
+         annotation_colors = anno_colors,
+         show_colnames     = FALSE,
+         col=viridis(20),
+         cutree_rows = 8,
+         cutree_cols = 8,
+         fontsize_row = 5)
+
+```
+
+### - Perform principal component analysis
+```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
+# transform data
+# load(file=paste(data,"deseq2.dds", sep="/"))
+vst_dat <- assay(vst(dds))
+# colData(dds) %>% head() %>% kable() %>% kable_styling("striped", full_width = T) %>% scroll_box(height = "400px")
+
+ens <- rownames(dds)
+  symbols <- mcols(dds)$SYMBOL
+  symbols <- symbols[!is.na(symbols)]
+  symbols <- symbols[match(rownames(vst_dat), names(symbols))]
+  rownames(vst_dat) <- symbols
+  keep <- !is.na(rownames(vst_dat))
+  vst_dat <- vst_dat[keep,]
+  keep2 <- rownames(vst_dat)
+  keep2 <- !duplicated(keep2)
+  vst_dat <- vst_dat[keep2,]
+
+### PCA with top 500 genes with highest row variance 
+pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE)
+# levels(pcaData$genotype)
+percentVar <- round(100 * attr(pcaData, "percentVar"))
+g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "treatment") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
+  scale_color_manual(values = c("lightcoral","skyblue1"))
+
+g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "experiment") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  scale_color_viridis_d(option ="viridis")
+
+g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "condition") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  # coord_fixed()
+  scale_color_viridis_d(option ="viridis")
+
+g4 <- ggplot(pcaData, aes(PC1, PC2, label=orig.name,color=condition, shape=genotype)) +
+  geom_text_repel(data         = subset(pcaData, experiment == "Simon"),
+                  segment.color = 'grey50',
+                  max.overlaps = 40,
+                  color="grey30",
+                  size          = 2.5) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "top 500 variance") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  # coord_fixed()
+  scale_color_viridis_d(option ="viridis")
+
+((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")
+
+```
+
+###### -- Advanced PCA
+```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
+# transform data
+# calculate PCA (all data)
+## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots
+
+# p <- pca(vst_dat, metadata = colData(dds))
+# vst <- assay(vst(dds))
+p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)
+
+# check different PCAs
+p1 <- pairsplot(p,colby = 'treatment', colkey = viridis(2),title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1 = treatment
+p2 <- pairsplot(p,colby = 'genotype', colkey = viridis(4),title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC1&3 = genotype
+p3 <- pairsplot(p,colby = 'experiment', colkey = viridis(4),title = 'experiment',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # -> PC4 = experiment
+p4 <- pairsplot(p,colby = 'condition', colkey = viridis(8),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
+                hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
+                plotaxes = FALSE,margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm')) # PC1 & PC3 = condition
+
+(p1+p2) / (p3+p4)
+
+# Determine optimum number of PCs to retain
+elbow <- findElbowPoint(p$variance)
+elbow
+
+# plot Scree
+# find explaining PCs
+# horn <- parallelPCA(vst_dat)
+# horn$n = 7
+# warnings()
+
+horn <- list()
+horn$n <- 7
+
+screeplot(p,
+    components = getComponents(p, 1:20),
+    vline = c(horn$n, elbow)) +
+    geom_label(aes(x = horn$n + 1, y = 30,
+      label = 'Horn\'s=7', vjust = -1, size = 8)) +
+    geom_label(aes(x = elbow + 1, y = 50,
+      label = 'Elbow=6', vjust = -1, size = 8))
+
+bi <- biplot(p,x="PC3",y="PC1",
+    lab = p$metadata$experiment,
+    colby = 'condition',colkey = viridis(8),
+    hline = 0, vline = 0,
+    encircle = TRUE, encircleFill = TRUE,
+    labSize = 3, legendIconSize = 4.0,
+    legendPosition = 'bottom',
+    sizeLoadingsNames = 3,
+    axisLabSize = 10,
+    captionLabSize = 1)
+
+bi <- bi +theme(panel.background = element_rect(fill='transparent'))
+
+pairs <- pairsplot(p,
+    components = getComponents(p, c(1:7)),
+    triangle = TRUE, trianglelabSize = 6,
+    hline = 0, vline = 0,
+    pointSize = 2,
+    gridlines.major = FALSE, gridlines.minor = FALSE,
+    colby = 'condition', colkey = viridis(8),
+    plotaxes = FALSE,
+    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))
+
+layout <- c(
+  area(t = 1, l = 1, b = 6, r = 6),
+  area(t = 5, l = 1, b = 7, r = 3)
+)
+pairs + bi + 
+  plot_layout(design = layout)
+
+# pairs + inset_element(bi, left = 0, bottom = 0, right = 0.5, top = 0.5)
+
+plotloadings(p,
+             components = getComponents(p)[1:7],
+             col = viridis(3),
+             labSize = 3)
+
+```
+
+###### -- #PCA gif
+```{r pca_gif, fig.height=8, fig.show="hold", warning=FALSE, out.width="100%", eval=FALSE, include=FALSE, echo=FALSE}
+
+saveGIF({
+  ani.options(nmax = 100)
+for (i in 1:100){
+  n <- i*310
+pcaData <- plotPCA(vsd, intgroup=colnames(colData(vsd)), returnData=TRUE,ntop=n)
+percentVar <- round(100 * attr(pcaData, "percentVar"))
+g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "treatment") +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
+  coord_cartesian(xlim = c(-100, 100),ylim = c(-100,100)) +
+  scale_color_manual(values = c("lightcoral","skyblue1"))
+
+g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "experiment") +
+  coord_cartesian(xlim = c(-100, 100),ylim = c(-100,100)) +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  scale_color_viridis_d(option ="viridis")
+
+g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
+  geom_point(size=5, alpha=0.7) +
+  labs(title = "condition") +
+  coord_cartesian(xlim = c(-100, 100),ylim = c(-100,100)) +
+  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
+  ylab(paste0("PC2: ",percentVar[2],"% variance")) +
+  # coord_fixed()
+  scale_color_viridis_d(option ="viridis")
+plot((g1+g2)/g3 +
+       plot_annotation(title = paste("number of variable genes: ",n)) +
+       plot_layout(guides = 'collect')
+     )
+ani.pause()
+}
+}, interval = 0.3, movie.name = file.path(gitdir,"1_data_processing/pca.gif"), ani.width = 600, ani.height = 600)
+
+```
+
+[PCA Gif](pca.gif){height=100%, width=100%}
+
+```{r pca_gif_include, eval=TRUE, echo=FALSE, out.width="100%", fig.height=8}
+knitr::include_graphics("pca.gif")
+
+```
+
+### - Plot example counts
+```{r example_counts, fig.show="hold", out.width="50%", eval=TRUE, include=TRUE, echo=FALSE}
+# CA9 & EPO
+# load(file=paste(data,"deseq2.dds", sep="/"))
+goi <- c("CA9","EPO")
+# (mcols(dds)$SYMBOL == "CA9") %>% summary()
+
+mcols(dds)$gene_id %>% length()
+mcols(dds)$SYMBOL %>% length()
+SYMBOLs <- mcols(dds)$SYMBOL %>% unique() %>% na.omit()
+SYMBOLs %>% length()
+SYMBOLs <- SYMBOLs[!(SYMBOLs == "")]
+SYMBOLs %>% length()
+ENSEMBLS <- as.character(mapIds(edb,keys = SYMBOLs, column = "GENEID", keytype = "SYMBOL", multiVals="first"))
+ENTREZ <- as.character(mapIds(edb,keys = SYMBOLs, column = "ENTREZID", keytype = "SYMBOL", multiVals="first"))
+s2e <- data.frame(symbol=SYMBOLs,
+                  ensemble=ENSEMBLS,
+                  entrez=ENTREZ)
+rownames(s2e) <- s2e$symbol
+# s2e %>% head() %>% kable() %>% kable_styling("striped", full_width = T) %>% scroll_box(height = "400px")
+
+# s2e[goi[1],2]
+par(mar=c(2,2,2,2)+.1)
+# plotCounts(dds,gene=s2e[goi[1],2],intgroup="condition",main=goi[1],col=dds$experiment)
+# plotCounts(dds,gene=s2e[goi[2],2],intgroup="condition",main=goi[2],col=dds$experiment)
+
+i <- 1
+l <- length(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=s2e[goi[i],2], intgroup=c("condition","experiment","genotype","treatment"), main=goi[i],returnData=TRUE)
+  d$Gene <- rep(goi[i],length(rownames(d)))
+  d$sample <- rownames(d)
+  # rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+cols = brewer.pal(n=8,name = 'Paired')
+
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition, color=condition)) +
+  geom_boxplot(color="black") +
+  geom_point(position=position_dodge(width=0.75), alpha=1) +
+  scale_fill_manual(values=cols) +
+  scale_color_manual(values=cols) +
+  scale_y_continuous(trans = "log2")
+
+gcounts1 <- ggplot(d, aes(x = condition, y = count, color=experiment)) +
+  geom_point() +
+  geom_text(label = d$sample) +
+  scale_fill_manual(values=cols) +
+  scale_y_continuous(trans = "log2")
+
+gcounts2 <- ggplot(all_counts, aes(x = condition, y = count, fill=Gene)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values=cols) +
+  scale_y_continuous(trans = "log2")
+
+gcounts3 <- ggplot(all_counts[all_counts$Gene=="CA9",], aes(x = genotype, y = count, fill=treatment, color=treatment)) +
+  geom_boxplot(color="black") +
+  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
+  scale_fill_manual(values=cols[c(1,5)]) +
+  scale_color_manual(values=cols[c(1,5)]) +
+  scale_y_continuous(trans = "log2") +
+  labs(title = "CA9")
+gcounts3
+
+gcounts4 <- ggplot(all_counts[all_counts$Gene=="EPO",], aes(x = genotype, y = count, fill=treatment, color=treatment)) +
+  geom_boxplot(color="black") +
+  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
+  scale_fill_manual(values=cols[c(1,5)]) +
+  scale_color_manual(values=cols[c(1,5)]) +
+  scale_y_continuous(trans = "log2") +
+  labs(title = "EPO")
+gcounts4
+
+```
+
+```{r}
+sessionInfo()
+```

### 2024-04-16 1_data_processing/1_data_processing.Rmd
- separate dds for cia5 & acry - Results Cia5

@@
 library(curl)
+library(DESeq2)
 
@@
 library(vsn)
-
+library(stringr)
+library(R.utils)
+library(RColorBrewer)
 library(PCAtools)
@@
 fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
-head(readLines(fastaPath,n=50))
+head(readLines(fastaPath,n=10))
 
 gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
-readLines(gtfPath,n=20)
+readLines(gtfPath,n=10)
 
@@
 sample.table <- data.frame(Samplefile[,-3])
-sample.table
+sample.table %>% kable()
 
@@
 ## 2b. Mapping Rates
-
 ```{r mapping_rates}
-library(stringr)
-library(R.utils)
-library(RColorBrewer)
 
@@
 
-# Colours
-
 # Plot
-par(mar=c(2,4,0,0)+.1)
-# plot(sample_table_all$mappingrates)
+par(mar=c(2,4,0,0)+.1) # set margins for figure
+plot(sample_table_all$mappingrates)
 # -> boring
@@
 files <- c(files,paste(i,"/quant.sf",sep=""))
-print(basename(i))
+# print(basename(i))
 # print(head(read.table(files[length(files)], header=T)))
-print (files[length(files)])
+# print (files[length(files)])
 }
@@
 coldata <- data.frame(coldata,sample.table)
-coldata %>% head()
+coldata %>% head() %>% kable()
 
@@
 gtfdata <- readLines(gtfPath)
-head(gtfdata, n=20)
+head(gtfdata, n=10)
 
@@
 
-colData(se)
+colData(se) %>% head() %>% kable()
 # meta infos sind da (genotype, treatment,...)
@@
 
-rowRanges(se)
+rowRanges(se) %>% kable()
 # genome info
@@
 # ?
-head(assays(se)[["counts"]])
+head(assays(se)[["counts"]]) %>% kable()
 # counts. THE DATA
@@
 names(metadata(se)[["quantInfo"]])
-str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
+# str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
 metadata(se)[["quantInfo"]]$percent_mapped
@@
 getwd()
-# save(gse,file=paste(dir,"tximeta.txm", sep="/"))
+# save(gse,file=paste(dir,"tximeta.RDS", sep="/"))
 # gse <- 1
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
+# load(file=paste(dir,"tximeta.RDS", sep="/"))
 
@@
 anno <- read.delim2(annopath, header = T, sep = "\t",)
-head(anno)
+head(anno)[1:5,1:9] %>% kable()
 
@@
 anno$prev.symbols <- prev.symbols
-head(anno)
+head(anno)[1:5,1:9] %>% kable()
 
@@
 
-# saveRDS(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
-# anno <- readRDS(paste(outdir,"anno.RDS",sep="/"))
+# save(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
+# load(paste(dir,"anno.RDS",sep="/"))
 
@@
 
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-gse
-anno <- readRDS(file=paste(dir,"/anno.RDS", sep="/"))
-# anno[1:10,1:10]
+# load(file=paste(dir,"tximeta.RDS", sep="/"))
+# load(file=paste(dir,"/anno.RDS", sep="/"))
+anno[1:3,1:8] %>% kable()
 
@@
 anno_join <- anno[,c("gene_id", "geneSymbol", "id.symbol","prev.symbols", "previousIdentifiers", "previousIdentifiers_list", "Description", "Comments", "TargetP", "Predalgo")]
-anno_join
+anno_join %>% head() %>% kable()
 
@@
 dim(new)
-head(new)
+head(new) %>% kable()
+
 new2 <- left_join(gsetable,anno_join,by = "gene_id")
 dim(new2)
-head(new2)
+head(new2) %>% kable()
 
 mcols(gse) <- left_join(as.data.frame(mcols(gse)),anno_join,by = "gene_id")
-mcols(gse)
-
-getwd()
-
-# save(gse,file=paste(dir,"tximeta.txm",sep="/"))
+mcols(gse) %>% head() %>% kable()
+
+save(gse,file=paste(dir,"tximeta.RDS",sep="/"))
 ```
@@
 getwd()
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-
-library(DESeq2)
-colData(gse)
-mcols(gse)
-gse$condition
-
+# load(file=paste(dir,"tximeta.RDS", sep="/"))
+
+colData(gse) %>% head() %>% kable()
+mcols(gse) %>% head() %>% kable()
+
+gse_acry <- gse[,colData(gse)$experiment=="acry"]
+colData(gse_acry) <- colData(gse_acry) %>% droplevels()
+colData(gse_acry)$genotype <- colData(gse_acry)$genotype %>% relevel(ref="WT")
+
+gse_cia5 <- gse[,colData(gse)$experiment=="cia5"]
+colData(gse_cia5) <- colData(gse_cia5) %>% droplevels()
+colData(gse_cia5)$genotype <- colData(gse_cia5)$genotype %>% relevel(ref="WT")
+
+```
+
+### Deseq2 acry
+```{r}
 design <- ~condition
@@
 
-dds <- DESeqDataSet(gse, design=design)
-colData(dds) %>% head()
+dds <- DESeqDataSet(gse_acry, design=design)
+colData(dds) %>% head() %>% kable()
 
@@
 
-colData(dds)
+colData(dds) %>% head() %>% kable()
 # add metadata
@@
 # search genes
-anno[str_detect(anno[["geneSymbol"]],"CRY"),]
-anno[str_detect(anno[["geneSymbol"]],"ROC"),]
-anno[str_detect(anno[["previousIdentifiers"]],"CRY"),]
+anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()
+
+anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
+anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()
 
@@
 # save(dds, file = paste(dir,"dds_23design.RDS",sep="/"), compress = FALSE)
-# save(dds, file = paste(dir,"dds.RDS",sep="/"), compress = FALSE)
-# dds <- readRDS(paste(outdir,"dds.RDS",sep="/"))
+# save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
+# load(paste(dir,"dds.RDS",sep="/"))
 # dds
@@
 
+### Deseq2 cia5
+```{r}
+design <- ~genotype
+dds <- DESeqDataSet(gse_cia5, design=design)
+colData(dds) %>% head() %>% kable()
+
+## filter all rows with rowsum = 0 ####
+par(mfrow=c(1,2))
+hist(log(counts(dds)), breaks=100, ylim = c(0,4000), xlim = c(0,10))
+hist(counts(dds), breaks=300000, ylim = c(0,10000), xlim = c(0,10))
+# -> dip at log(counts(dds)=2-3)
+par(mfrow=c(1,1))
+
+# min counts
+# at least 5 counts in 3 samples
+keep.sn <- rowSums(counts(dds) >= 5) >= 3
+keep.sn %>% summary()
+dds <- dds[keep.sn,]
+
+dds <- estimateSizeFactors(dds)
+dds
+head(assays(dds)[["counts"]])[1:5,1:5]
+length(rownames(counts(dds)))
+
+colData(dds) %>% head() %>% kable()
+
+# run DESeq
+dds <- DESeq(dds)
+
+# qc check
+DESeq2::plotMA(dds)
+plotDispEsts(dds)
+
+# counts
+plotCounts(dds, gene = "Cre01.g000150", intgroup = "genotype", col=colData(dds)$genotype)
+
+save(dds, file = paste(dir,"dds_cia5.RDS",sep="/"), compress = FALSE)
+
+```
+
 # 3. Pre-Analysis
+### load data
+```{r}
+load(file=paste(dir,"dds_cia5.RDS", sep="/"))
+load(file=paste(dir,"dds_acry.RDS", sep="/"))
+load(file=paste(dir,"anno.RDS", sep="/"))
+
+```
+
 ### - Data transformations
 ```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
-# load(file=paste(data,"deseq2.dds", sep="/"))
 vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
@@
 anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype")])
-anno_colors <- list(treatment = c("grey20","skyblue1","lightcoral","ivory","yellow1","gold"),
+anno_colors <- list(treatment = c("grey20","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
                     genotype = c("grey","seagreen3","tan2"))
@@
          col=viridis(20),
-         cutree_rows = 8,
-         cutree_cols = 8,
+         cutree_rows = 6,
+         cutree_cols = 6,
          fontsize_row = 5)
@@
 
-p4 <- pairsplot(p,colby = 'condition', colkey = viridis(9),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
+p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
                 hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
@@
 # horn <- parallelPCA(vst_dat)
-# horn$n = 4
+# horn$n = 2
 # warnings()
@@
 horn <- list()
-horn$n <- 4
+horn$n <- 2
 
@@
     geom_label(aes(x = horn$n + 1, y = 30,
-      label = 'Horn\'s=4', vjust = -1, size = 8)) +
+      label = 'Horn\'s=2', vjust = -1, size = 8)) +
     geom_label(aes(x = elbow + 1, y = 50,
-      label = 'Elbow=5', vjust = -1, size = 8))
+      label = 'Elbow=4', vjust = -1, size = 8))
 
@@
     lab = p$metadata$condition,
-    colby = 'condition',colkey = viridis(9),
+    colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
     hline = 0, vline = 0,
@@
     gridlines.major = FALSE, gridlines.minor = FALSE,
-    colby = 'condition', colkey = viridis(9),
+    colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),
     plotaxes = FALSE,
@@
 ### - Plot example counts
-```{r example_counts, fig.show="hold", out.width="50%", eval=TRUE, include=TRUE, echo=FALSE}
+#### cia5
+
+```{r example_counts, fig.show="hold", out.width="50%", eval=FALSE, include=TRUE, echo=FALSE}
+
+load(file=paste(dir,"dds_cia5.RDS", sep="/"))
+
+## cia5
+
+# plap6
+PLAP6 <- "Cre03.g188700"
+d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)
+
+resultsNames(dds)
+res <- results(dds,contrast = c("genotype","cia5","WT"))
+res["Cre03.g188700",]
+ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
+    geom_boxplot(color="black") +
+    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
+   # scale_fill_manual(values=cols[c(1,5)]) +
+  #  scale_color_manual(values=cols[c(1,5)]) +
+   scale_y_continuous(trans = "log2") +
+  scale_fill_manual(values = c("grey30","salmon")) + 
+  scale_color_manual(values = c("grey30","salmon")) + 
+    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
+  assign(paste("gcounts_",ig,sep=""),gcounts)
+
+  gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  scale_y_continuous(trans = "log2")
+gcounts
+# ggsave(paste(outdir,"graphs/2023_11_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
+#   width = 18,
+#   height = 10)
+
 goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "DCRY2", "PHOT1", "UVR8", "HKR1") #"HKR2"

### 2024-04-16 2_A_Results_cia5/2_results_cia5.Rmd
- separate dds for cia5 & acry - Results Cia5

--- 2024-04-09 1_data_processing/1_data_processing.Rmd
@@
 ---
-  title: "RNAseq aCRY processing"
+title: "RNAseq cia5 results"
 author: "Kelterborn"
-date: "2024-03-15"
+date: "2024-04-16"
 output:
@@
 
-## -Linux folders
-```{bash linux_folders, eval=FALSE}
-fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
-fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;
-
-dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY";
-
-quantdir="$dir/quants";
-
-indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";
+# 1. Load data
+```{r load_data}
+load(file=paste(dir,"dds_cia5.RDS", sep="/"))
+load(paste(dir,"anno.RDS",sep="/"))
 
@@
 
-# 1. Raw data processing
-## 1a. Download data (AWS CLI)
+### - Plot counts
+#### -- cia5 & rbcs
 
-The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
-Index in AWS: bmkdatarelease-37/delivery_20240313120107785
-Access key ID: AKIA45O7KHZWJCKW62VE
-Secret Access Key: 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
-Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
-Username:delivery_20240313120107785
-Password:Pwd2024@x4S416
+```{r example_counts, fig.show="hold", out.width="50%", eval=TRUE, include=TRUE, echo=FALSE}
 
-```{bash download data, eval=FALSE}
+## cia5
 
-# install aws
-curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
-unzip awscliv2.zip
-sudo ./aws/install
+# plap6
+PLAP6 <- "Cre03.g188700"
+d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)
+g1 <- ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
+    geom_boxplot(color="black") +
+    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
+   scale_y_continuous(limits = c(60,150),trans = "log2") +
+  scale_fill_manual(values = c("grey30","salmon")) + 
+    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
 
-aws --version
-
-aws configure
-AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
-AWS Secret Access Key [None]: 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
-Default region name [None]: eu-north-1
-Default output format [None]: table
-
-fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
-cd $fastq
-
-# test
-aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt
-
-# all files
-aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive
+# rbcl
+anno[str_detect(anno[["Description"]],"RuBisCO"),1:9]
+rbcl <- "CreCp.g802313"
+d <- plotCounts(dds,gene = rbcl, intgroup= c("treatment","genotype","condition"), returnData=TRUE)
+g2 <- ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
+    geom_boxplot(color="black") +
+    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
+   scale_y_continuous(trans = "log2") +
+  scale_fill_manual(values = c("grey30","salmon")) + 
+#  scale_color_manual(values = c("grey30","salmon")) + 
+    labs(title = paste("rbcl (",rbcl,")",sep=""))
+g1
+g2
+g1+g2 + plot_layout(guides = "collect", axes="collect", axis_titles="collect") & theme(legend.position = 'right')
 
@@
 
-### check md5
-```{bash md5, eval=FALSE}
-cd $fastqdir
-ls
-# 1 file
-md5sum Unknown_BU327-002T0008_1.fq.gz
-
-# Check downloads
-
-# Path to the text file containing the MD5 sums
-md5_file="data_md5.txt"
-
-# Function to check the MD5 sum of a file
-check_md5() {
-    file_name=$1
-    expected_sum=$2
-
-    md5_sum=$(md5sum "$file_name" | awk '{print $1}')
-    if [[ $md5_sum == $expected_sum ]]; then
-        echo "$file_name: MD5 sum is correct."
-    else
-        echo "$file_name: MD5 sum is incorrect."
-    fi
-}
-
-# Read the MD5 sums from the text file and check them
-while IFS= read -r line; do
-    file_name=$(echo "$line" | awk '{print $2}')
-    expected_sum=$(echo "$line" | awk '{print $1}')
-    if [[ -f $file_name ]]; then
-        check_md5 "$file_name" "$expected_sum"
-    else
-        echo "$file_name: File not found."
-    fi
-done < "$md5_file" | tee -a $log
+# 2. Results
+```{r results}
+resultsNames(dds)
+res <- results(dds,contrast = c("genotype","cia5","WT"))
+res["Cre03.g188700",]
 
@@
 
-## 1b. Fastqc
-
-## 1c. Mapping
-```{bash mapping, eval=FALSE}
-mamba activate salmon;
-
-fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
-fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);
-
-# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
-fastqfiles_all=$fastqfiles;
-
-echo "${fastqfiles:0:100}";
-echo "${fastqfiles2:0:100}";
-echo "${fastqfiles_all:0:100}";
-
-# files: Unknown_BU327-002T0001_1.fq.gz;
-for fn in $fastqfiles_all;
-do
-bfn=$(echo "`basename ${fn}`");
-samp=${bfn%"_1.fq.gz"};
-R1=$fn;
-R2=$(echo "$R1" | sed 's/_1/_2/');
-echo "Processing Sample: $samp";
-test -f $R1 && echo "--> File: $R1 exists"
-test -f $R2 && echo "--> File: $R2 exists"
-outquant=${quantdir}/${samp}_quant
-echo $outquant
-echo $indexdir
-
-salmon quant -i $indexdir -l A \
-  -1 $R1 \
-  -2 $R2 \
-  -p 25 --validateMappings --gcBias -o $outquant
-
-done
-
+```{r}
+sessionInfo()
 ```
-
-# 2. Process Data
-## 2a. Sample names
-```{r sample_table}
-# Get Sample table
-# Filenames
-
-quantdir <- paste(dir,"quants",sep="/")
-
-f <- list.files(path = quantdir)
-files <- as.factor(str_remove(f,pattern ="_quant"))
-# short form
-sample <- str_remove(files,pattern ="Unknown_")
-sample
-# condition
-library(readxl)
-Samplepath <- filePath(path=dir, file="sample_table.xlsx")
-Samplefile <- read_xlsx(Samplepath)
-samplename <- Samplefile$clientName
-Samplefile$treatment
-Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor() %>% relevel(ref="WT_dark")
-# levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red")
-
-# Make table
-sample.table <- data.frame(Samplefile[,-3])
-sample.table
-
-sample.table <- data.frame(lapply(sample.table, factor))
-
-# Check the updated structure of the data frame
-str(sample.table)
-
-```
-
-## 2b. Mapping Rates
-
-```{r mapping_rates}
-library(stringr)
-library(R.utils)
-library(RColorBrewer)
-
-samplelist <- {}
-mappingrates <- {}
-for (i in list.files(path = quantdir)){
-print(i)
-si <- paste("",str_sub(i,11,-7),sep = "")
-si
-samplelist <- c(samplelist,si)
-f <- readLines(paste(quantdir,i,"logs/salmon_quant.log", sep="/"))
-line <- grep("Mapping rate = ",f,value=TRUE)
-sl <- str_length(line)
-sl
-notime <- substring(line,30,sl)
-notime
-manual <- substring(line,sl-7,sl-1)
-val <- as.numeric(str_extract(notime,"[0-9.]+"))
-val
-valr<-round(val, digits=2)
-print(paste("Mapping rate of ",si," is: ",valr," %"))
-mappingrates <- c(mappingrates,valr)
-}
-sample.table$mappingrates <- mappingrates
-# Make table
-
-m.table <- data.frame(sample.table,mappingrates)
-```
-
-### Plot mapping rates
-```{r plot_mappingrates}
-
-# Colours
-
-# Plot
-par(mar=c(2,4,0,0)+.1)
-# plot(sample_table_all$mappingrates)
-# -> boring
-
-# increase margin for longer names
-par(mar=c(2,5,2,2)+.1)
-# barplot(height=mappingrates, names=samplename, horiz=T, las=1)
-
-xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename, horiz=T, las=1, xlim=c(0,100))
-text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.5, col = "red")
-
-par(mar=c(4,4,4,4)+.1)
-
-```
-
-## 2c. Tximeta
-
-```{r tximeta}
-library(sessioninfo)
-library(data.table)
-library(plyr)
-library(tidyverse)
-library(tximeta)
-library(tximport)
-library(curl)
-
-getwd()
-example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")
-
-list.files(quantdir)[1]
-file.exists(example.quant)
-
-# first file:
-example.table <- read.table(example.quant, header=T)
-head(example.table)
-basename(example.quant)
-dirname(example.quant)
-basename(dirname(example.quant))
-
-# generate file list & prepare variables
-files={}
-for (i in list.dirs(path = dirname(dirname(example.quant)), full.names = TRUE, recursive = FALSE)) {
-files <- c(files,paste(i,"/quant.sf",sep=""))
-print(basename(i))
-# print(head(read.table(files[length(files)], header=T)))
-print (files[length(files)])
-}
-head(files)
-
-dirfiles <- dirname(files)
-head(dirfiles)
-head(basename(dirfiles))
-run <- basename(dirfiles)
-dirdir <- dirname(dirfiles)
-head(dirdir)
-
-tximeta_files <- file.path(dirdir, run, "quant.sf") 
-file.exists(tximeta_files)
-file.exists(files)
-
-# names as working sample names
-# see 'm.table' at mapping rates
-coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
-coldata <- data.frame(coldata,sample.table)
-coldata %>% head()
-
-# load tximeta
-# with linked Transcriptome
-## Chlamy
-file.exists(indexDir, fastaPath, gtfPath)
-
-makeLinkedTxome(indexDir=indexDir,
-                source="Phytozome",
-                organism="Chlamydomonas reinhardtii",
-                release="v6.1",
-                genome="CreinhardtiiCC_4532_707",
-                fasta=fastaPath,
-                gtf=gtfPath,
-                write=FALSE)
-
-gtfdata <- readLines(gtfPath)
-head(gtfdata, n=20)
-
-# gene_name info sind in gff3 abder nicht in se?
-coldata_Cre <- filter(coldata, species == "Cre")
-se <- tximeta(coldata_Cre, useHub=T)
-se
-
-library(SummarizedExperiment)
-
-colData(se)
-# meta infos sind da (genotype, treatment,...)
-
-# rename Samples
-rownames(colData(se)) <- colData(se)$clientName
-
-rowRanges(se)
-# genome info
-seqinfo(se)
-# ?
-head(assays(se)[["counts"]])
-# counts. THE DATA
-
-# Mapping infos:
-names(metadata(se)[["quantInfo"]])
-str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
-metadata(se)[["quantInfo"]]$percent_mapped
-metadata(se)[["quantInfo"]]$num_processed
-
-par(mfrow=c(1,2))
-barplot(metadata(se)[["quantInfo"]]$percent_mapped, main="Mapping Rate")
-barplot(metadata(se)[["quantInfo"]]$num_processed/1000000, main="Mio. Reads")
-par(mfrow=c(1,1))
-
-edb <- retrieveDb(se)
-class(edb)
-genes(edb)
-columns(edb)
-
-se.exons <- addExons(se)
-rowRanges(se.exons)[[1]]
-gse <- summarizeToGene(se)
-rowRanges(gse)
-
-head(assays(gse)[["counts"]])[1:5,1:5]
-head(assays(gse)[["abundance"]])[1:5,1:5]
-head(assays(gse)[["length"]])[1:5,1:5]
-
-#### add gene symbol
-columns(retrieveDb(se))
-edb
-metadata(se)$txomeInfo$source
-
-TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="first"))
-# select()' returned 1:many mapping between keys and columns = 1 gene has many transcripts...
-TXNAME.list <- (mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="list"))
-head(TXNAME.list)
-# info is already there
-CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
-head(CDSID)
-# ?
-CDSNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSNAME", keytype = "GENEID", multiVals="first"))
-head(CDSNAME)
-# same as gene_id
-EXONNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "EXONNAME", keytype = "GENEID", multiVals="first"))
-head(EXONNAME)
-TXTYPE <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXTYPE", keytype = "GENEID", multiVals="first"))
-head(TXTYPE)
-# no useful info... no gene Symbol!!
-
-# add CDSID
-mcols(gse)$CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
-
-colnames(mcols(gse))
-head(rownames(gse))
-
-getwd()
-# save(gse,file=paste(dir,"tximeta.txm", sep="/"))
-# gse <- 1
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-
-```
-
-### -Add gene info from Master annotation
-
-Use other files from Phytozome database: CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv
-
-```{r anno}
-library(data.table)
-library(plyr)
-library(tidyverse)
-library(ape)
-
-# more gene info from master_annotation_table
-annopath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")
-
-anno <- read.delim2(annopath, header = T, sep = "\t",)
-head(anno)
-
-# search for genes
-anno[str_detect(anno[["locusName_4532"]],"Cre01.g000050"),]
-anno[str_detect(anno[["geneSymbol"]],"HKR"),]
-anno[str_detect(anno[["previousIdentifiers"]],"HKR"),]
-anno[str_detect(anno[["geneSymbol"]],"COP"),]
-anno[str_detect(anno[["previousIdentifiers"]],"COP1"),]
-anno[str_detect(anno[["geneSymbol"]],"cop"),]
-anno[str_detect(anno[["geneSymbol"]],"PLAP6"),]
-anno[str_detect(anno[["geneSymbol"]],"COQ3"),]
-anno[str_detect(anno[["geneSymbol"]],"PCRY"),]
-anno[str_detect(anno[["geneSymbol"]],"CRY"),]
-anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
-anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH"),]
-
-dim(anno)
-
-# Remove "_4532"
-anno$gene_id <- str_remove(anno$locusName_4532,pattern = "_4532")
-rownames(anno) <- anno$gene_id
-
-# convert previousIdentifiers to list
-anno$previousIdentifiers_list <- strsplit(anno$previousIdentifiers, "#")
-head(anno)
-anno$previousIdentifiers_list[1]
-
-# polish "previousIdentifiers"
-# remove Cre & g-numbers
-head(anno$previousIdentifiers)
-tail(anno$previousIdentifiers)
-prev.symbols <-  
-  str_replace_all(anno$previousIdentifiers,pattern="Cre\\d+.g\\d+.t\\d+.\\d+", replacement="?") %>% 
-  str_replace_all(pattern="Cre...g\\d+", replacement="?") %>% 
-  str_replace_all(pattern="g\\d+.t\\d+", replacement="?") %>% 
-  str_remove_all(pattern="\\?") %>%
-  str_remove_all("^#|#$| ") %>% str_remove_all("^#|#$") %>% str_remove_all("^#|#$|^0") %>% 
-  str_replace_all(pattern="##","#") %>% str_replace_all(pattern="##","#")
-prev.symbols[str_detect(prev.symbols,"#")] %>% head(n=20)
-prev.symbols[str_detect(prev.symbols,"0")] %>% head(n=20)
-prev.symbols[str_detect(prev.symbols,"Cre")] %>% head(n=20)
-tail(anno$previousIdentifiers[str_detect(anno$previousIdentifiers,"Cre")])
-
-anno$prev.symbols <- prev.symbols
-head(anno)
-
-# Make colum with gene symbol or prev.symbol or Gene id
-anno$combname <- anno$geneSymbol
-
-summary(anno$combname)
-summary(anno$combname=="")
-summary(anno$combname==" ")
-summary(is.na(anno$combname))
-summary(anno$combname=="0")
-
-# replace "" with prev.symbol
-summary(anno$combname=="")
-  length(anno$combname[anno$combname==""])
-  length(str_split(
-  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1])
-anno$combname[anno$combname==""] <- str_split(
-  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1]
-summary(anno$combname=="")
-
-anno$combname[anno$combname==""] <- anno$gene_id[anno$combname==""]
-summary(anno$combname=="")
-subset(anno,anno$combname=="")
-
-summary(duplicated(anno$combname))
-# some duplicates
-
-colnames(anno)[which(names(anno) == "combname")] <- "id.symbol"
-
-summary(anno$geneSymbol!="")
-# 4408 gene Symbols
-summary(anno$previousIdentifiers!="")
-# all (!) 17712 genes have "previousIdentifiers"
-summary(anno$prev.symbols!="")
-# 7236 have alternative/old symbols
-summary(anno$id.symbol)
-
-dim(anno)
-
-anno %>% head()
-
-# saveRDS(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
-# anno <- readRDS(paste(outdir,"anno.RDS",sep="/"))
-
-```
-
-### -add anno data into gse ####
-```{r anno_gse}
-
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-gse
-anno <- readRDS(file=paste(dir,"/anno.RDS", sep="/"))
-anno %>% head()
-
-# Remove "_4532"
-mcols(gse)$gene_id2 <- mcols(gse)[,"gene_id"]
-mcols(gse)$gene_id <- str_remove(mcols(gse)[,"gene_id2"], pattern = "_4532")
-rownames(gse) <- mcols(gse)$gene_id
-
-summary(anno$gene_id %in% mcols(gse)$gene_id)
-summary(mcols(gse)$gene_id %in% anno$gene_id)
-
-# choose info
-gsetable <- as.data.frame(mcols(gse)[,c("gene_id","gene_id2")])
-
-anno_join <- anno[,c("gene_id", "geneSymbol", "id.symbol","prev.symbols", "previousIdentifiers", "previousIdentifiers_list", "Description", "Comments", "TargetP", "Predalgo")]
-anno_join
-
-class(anno_join)
-dim(anno_join)
-
-# any duplicated:
-summary(duplicated(anno$gene_id))
-
-new <- plyr::join(gsetable,anno_join, by = "gene_id", type = "left")
-dim(new)
-head(new)
-new2 <- left_join(gsetable,anno_join,by = "gene_id")
-dim(new2)
-head(new2)
-
-mcols(gse) <- left_join(as.data.frame(mcols(gse)),anno_join,by = "gene_id")
-mcols(gse)
-
-getwd()
-
-# save(gse,file=paste(dir,"tximeta.txm",sep="/"))
-```
-
-## 2c. DESeq2 Analysis
-
-```{r deseq2}
-getwd()
-# load(file=paste(outdir,"tximeta.txm", sep="/"))
-
-library(DESeq2)
-colData(gse)
-mcols(gse)
-gse$condition
-
-design <- ~condition
-# design <- ~genotype+treatment+genotype:treatment
-
-dds <- DESeqDataSet(gse, design=design)
-colData(dds) %>% head()
-
-##########################################\n
-## filter all rows with rowsum = 0 ####
-par(mfrow=c(1,2))
-hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
-hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
-# -> dip at log(counts(dds)=2-3)
-par(mfrow=c(1,1))
-
-# min counts
-# at least 5 counts in 3 samples
-keep.sn <- rowSums(counts(dds) >= 5) >= 3
-keep.sn %>% summary()
-dds <- dds[keep.sn,]
-
-dds <- estimateSizeFactors(dds)
-dds
-head(assays(dds)[["counts"]])[1:5,1:5]
-length(rownames(counts(dds)))
-
-colData(dds)
-# add metadata
-# colData(dds)$strain <- factor(colData(dds)$strain)
-
-# ## combine technical replicates #####
-# colData(dds)
-# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
-# colData(dds)
-# dds$runsCollapsed
-# colnames(dds)
-# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
-# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))
-
-###########
-# run DESeq
-dds <- DESeq(dds)
-DESeq2::plotMA(dds)
-
-plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition")
-
-head(mcols(dds)$geneSymbol)
-
-anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
-# no hit
-g <- anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH1"),"gene_id"]
-g
-plotCounts(dds, gene = g, intgroup = "condition")
-g <- anno[str_detect(anno[["geneSymbol"]],"PCRY"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ACRY"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ROC15"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ROC40"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-
-# search genes
-anno[str_detect(anno[["geneSymbol"]],"CRY"),]
-anno[str_detect(anno[["geneSymbol"]],"ROC"),]
-anno[str_detect(anno[["previousIdentifiers"]],"CRY"),]
-
-###############
-# save dds #####
-
-# save(dds, file = paste(dir,"dds_23design.RDS",sep="/"), compress = FALSE)
-# save(dds, file = paste(dir,"dds.RDS",sep="/"), compress = FALSE)
-# dds <- readRDS(paste(outdir,"dds.RDS",sep="/"))
-# dds
-
-```

### 2024-04-17 1_data_processing/1_data_processing.Rmd
counts figures upload

@@
 par(mar=c(2,4,0,0)+.1) # set margins for figure
-plot(sample_table_all$mappingrates)
+plot(m.table$mappingrates)
 # -> boring
@@
 ### Deseq2 acry
-```{r}
+```{r dds_acry}
 design <- ~condition
@@
 ##########################################\n
-## filter all rows with rowsum = 0 ####
+# filter all rows with rowsum = 0 ####
 par(mfrow=c(1,2))

### 2024-04-20 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
acry pcry combine

--- 2024-04-17 1_data_processing/1_data_processing.Rmd
@@
 ---
-title: "RNAseq aCRY processing"
+title: "RNAseq aCRY pCRY data dive"
 author: "Kelterborn"
-date: "2024-03-15"
+date: "2024-04-19"
 output:
@@
 knit: (function(input_file, encoding) {
-    rmarkdown::render(input_file,output_file= 'README.md')
+    rmarkdown::render(input_file,output_file= 'Readme.md')
     })
@@
 
-# 0. Prepare System
-BiocManager::install()
+BiocManager::install("ggalt") # , type="source"
 
+# Load System
 ## -R_libraries
 ```{r r_libraries}
+library(PCAtools)
 library(stringr)
@@
 library(patchwork)
-library("ggpubr")
+library(ggpubr)
 library(vsn)
@@
 library(RColorBrewer)
-library(PCAtools)
 
@@
 ```{r r_folders}
+
+# Linux Workstation
 ifelse(Sys.info()["sysname"]== "Linux",
        s <- "/mnt/s",
-       s <- "S:")
+    ifelse(Sys.info()["sysname"]== "Darwin",
+       s <- "/mnt/s",
+       s <- "S:"))
 dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
-list.files(dir) %>% head()
+# list.files(dir) %>% head()
 gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
-list.files(gitdir) %>% head()
+# list.files(gitdir) %>% head()
 
-fastqdir <- paste(dir,"fastq",sep="/")
-list.files(fastqdir) %>% tail()
-fastqdir2 <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/",sep="/")
-list.files(fastqdir2) %>% head()
-
-indexDir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")
-
-fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
-head(readLines(fastaPath,n=10))
-
-gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
-readLines(gtfPath,n=10)
-
-quantdir <- paste(dir,"quants",sep="/")
+# Macbook
+scriptdir <- rstudioapi::getSourceEditorContext()$path
+gitdir <- scriptdir %>% dirname() %>% dirname()
+dir <- gitdir %>% dirname()
 
@@
 
-## -Linux folders
-```{bash linux_folders, eval=FALSE}
-fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
-fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;
+# 1. Load aCRY & pCRY
+## - Tximeta files
+```{r get_tximeta_files}
 
-dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY";
+# Annotation file
+load(file=paste(dir,"anno.RDS", sep="/"))
 
-quantdir="$dir/quants";
+# Seq 1
+tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm"
+load(file=tximeta_pcry_file)
+gse_pcry <- gse
 
-indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";
-
-```
-
-# 1. Raw data processing
-## 1a. Download data (AWS CLI)
-
-The Raw data for your project has also been uploaded completely. Please find the data download guide attached.
-Index in AWS: bmkdatarelease-37/delivery_20240313120107785
-Access key ID: AKIA45O7KHZWJCKW62VE
-Secret Access Key: 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
-Web log-in link: https://sctrack.sendcloud.net/track/click2/eNpFj8FOxDAMRP8lgls2jR0aNzf2O1YoSlOzVNs6qOkKBOLfyS4HJGsO1jzP-BT60A9aKa2AwDobPBI5H58GCD4ChIAQsamp-QA2hggY0frDLGO5ymQfljnJ-VOex7msabvwZnJZTRalrVZv-_5eH92x64aBArQwR9RuzWeZxaSPatKavorcmC4XqWXhO_ialsq3VmgNODR9b4JX__s2TI5Hytxqkwcm5JYw4EgJJ57Q_blq3reUL6ayTHkp18kI70p__-j7i9qFl18qJUo2.html
-Username:delivery_20240313120107785
-Password:Pwd2024@x4S416
-
-```{bash download data, eval=FALSE}
-
-# install aws
-curl "https://awscli.amazonaws.com/awscli-exe-linux-x86_64.zip" -o "awscliv2.zip"
-unzip awscliv2.zip
-sudo ./aws/install
-
-aws --version
-
-aws configure
-AWS Access Key ID [None]: AKIA45O7KHZWJCKW62VE
-AWS Secret Access Key [None]: 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
-Default region name [None]: eu-north-1
-Default output format [None]: table
-
-fastq="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq"
-cd $fastq
-
-# test
-aws s3api get-object --bucket bmkdatarelease-37 --key delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/sampleName_clientId.txt sampleName_clientId.txt
-
-# all files
-aws s3 cp s3://bmkdatarelease-37/delivery_20240313120107785/BMK231214-BU327-ZX01-0201/BMK_DATA_20240313120107_1/Data/rawdata/ . --recursive
-
-```
-
-### check md5
-```{bash md5, eval=FALSE}
-cd $fastqdir
-ls
-# 1 file
-md5sum Unknown_BU327-002T0008_1.fq.gz
-
-# Check downloads
-
-# Path to the text file containing the MD5 sums
-md5_file="data_md5.txt"
-
-# Function to check the MD5 sum of a file
-check_md5() {
-    file_name=$1
-    expected_sum=$2
-
-    md5_sum=$(md5sum "$file_name" | awk '{print $1}')
-    if [[ $md5_sum == $expected_sum ]]; then
-        echo "$file_name: MD5 sum is correct."
-    else
-        echo "$file_name: MD5 sum is incorrect."
-    fi
-}
-
-# Read the MD5 sums from the text file and check them
-while IFS= read -r line; do
-    file_name=$(echo "$line" | awk '{print $2}')
-    expected_sum=$(echo "$line" | awk '{print $1}')
-    if [[ -f $file_name ]]; then
-        check_md5 "$file_name" "$expected_sum"
-    else
-        echo "$file_name: File not found."
-    fi
-done < "$md5_file" | tee -a $log
-
-```
-
-## 1b. Fastqc
-
-## 1c. Mapping
-```{bash mapping, eval=FALSE}
-mamba activate salmon;
-
-fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
-fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);
-
-# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
-fastqfiles_all=$fastqfiles;
-
-echo "${fastqfiles:0:100}";
-echo "${fastqfiles2:0:100}";
-echo "${fastqfiles_all:0:100}";
-
-# files: Unknown_BU327-002T0001_1.fq.gz;
-for fn in $fastqfiles_all;
-do
-bfn=$(echo "`basename ${fn}`");
-samp=${bfn%"_1.fq.gz"};
-R1=$fn;
-R2=$(echo "$R1" | sed 's/_1/_2/');
-echo "Processing Sample: $samp";
-test -f $R1 && echo "--> File: $R1 exists"
-test -f $R2 && echo "--> File: $R2 exists"
-outquant=${quantdir}/${samp}_quant
-echo $outquant
-echo $indexdir
-
-salmon quant -i $indexdir -l A \
-  -1 $R1 \
-  -2 $R2 \
-  -p 25 --validateMappings --gcBias -o $outquant
-
-done
-
-```
-
-# 2. Process Data
-## 2a. Sample names
-```{r sample_table}
-# Get Sample table
-# Filenames
-
-quantdir <- paste(dir,"quants",sep="/")
-
-f <- list.files(path = quantdir)
-files <- as.factor(str_remove(f,pattern ="_quant"))
-# short form
-sample <- str_remove(files,pattern ="Unknown_")
-sample
-# condition
-library(readxl)
-Samplepath <- filePath(path=dir, file="sample_table.xlsx")
-Samplefile <- read_xlsx(Samplepath)
-samplename <- Samplefile$clientName
-Samplefile$treatment
-Samplefile$condition <- paste(Samplefile$genotype, Samplefile$treatment, sep="_") %>% factor(levels = c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red","WT_3HL","WT_16HL","cia5_16HL"))
-Samplefile$treatment <- Samplefile$treatment %>% factor(levels = c("dark","blue","red","light","3HL","16HL"))
-# Make table
-sample.table <- data.frame(Samplefile[,-3])
-sample.table %>% kable()
-
-sample.table <- data.frame(lapply(sample.table, factor))
-
-# Check the updated structure of the data frame
-str(sample.table)
-
-```
-
-## 2b. Mapping Rates
-```{r mapping_rates}
-
-samplelist <- {}
-mappingrates <- {}
-for (i in list.files(path = quantdir)){
-print(i)
-si <- paste("",str_sub(i,11,-7),sep = "")
-si
-samplelist <- c(samplelist,si)
-f <- readLines(paste(quantdir,i,"logs/salmon_quant.log", sep="/"))
-line <- grep("Mapping rate = ",f,value=TRUE)
-sl <- str_length(line)
-sl
-notime <- substring(line,30,sl)
-notime
-manual <- substring(line,sl-7,sl-1)
-val <- as.numeric(str_extract(notime,"[0-9.]+"))
-val
-valr<-round(val, digits=2)
-print(paste("Mapping rate of ",si," is: ",valr," %"))
-mappingrates <- c(mappingrates,valr)
-}
-sample.table$mappingrates <- mappingrates
-# Make table
-
-m.table <- data.frame(sample.table,mappingrates)
-```
-
-### Plot mapping rates
-```{r plot_mappingrates}
-
-# Plot
-par(mar=c(2,4,0,0)+.1) # set margins for figure
-plot(m.table$mappingrates)
-# -> boring
-
-# increase margin for longer names
-par(mar=c(2,5,2,2)+.1)
-# barplot(height=mappingrates, names=samplename, horiz=T, las=1)
-
-xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename, horiz=T, las=1, xlim=c(0,100))
-text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.5, col = "red")
-
-par(mar=c(4,4,4,4)+.1)
-
-```
-
-## 2c. Tximeta
-
-```{r tximeta}
-library(sessioninfo)
-library(data.table)
-library(plyr)
-library(tidyverse)
-library(tximeta)
-library(tximport)
-library(curl)
-
-getwd()
-example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")
-
-list.files(quantdir)[1]
-file.exists(example.quant)
-
-# first file:
-example.table <- read.table(example.quant, header=T)
-head(example.table)
-basename(example.quant)
-dirname(example.quant)
-basename(dirname(example.quant))
-
-# generate file list & prepare variables
-files={}
-for (i in list.dirs(path = dirname(dirname(example.quant)), full.names = TRUE, recursive = FALSE)) {
-files <- c(files,paste(i,"/quant.sf",sep=""))
-# print(basename(i))
-# print(head(read.table(files[length(files)], header=T)))
-# print (files[length(files)])
-}
-head(files)
-
-dirfiles <- dirname(files)
-head(dirfiles)
-head(basename(dirfiles))
-run <- basename(dirfiles)
-dirdir <- dirname(dirfiles)
-head(dirdir)
-
-tximeta_files <- file.path(dirdir, run, "quant.sf") 
-file.exists(tximeta_files)
-file.exists(files)
-
-# names as working sample names
-# see 'm.table' at mapping rates
-coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
-coldata <- data.frame(coldata,sample.table)
-coldata %>% head() %>% kable()
-
-# load tximeta
-# with linked Transcriptome
-## Chlamy
-file.exists(indexDir, fastaPath, gtfPath)
-
-makeLinkedTxome(indexDir=indexDir,
-                source="Phytozome",
-                organism="Chlamydomonas reinhardtii",
-                release="v6.1",
-                genome="CreinhardtiiCC_4532_707",
-                fasta=fastaPath,
-                gtf=gtfPath,
-                write=FALSE)
-
-gtfdata <- readLines(gtfPath)
-head(gtfdata, n=10)
-
-# gene_name info sind in gff3 abder nicht in se?
-coldata_Cre <- filter(coldata, species == "Cre")
-se <- tximeta(coldata_Cre, useHub=T)
-se
-
-library(SummarizedExperiment)
-
-colData(se) %>% head() %>% kable()
-# meta infos sind da (genotype, treatment,...)
-
-# rename Samples
-rownames(colData(se)) <- colData(se)$clientName
-
-rowRanges(se) %>% kable()
-# genome info
-seqinfo(se)
-# ?
-head(assays(se)[["counts"]]) %>% kable()
-# counts. THE DATA
-
-# Mapping infos:
-names(metadata(se)[["quantInfo"]])
-# str(metadata(se)[["quantInfo"]]) # Infos from Salmon Mapping
-metadata(se)[["quantInfo"]]$percent_mapped
-metadata(se)[["quantInfo"]]$num_processed
-
-par(mfrow=c(1,2))
-barplot(metadata(se)[["quantInfo"]]$percent_mapped, main="Mapping Rate")
-barplot(metadata(se)[["quantInfo"]]$num_processed/1000000, main="Mio. Reads")
-par(mfrow=c(1,1))
-
-edb <- retrieveDb(se)
-class(edb)
-genes(edb)
-columns(edb)
-
-se.exons <- addExons(se)
-rowRanges(se.exons)[[1]]
-gse <- summarizeToGene(se)
-rowRanges(gse)
-
-head(assays(gse)[["counts"]])[1:5,1:5]
-head(assays(gse)[["abundance"]])[1:5,1:5]
-head(assays(gse)[["length"]])[1:5,1:5]
-
-#### add gene symbol
-columns(retrieveDb(se))
-edb
-metadata(se)$txomeInfo$source
-
-TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="first"))
-# select()' returned 1:many mapping between keys and columns = 1 gene has many transcripts...
-TXNAME.list <- (mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="list"))
-head(TXNAME.list)
-# info is already there
-CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
-head(CDSID)
-# ?
-CDSNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSNAME", keytype = "GENEID", multiVals="first"))
-head(CDSNAME)
-# same as gene_id
-EXONNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "EXONNAME", keytype = "GENEID", multiVals="first"))
-head(EXONNAME)
-TXTYPE <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXTYPE", keytype = "GENEID", multiVals="first"))
-head(TXTYPE)
-# no useful info... no gene Symbol!!
-
-# add CDSID
-mcols(gse)$CDSID <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "CDSID", keytype = "GENEID", multiVals="first"))
-
-colnames(mcols(gse))
-head(rownames(gse))
-
-getwd()
-# save(gse,file=paste(dir,"tximeta.RDS", sep="/"))
-# gse <- 1
-# load(file=paste(dir,"tximeta.RDS", sep="/"))
-
-```
-
-### -Add gene info from Master annotation
-
-Use other files from Phytozome database: CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv
-
-```{r anno}
-library(data.table)
-library(plyr)
-library(tidyverse)
-library(ape)
-
-# more gene info from master_annotation_table
-annopath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")
-
-anno <- read.delim2(annopath, header = T, sep = "\t",)
-head(anno)[1:5,1:9] %>% kable()
-
-# search for genes
-anno[str_detect(anno[["locusName_4532"]],"Cre01.g000050"),]
-anno[str_detect(anno[["geneSymbol"]],"HKR"),]
-anno[str_detect(anno[["previousIdentifiers"]],"HKR"),]
-anno[str_detect(anno[["geneSymbol"]],"COP"),]
-anno[str_detect(anno[["previousIdentifiers"]],"COP1"),]
-anno[str_detect(anno[["geneSymbol"]],"cop"),]
-anno[str_detect(anno[["geneSymbol"]],"PLAP6"),]
-anno[str_detect(anno[["geneSymbol"]],"COQ3"),]
-anno[str_detect(anno[["geneSymbol"]],"PCRY"),]
-anno[str_detect(anno[["geneSymbol"]],"CRY"),]
-anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
-anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH"),]
-
-dim(anno)
-
-# Remove "_4532"
-anno$gene_id <- str_remove(anno$locusName_4532,pattern = "_4532")
-rownames(anno) <- anno$gene_id
-
-# convert previousIdentifiers to list
-anno$previousIdentifiers_list <- strsplit(anno$previousIdentifiers, "#")
-head(anno)
-anno$previousIdentifiers_list[1]
-
-# polish "previousIdentifiers"
-# remove Cre & g-numbers
-head(anno$previousIdentifiers)
-tail(anno$previousIdentifiers)
-prev.symbols <-  
-  str_replace_all(anno$previousIdentifiers,pattern="Cre\\d+.g\\d+.t\\d+.\\d+", replacement="?") %>% 
-  str_replace_all(pattern="Cre...g\\d+", replacement="?") %>% 
-  str_replace_all(pattern="g\\d+.t\\d+", replacement="?") %>% 
-  str_remove_all(pattern="\\?") %>%
-  str_remove_all("^#|#$| ") %>% str_remove_all("^#|#$") %>% str_remove_all("^#|#$|^0") %>% 
-  str_replace_all(pattern="##","#") %>% str_replace_all(pattern="##","#")
-prev.symbols[str_detect(prev.symbols,"#")] %>% head(n=20)
-prev.symbols[str_detect(prev.symbols,"0")] %>% head(n=20)
-prev.symbols[str_detect(prev.symbols,"Cre")] %>% head(n=20)
-tail(anno$previousIdentifiers[str_detect(anno$previousIdentifiers,"Cre")])
-
-anno$prev.symbols <- prev.symbols
-head(anno)[1:5,1:9] %>% kable()
-
-# Make colum with gene symbol or prev.symbol or Gene id
-anno$combname <- anno$geneSymbol
-
-summary(anno$combname)
-summary(anno$combname=="")
-summary(anno$combname==" ")
-summary(is.na(anno$combname))
-summary(anno$combname=="0")
-
-# replace "" with prev.symbol
-summary(anno$combname=="")
-  length(anno$combname[anno$combname==""])
-  length(str_split(
-  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1])
-anno$combname[anno$combname==""] <- str_split(
-  anno$prev.symbols[anno$combname==""],pattern ="#", simplify = T)[,1]
-summary(anno$combname=="")
-
-anno$combname[anno$combname==""] <- anno$gene_id[anno$combname==""]
-summary(anno$combname=="")
-subset(anno,anno$combname=="")
-
-summary(duplicated(anno$combname))
-# some duplicates
-
-colnames(anno)[which(names(anno) == "combname")] <- "id.symbol"
-
-summary(anno$geneSymbol!="")
-# 4408 gene Symbols
-summary(anno$previousIdentifiers!="")
-# all (!) 17712 genes have "previousIdentifiers"
-summary(anno$prev.symbols!="")
-# 7236 have alternative/old symbols
-summary(anno$id.symbol)
-
-dim(anno)
-
-# anno[1:10,1:10]
-
-# save(anno, file = paste(dir,"anno.RDS",sep="/"), compress = FALSE)
-# load(paste(dir,"anno.RDS",sep="/"))
-
-```
-
-### -add anno data into gse ####
-```{r anno_gse}
-
-# load(file=paste(dir,"tximeta.RDS", sep="/"))
-# load(file=paste(dir,"/anno.RDS", sep="/"))
-anno[1:3,1:8] %>% kable()
-
-# Remove "_4532"
-mcols(gse)$gene_id2 <- mcols(gse)[,"gene_id"]
-mcols(gse)$gene_id <- str_remove(mcols(gse)[,"gene_id2"], pattern = "_4532")
-rownames(gse) <- mcols(gse)$gene_id
-
-summary(anno$gene_id %in% mcols(gse)$gene_id)
-summary(mcols(gse)$gene_id %in% anno$gene_id)
-
-# choose info
-gsetable <- as.data.frame(mcols(gse)[,c("gene_id","gene_id2")])
-
-anno_join <- anno[,c("gene_id", "geneSymbol", "id.symbol","prev.symbols", "previousIdentifiers", "previousIdentifiers_list", "Description", "Comments", "TargetP", "Predalgo")]
-anno_join %>% head() %>% kable()
-
-class(anno_join)
-dim(anno_join)
-
-# any duplicated:
-summary(duplicated(anno$gene_id))
-
-new <- plyr::join(gsetable,anno_join, by = "gene_id", type = "left")
-dim(new)
-head(new) %>% kable()
-
-new2 <- left_join(gsetable,anno_join,by = "gene_id")
-dim(new2)
-head(new2) %>% kable()
-
-mcols(gse) <- left_join(as.data.frame(mcols(gse)),anno_join,by = "gene_id")
-mcols(gse) %>% head() %>% kable()
-
-save(gse,file=paste(dir,"tximeta.RDS",sep="/"))
-```
-
-## 2c. DESeq2 Analysis
-
-```{r deseq2}
-getwd()
-# load(file=paste(dir,"tximeta.RDS", sep="/"))
-
-colData(gse) %>% head() %>% kable()
-mcols(gse) %>% head() %>% kable()
-
+# Seq 2 (acry + cia5)
+load(file=paste(dir,"tximeta.RDS", sep="/"))
 gse_acry <- gse[,colData(gse)$experiment=="acry"]
@@
 
-gse_cia5 <- gse[,colData(gse)$experiment=="cia5"]
-colData(gse_cia5) <- colData(gse_cia5) %>% droplevels()
-colData(gse_cia5)$genotype <- colData(gse_cia5)$genotype %>% relevel(ref="WT")
+# combine
+( mcols(gse_acry) == mcols(gse_pcry) ) %>% summary()
+# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
+names(colData(gse_pcry)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
+colData(gse_acry) <- colData(gse_acry)[,c(1:4,7:12)]
+colData(gse_pcry)$experiment <- "pcry"
+colData(gse_pcry) <- colData(gse_pcry)[,c(1,3,2,4:6,8,12,7,11)]  
+# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
+
+# now same metadata
+gse <- cbind(gse_acry,gse_pcry)
+
+colData(gse)$genotype
+
+colData(gse)$treatment <- colData(gse)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
+  str_replace(pattern="BL", replacement = "blue") %>%
+  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))
+
+colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","pcry_dark","WT_blue","acry_blue","pcry_blue","WT_red","acry_red","pcry_red"))
+
+colData(gse)$experiment
 
@@
 
-### Deseq2 acry
-```{r dds_acry}
-design <- ~condition
-# design <- ~genotype+treatment+genotype:treatment
+## Run Deseq2
+```{r dds_acry_pcry}
+# design <- ~condition
+design <- ~genotype+treatment+genotype:treatment
 
-dds <- DESeqDataSet(gse_acry, design=design)
+dds <- DESeqDataSet(gse, design=design)
 colData(dds) %>% head() %>% kable()
@@
 ##########################################\n
-# filter all rows with rowsum = 0 ####
+# filter all rows with rowsum = 0 ####\n
 par(mfrow=c(1,2))
@@
 
-colData(dds) %>% head() %>% kable()
+# colData(dds) %>% head() %>% kable()
 # add metadata
@@
 
-# ## combine technical replicates #####
+# ## combine technical replicates #####\n
 # colData(dds)
@@
 
-###########
+###########\n
 # run DESeq
 dds <- DESeq(dds)
+
+plotDispEsts(dds)
+resultsNames(dds)
 DESeq2::plotMA(dds)
 
-plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition")
+plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$experiment)
 
@@
 
-anno[str_detect(anno[["geneSymbol"]],"CRY-DASH"),]
-# no hit
-g <- anno[str_detect(anno[["previousIdentifiers"]],"CRY-DASH1"),"gene_id"]
-g
-plotCounts(dds, gene = g, intgroup = "condition")
-g <- anno[str_detect(anno[["geneSymbol"]],"PCRY"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ACRY"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ROC15"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-g <- anno[str_detect(anno[["geneSymbol"]],"ROC40"),"gene_id"]
-plotCounts(dds, gene = g, intgroup = "condition", col=dds$genotype, main =anno[g,"geneSymbol"])
-
-# search genes
-anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()
-
-anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
-anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()
-
-###############
+###############\n
 # save dds #####
 
-# save(dds, file = paste(dir,"dds_23design.RDS",sep="/"), compress = FALSE)
+# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
 # save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
@@
 
-### Deseq2 cia5
-```{r}
-design <- ~genotype
-dds <- DESeqDataSet(gse_cia5, design=design)
-colData(dds) %>% head() %>% kable()
-
-## filter all rows with rowsum = 0 ####
-par(mfrow=c(1,2))
-hist(log(counts(dds)), breaks=100, ylim = c(0,4000), xlim = c(0,10))
-hist(counts(dds), breaks=300000, ylim = c(0,10000), xlim = c(0,10))
-# -> dip at log(counts(dds)=2-3)
-par(mfrow=c(1,1))
-
-# min counts
-# at least 5 counts in 3 samples
-keep.sn <- rowSums(counts(dds) >= 5) >= 3
-keep.sn %>% summary()
-dds <- dds[keep.sn,]
-
-dds <- estimateSizeFactors(dds)
-dds
-head(assays(dds)[["counts"]])[1:5,1:5]
-length(rownames(counts(dds)))
-
-colData(dds) %>% head() %>% kable()
-
-# run DESeq
-dds <- DESeq(dds)
-
-# qc check
-DESeq2::plotMA(dds)
-plotDispEsts(dds)
-
-# counts
-plotCounts(dds, gene = "Cre01.g000150", intgroup = "genotype", col=colData(dds)$genotype)
-
-save(dds, file = paste(dir,"dds_cia5.RDS",sep="/"), compress = FALSE)
-
-```
-
-# 3. Pre-Analysis
-### load data
-```{r}
-load(file=paste(dir,"dds_cia5.RDS", sep="/"))
-load(file=paste(dir,"dds_acry.RDS", sep="/"))
-load(file=paste(dir,"anno.RDS", sep="/"))
-
-```
-
+# 2. Pre-Analysis
 ### - Data transformations
@@
 
-```{r pre_trans_fig, figures-side, fig.show="hold", out.width="33%"}
+```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
 # load(file=paste(data,"rlog.rld", sep="/"))
+meanSdPlot(counts(dds, normalized =TRUE))
 meanSdPlot(assay(ntd))
@@
 colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
-anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype")])
-anno_colors <- list(treatment = c("grey20","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
-                    genotype = c("grey","seagreen3","tan2"))
+anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
+anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
+                    genotype = c("green4","orchid","tan2"))
 
@@
 
-g2 <- ggplot(pcaData, aes(PC1, PC2, color=genotype, shape=genotype)) +
+g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
   geom_point(size=5, alpha=0.7) +
-  labs(title = "genotype") +
+  labs(title = "experiment") +
   xlab(paste0("PC1: ",percentVar[1],"% variance")) +
@@
 # find explaining PCs
-# horn <- parallelPCA(vst_dat)
+horn <- parallelPCA(vst_dat)
 # horn$n = 2
@@
 
-horn <- list()
-horn$n <- 2
+# horn <- list()
+# horn$n <- 2
 
@@
     geom_label(aes(x = horn$n + 1, y = 30,
-      label = 'Horn\'s=2', vjust = -1, size = 8)) +
+      label = paste('Horn',horn$n,sep="="), vjust = -1, size = 8)) +
     geom_label(aes(x = elbow + 1, y = 50,
-      label = 'Elbow=4', vjust = -1, size = 8))
+      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8))
 
@@
 
-### - Plot example counts
-#### cia5
+### 3. Plot counts
+####
 
@@
 
-load(file=paste(dir,"dds_cia5.RDS", sep="/"))
+# search genes
+anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()
 
-## cia5
+anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
+anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()
 

### 2024-04-21 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
Rstudio fix
move folder

@@
 knit: (function(input_file, encoding) {
+    })
     rmarkdown::render(input_file,output_file= 'Readme.md')
-    })
 ---
@@
 gitdir <- scriptdir %>% dirname() %>% dirname()
-dir <- gitdir %>% dirname()
+dir <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq aCRY"
 
@@
 
-### 3. Plot counts
-####
+# 3. Results
+## - Plot Counts
 

### 2024-04-21 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
include pCRY figures

@@
 # 3. Results
-## - Plot Counts
 
-```{r example_counts, fig.show="hold", out.width="50%", eval=FALSE, include=TRUE, echo=FALSE}
-
-# search genes
-anno[str_detect(anno[["geneSymbol"]],"CRY"),1:9] %>% kable()
-
-anno[str_detect(anno[["geneSymbol"]],"ROC"),1:9] %>% kable()
-anno[str_detect(anno[["previousIdentifiers"]],"CRY"),1:9] %>% kable()
-
-# plap6
-PLAP6 <- "Cre03.g188700"
-d <- plotCounts(dds,gene = PLAP6, intgroup= c("treatment","genotype","condition"), returnData=TRUE)
-
-resultsNames(dds)
-res <- results(dds,contrast = c("genotype","cia5","WT"))
-res["Cre03.g188700",]
-ggplot(d, aes(x = genotype, y = count, fill=genotype, color=genotype)) +
-    geom_boxplot(color="black") +
-    geom_point(shape=21,color="black",aes(fill=genotype),position=position_dodge(width=0.75), alpha=1) +
-   # scale_fill_manual(values=cols[c(1,5)]) +
-  #  scale_color_manual(values=cols[c(1,5)]) +
-   scale_y_continuous(trans = "log2") +
-  scale_fill_manual(values = c("grey30","salmon")) + 
-  scale_color_manual(values = c("grey30","salmon")) + 
-    labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
-  assign(paste("gcounts_",ig,sep=""),gcounts)
-
-  gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
-  geom_boxplot(fatten = 1) +
-  scale_fill_manual(values = group.colors) +
-  scale_y_continuous(trans = "log2")
-gcounts
-# ggsave(paste(outdir,"graphs/2023_11_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
-#   width = 18,
-#   height = 10)
-
-goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "DCRY2", "PHOT1", "UVR8", "HKR1") #"HKR2"
-
-anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
-rownames(anno_phot) <- anno_phot$geneSymbol
-anno_phot <- anno_phot[goi_phot,]
-
-rownames(anno_phot) <- anno_phot$geneSymbol
-anno_phot <- anno_phot[goi_phot,]
-
-# Get Count data
-i <- 1
-l <- length(goi_phot)
-all_counts <- {}
-for (i in 1:l){
-  d <-  plotCounts(dds, gene=anno_phot[i,"gene_id"], intgroup=c("condition","genotype","treatment"), main=anno_phot[i,"id.symbol"],returnData=TRUE)
-  d$Gene <- rep(anno_phot[i,"id.symbol"],length(rownames(d)))
-  d$sample <- rownames(d)
-  # rownames(d) <- {}
-  all_counts <- bind_rows(all_counts,d)
-  }
+## Plot Counts
+```{r}
 
@@
 
-goi_tf <- c("CCM1","LCR1", "HY5", "PHOT1", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
+goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
 anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
@@
 for (i in 1:l){
-  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","treatment","genotype"),col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
   d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
@@
 levels(all_counts$condition)
-all_counts$Gene <- factor(all_counts$Gene, levels = anno_tf$geneSymbol[c(9,6,8,7,15,2,1,4,3,5,13,11,14,10,12)])
+all_counts$Gene <- factor(all_counts$Gene)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
 levels(all_counts$Gene)
@@
 # colours
-# group.colors <- rep(anno_colors$treatment, each = 2)
-group.colors <- rep(anno_colors$treatment, each = 2)[c(1:6,7,9,11)]
-names(group.colors) <- colData(dds)$condition %>% levels()
-anno_colors$condition <- group.colors
-
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
 max_val <- 1.0*max(all_counts$count)
@@
 gcounts
+ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
+  width = 12,
+  height = 8)
 
-cols = brewer.pal(n=9,name = 'Paired')
+## Phot genes all-in-one
+goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",
 
-gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition, color=condition)) +
-  geom_boxplot(color="black") +
-  geom_point(position=position_dodge(width=0.75), alpha=1) +
-  scale_fill_manual(values=cols) +
-  scale_color_manual(values=cols) +
+anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
+rownames(anno_phot) <- anno_phot$geneSymbol
+anno_phot <- anno_phot[goi_phot,]
+
+# Combined counts table TFs
+
+goi <- anno_phot
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
+levels(all_counts$Gene)
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+max_val <- 1.0*max(all_counts$count)
+
+# Plot
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
   scale_y_continuous(trans = "log2")
 gcounts
+ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
+  width = 12,
+  height = 8)
 
-gcounts2 <- ggplot(all_counts, aes(x = condition, y = count, fill=Gene)) +
+```
+
+# ---------------------------
+# COPY & PASTE from pCRY
+## Make results
+```{r, eval=FALSE, include=FALSE, echo=FALSE}
+outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
+setwd(outdir)
+
+dds <- readRDS(paste(outdir,"dds_23design.RDS",sep="/"))
+anno <- readRDS(file=paste(outdir,"anno.RDS", sep="/"))
+
+design(dds)
+resultsNames(dds)
+cbind(resultsNames(dds),c(rep(0,6)))
+
+# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))
+
+# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
+res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
+res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")
+
+res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))
+
+res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))
+
+# Unterschiede zwischen versch. Belichtungen über beide Genotypen
+# D vs. (BL+R)
+res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))
+
+# D vs. BL
+res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))
+
+# D vs. R
+res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))
+
+# WT vs. pCRY
+res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))
+
+## Make top lists:
+
+topgenes <- function(res){
+a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
+a$id.symbol <- anno[rownames(a),"id.symbol"]
+a$symbol <- anno[rownames(a),"geneSymbol"]
+a <- a[order(a$baseMean, decreasing = T),]
+  a$rank.bm <- seq(1:length(rownames(a)))
+a <- a[order(a$padj, decreasing = F),]
+  a$rank.padj <- seq(1:length(rownames(a)))
+a <- a[order(abs(a$log2FoldChange), decreasing = T),]
+  a$rank.l2FC <- seq(1:length(rownames(a)))
+a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
+  a <- a[order(a$rank.sum),]
+print(a)
+}
+
+top.res1 <- topgenes(res=res1)
+summary(res1)
+top.res2 <- topgenes(res=res2)
+summary(res2)
+top.res3 <- topgenes(res=res3)
+summary(res3)
+top.res4 <- topgenes(res=res4)
+summary(res4)
+top.res5 <- topgenes(res=res5)
+summary(res5)
+top.res6 <- topgenes(res=res6)
+summary(res6)
+top.res7 <- topgenes(res=res7)
+summary(res7)
+
+top.all <- rbind(top.res1,
+         top.res2,
+         top.res3,
+         top.res4,
+         top.res5,
+         top.res6,
+         top.res7)
+top.all <- top.all[order(top.all$rank.sum),]
+head(top.all, n=185) %>% rownames() %>% unique() %>% length()
+top.all <- top.all[unique(rownames(top.all)),]
+
+# top.all <- head(top.all, n=100)
+
+## Shrinkage
+res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
+res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
+res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
+res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
+res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
+res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
+res7_shrink <- lfcShrink(dds, res=res7, type="ashr")
+
+```
+
+### combine L2FC with padj
+```{r}
+
+res3_shrink
+subset(mcols(dds), id.symbol =="SELU1") %>% rownames()
+res3_shrink["Cre06.g263550",]
+
+# DataFrame with 1 row and 5 columns
+#                baseMean log2FoldChange     lfcSE      pvalue        padj
+#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
+# Cre06.g263550   3566.14        2.90397  0.167321 1.72033e-71 1.97322e-68
+
+plot(res3_shrink$log2FoldChange,-log(res3_shrink$padj))
+res3_shrink$logpadj <- -log(res3_shrink$padj, base=10)
+
+-log(1.38379e-06,base = 10)
+
+x=-2.5
+60/(3^2)
+
+y=-8*x^2 + 50
+y
+
+res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < res3_shrink$logpadj,TRUE,FALSE)
+summary(res3_shrink$ishs)
+subset(res3_shrink,ishs==TRUE & logpadj< 50 & abs(log2FoldChange)<2.5)
+subset(res3_shrink,ishs==TRUE & logpadj< 50)
+
+# for log2 -/+ 5 and padj 100 
+res1_shrink$ishs <- ifelse((-20/3*(res1_shrink$log2FoldChange^2) + 60) < -log(res1_shrink$padj, base=10),TRUE,FALSE)
+res2_shrink$ishs <- ifelse((-20/3*(res2_shrink$log2FoldChange^2) + 60) < -log(res2_shrink$padj, base=10),TRUE,FALSE)
+res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < -log(res3_shrink$padj, base=10),TRUE,FALSE)
+res4_shrink$ishs <- ifelse((-20/3*(res4_shrink$log2FoldChange^2) + 60) < -log(res4_shrink$padj, base=10),TRUE,FALSE)
+res5_shrink$ishs <- ifelse((-20/3*(res5_shrink$log2FoldChange^2) + 60) < -log(res5_shrink$padj, base=10),TRUE,FALSE)
+res6_shrink$ishs <- ifelse((-20/3*(res6_shrink$log2FoldChange^2) + 60) < -log(res6_shrink$padj, base=10),TRUE,FALSE)
+res7_shrink$ishs <- ifelse((-20/3*(res7_shrink$log2FoldChange^2) + 60) < -log(res7_shrink$padj, base=10),TRUE,FALSE)
+
+```
+
+```{r}
+outdir
+figures <- paste(outdir,"figures",sep="/")
+dir.create(figures)
+
+```
+
+## Colours
+```{r}
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+```
+
+## PCA (Panel B)
+```{r}
+library("PCAtools")
+library(ggalt)
+
+vsd <- vst(dds, blind=FALSE)
+length(vsd)
+
+dubs <- duplicated(mcols(vsd)$id.symbol)
+length(dubs)
+
+unique <- unique(mcols(vsd)$id.symbol)
+length(unique)
+
+vsd2 <- vsd[!duplicated(mcols(vsd)$id.symbol)]
+rownames(vsd2) <- mcols(vsd2)$id.symbol
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+names(group.colors) <- levels(colData(dds)$condition)
+group.colors
+
+plotPCA(vsd2, intgroup="condition")
+vst <- assay(vsd2)
+p <- pca(vst, metadata = colData(dds), removeVar = 0.1)
+screeplot(p, axisLabSize = 18, titleLabSize = 22)
+biplot(p, colby = "condition")
+biplot(p, showLoadings = TRUE,labSize = 5, pointSize = 5, sizeLoadingsNames = 5)
+
+biplot(p,x= "PC1", y="PC3",
+       lab = p$metadata$sample,
+    colby = 'condition', colkey = group.colors,
+    encircle = TRUE,
+    encircleFill = TRUE,
+    legendPosition = 'right',
+    labSize = 4)
+ggsave(paste(figures,"2023_12_PCA1-3.png",sep="/"),
+  width = 6,
+  height = 6)
+
+biplot(p,x= "PC3", y="PC4",
+       lab = p$metadata$sample,
+    colby = 'condition', colkey = group.colors,
+    encircle = TRUE,
+    encircleFill = TRUE,
+    legendPosition = 'right')
+ggsave(paste(figures,"2023_12_PCA3-4.pdf",sep="/"),
+  width = 12,
+  height = 12)
+
+pairsplot(p,colby = 'condition', colkey = group.colors)
+ggsave(paste(figures,"2023_12_pairsplot.pdf",sep="/"),
+  width = 12,
+  height = 12)
+
+plotloadings(p,
+    rangeRetain = 0.01,
+    labSize = 2.0,
+    title = 'Loadings plot',
+    subtitle = 'PC1, PC2, PC3, PC4, PC5',
+    caption = 'Top 1% variables',
+    drawConnectors = TRUE)
+ggsave(paste(figures,"2023_12_loadings.pdf",sep="/"),
+  width = 12,
+  height = 12)
+
+# Determine optimum number of PCs to retain
+elbow <- findElbowPoint(p$variance)
+elbow
+
+anno["Cre01.g016600",]
+
+ biplot(p,
+    lab = p$metadata$condition,
+    colby = 'condition', colkey = group.colors,
+    hline = 0, vline = 0,
+    encircle = TRUE, encircleFill = TRUE,
+    legendLabSize = 16, legendIconSize = 8.0,
+    legendPosition = 'right')
+
+pairsplot(p,
+    components = getComponents(p, c(1:6)),
+    triangle = TRUE, trianglelabSize = 12,
+    hline = 0, vline = 0,
+    pointSize = 2,
+    gridlines.major = FALSE, gridlines.minor = FALSE,
+    colby = 'condition', colkey = group.colors,
+    title = 'Pairs plot', plotaxes = FALSE,
+    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))
+
+```
+
+## Counts
+```{r}
+
+# TF genes
+
+goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
+anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
+anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
+anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
+summary(anno_tf$gene_id %in% rownames(dds))
+anno_tf[!anno_tf$gene_id %in% rownames(dds),]
+anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]
+
+# Combined counts table TFs
+
+goi <- anno_tf
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+all_counts$Gene <- factor(all_counts$Gene)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
+levels(all_counts$Gene)
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+max_val <- 1.0*max(all_counts$count)
+
+# Plot
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
   geom_boxplot(fatten = 1) +
-  scale_fill_viridis_d() +
+  scale_fill_manual(values = group.colors) +
   scale_y_continuous(trans = "log2")
-gcounts2
+gcounts
+ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
+  width = 12,
+  height = 8)
 
-gcounts3 <- ggplot(all_counts[all_counts$Gene=="ROC40",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
-  geom_boxplot(color="black") +
-  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
-  scale_fill_manual(values=anno_colors$treatment) +
-  scale_color_manual(values=anno_colors$treatment) +
-  scale_y_continuous(trans = "log2") +
-  labs(title = "ROC40") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-gcounts3
+## Phot genes all-in-one
+goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",
 
-gcounts4 <- ggplot(all_counts[all_counts$Gene=="ROC114",], aes(x = condition, y = count, fill=treatment, color=treatment)) +
-  geom_boxplot(color="black") +
-  geom_point(shape=21,color="black",aes(fill=treatment),position=position_dodge(width=0.75), alpha=1) +
-  scale_fill_manual(values=anno_colors$treatment) +
-  scale_color_manual(values=anno_colors$treatment) +
-  scale_y_continuous(trans = "log2") +
-  labs(title = "ROC114") +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-gcounts4
+anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
+rownames(anno_phot) <- anno_phot$geneSymbol
+anno_phot <- anno_phot[goi_phot,]
+
+# Combined counts table TFs
+
+goi <- anno_phot
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
+levels(all_counts$Gene)
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+max_val <- 1.0*max(all_counts$count)
+
+# Plot
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  scale_y_continuous(trans = "log2")
+gcounts
+ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
+  width = 12,
+  height = 8)
+
+```
+
+## Volcanos
+### dark
+```{r}
+
+# dark
+res <- res1_shrink
+n <- 'WT vs. pCRY in dark'
+nf <- 'figures/Volcano_dark.png'
+l <- nrow(top.res1)
+pcol <- "black"
+lcol <- colours[1+1]
+
+rownames(res) <- mcols(dds)$id.symbol
+res <- res[!is.na(res$padj),]
+res <- res[!is.na(res$log2FoldChange),]
+
+shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
+                ifelse(res$padj < 1e-100,6,16))
+summary(is.na(shape))
+# shape[is.na(shape)] <- 2
+names(shape)[shape == 6] <- 'outliner'
+names(shape)[shape == 16] <- 'in range'
+
+res$log2FoldChange[res$log2FoldChange > 5] <- 5
+res$log2FoldChange[res$log2FoldChange < -5] <- -5
+res$padj[res$padj < 1e-100] <- 1e-100
+summary(res$padj < 1e-100)
+
+keyvals <- ifelse(
+    res$ishs == TRUE, pcol,
+      'grey')
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == pcol] <- 'highly significant'
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+evd <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = rownames(res),
+    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
+    # selectLab = anno_tf$geneSymbol,
+    # boxedLabels = TRUE,
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 10,
+    colCustom = keyvals,
+    xlim = c(-5.1, 5.1),
+    ylim = c(0, 101),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("DE genes:",l),
+    # sub = "SVF",
+    pCutoff = 10^(-60),
+    FCcutoff = 3,
+   shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
+                   'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+evd
+
+```
+
+### blue
+```{r}
+
+# blue
+res <- res2_shrink
+n <- 'WT vs. pCRY in blue'
+nf <- 'figures/Volcano_blue.png'
+l <- nrow(top.res2)
+pcol <- colours[3]
+lcol <- colours[3+1]
+
+rownames(res) <- mcols(dds)$id.symbol
+res <- res[!is.na(res$padj),]
+res <- res[!is.na(res$log2FoldChange),]
+
+shape <- ifelse(abs(res$log2FoldChange) > 5, 17,
+                ifelse(res$padj < 1e-100,17,16))
+summary(is.na(shape))
+# shape[is.na(shape)] <- 2
+names(shape)[shape == 6] <- 'outliner'
+names(shape)[shape == 16] <- 'in range'
+
+res$log2FoldChange[res$log2FoldChange > 5] <- 5
+res$log2FoldChange[res$log2FoldChange < -5] <- -5
+res$padj[res$padj < 1e-100] <- 1e-100
+summary(res$padj < 1e-100)
+
+keyvals <- ifelse(
+    res$ishs == TRUE, pcol,
+      'grey')
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == pcol] <- 'highly significant'
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+evb <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = rownames(res),
+    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
+    # selectLab = anno_tf$geneSymbol,
+    # boxedLabels = TRUE,
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 10,
+    colCustom = keyvals,
+    xlim = c(-5.1, 5.1),
+    ylim = c(0, 101),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("DE genes:",l),
+    # sub = "SVF",
+    pCutoff = 10^(-50),
+    FCcutoff = 2.5,
+   shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
+                   'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+evb
+
+```
+
+### red
+```{r}
+
+# red
+res <- res3_shrink
+n <- 'WT vs. pCRY in red'
+nf <- 'figures/Volcano_r.png'
+l <- nrow(top.res3)
+pcol <- colours[5]
+lcol <- colours[5+1]
+
+rownames(res) <- mcols(dds)$id.symbol
+res <- res[!is.na(res$padj),]
+res <- res[!is.na(res$log2FoldChange),]
+
+shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
+                ifelse(res$padj < 1e-100,6,16))
+summary(is.na(shape))
+# shape[is.na(shape)] <- 2
+names(shape)[shape == 6] <- 'outliner'
+names(shape)[shape == 16] <- 'in range'
+
+res$log2FoldChange[res$log2FoldChange > 5] <- 5
+res$log2FoldChange[res$log2FoldChange < -5] <- -5
+res$padj[res$padj < 1e-100] <- 1e-100
+summary(res$padj < 1e-100)
+
+keyvals <- ifelse(
+    res$ishs == TRUE, pcol,
+      'grey')
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == pcol] <- 'highly significant'
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+evr <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = rownames(res),
+    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
+    # selectLab = anno_tf$geneSymbol,
+    # boxedLabels = TRUE,
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 10,
+    colCustom = keyvals,
+    xlim = c(-5.1, 5.1),
+    ylim = c(0, 101),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("DE genes:",l),
+    # sub = "SVF",
+    pCutoff = 10^(-50),
+    FCcutoff = 2.5,
+   shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
+                   'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+evr
+
+#ggsave(file="figures/Vulcano_rtest.pdf",
+#       width = 12,
+#       height = 12)
+
+```
+
+### all
+```{r}
+
+ga <- ggarrange(evb,
+                evd+rremove("ylab"),
+                evr+rremove("ylab"),
+                ncol = 3, nrow = 1,common.legend = TRUE, legend = "bottom",
+                widths = c(1.05,1,1))
+
+ggsave(ga, file="figures/Vulcano_all.pdf",
+       width = 12,
+       height = 8)
+
+# + ggplot2::coord_cartesian(xlim=c(-10, 10))
+# ggplot2::coord_cartesian(xlim=c(-5, 5)
+
+# with ggplot
+ggplot(as.data.frame(resLFC_WT_BL_vs.pcry_BL),aes(log2FoldChange,-log(padj))) +
+  geom_point() +
+  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rownames(top_WT_BL_vs.pcry_BL),]), colour = "blue") + 
+  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), colour = "red") +
+  geom_text_repel(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), aes(label=anno[rocs,"id.symbol"]),colour = "red",hjust=-0.5, vjust=-1) +
+  xlim(-5,5)
+getwd()
+ggsave("graphs/Vulcano ROCs-2.pdf",
+       width = 10,
+       height = 6)
+
+## enhanced vulcano
+library(airway)
+  library(magrittr)
+
+ library('DESeq2')
+
+  dds <- DESeqDataSet(airway, design = ~ cell + dex)
+  dds <- DESeq(dds, betaPrior=FALSE)
+  res <- results(dds,
+    contrast = c('dex','trt','untrt'))
+  res <- lfcShrink(dds,
+    contrast = c('dex','trt','untrt'), res=res, type = 'normal')
+ EnhancedVolcano(res,
+    lab = rownames(res),
+    x = 'log2FoldChange',
+    y = 'pvalue')
+
+```
+
+## Heatmap
+
+### plot heatmap
+```{r}
+library("pheatmap")
+ntd <- normTransform(dds)
+top.all.hm <- head(top.all, n=100)
+df <- assay(ntd)[rownames(top.all.hm),]
+rownames(df) <- mcols(dds)[rownames(df),"id.symbol"]
+
+# anno_col <- as.data.frame(colData(dds)[,c("condition","treatment","genotype")])
+anno_col <- data.frame(condition = dds$condition, treatment = dds$treatment)
+rownames(anno_col) <- rownames(colData(dds))
+
+# column order
+anno_col <- arrange(anno_col, match(anno_col$condition, levels(anno_col$condition)))
+levels(anno_col$condition)
+df <- df[,rownames(anno_col)]
+
+annot_colors=list(condition=c(Cancer="#F0978D",Healthy="#63D7DE"))
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+
+mycolors2 <- list(genotype = c("#525252","#969696"),
+     treatment = c("#525252","#1F78B4","#E31A1C"),
+     condition = colours)
+
+names(mycolors2$genotype) <- levels(anno_col$genotype)
+names(mycolors2$treatment) <- levels(anno_col$treatment)
+names(mycolors2$condition) <- levels(anno_col$condition)
+mycolors2
+
+xx <- pheatmap(df, cluster_rows=TRUE, show_rownames=TRUE,
+         cluster_cols=FALSE,
+         annotation_col=anno_col,
+         color=colorRampPalette(c("navy", "white", "red"))(50),
+         annotation_colors=mycolors2
+         )
+
+ggsave("figures/2023_12_Heatmap_top.png",plot=xx,
+       width = 5,
+       height = 15)
+
+```
+
+### Complex Heatmap
+```{r}
+
+df
+genes <- anno_tf
+
+# collapse replicates
+colData(dds)
+dds_heat <- collapseReplicates(dds, dds$condition,dds$sample)
+dds_heat$sample <- factor(dds_heat$sample, levels=c("WT-D1","pcry-D1","WT-BL1","pcry-BL1","WT-R1","pcry-R1"))
+colData(dds_heat)
+dds_heat$runsCollapsed
+colnames(dds_heat)
+matchFirstLevel <- dds_heat$sample == levels(dds_heat$sample)[1]
+stopifnot(all(rowSums(counts(dds_heat[,matchFirstLevel])) == counts(dds_heat[,1])))
+
+mat <- counts(dds_heat)
+length(rownames(counts(dds_heat)))
+length(rownames(df))
+rownames(mat) <- mcols(dds_heat)$id.symbol
+length(rownames(mat))
+
+mat <- mat[df,]
+
+position <- mat
+l <- length(rownames(mat))
+pos <- data.frame(genes = rownames(mat),pos = 1:l)
+rownames(pos) <- rownames(mat)
+pos <- pos[genes,2]
+
+# mat <- cbind(res,counts(dds_heat))
+# mat <- subset(mat,!SYMBOL2=="" & SYMBOL2 %in% df$SYMBOL2)
+# rownames(mat) <- mat$SYMBOL2
+# mat <- mat[,c(12:15)]
+heat <- t(scale(t(mat)))
+Heatmap(heat)
+
+summary(heat)
+myCol <- colorRampPalette(c('brown3', 'yellow2', 'darkolivegreen3'))(100)
+myBreaks <- seq(-1.5, 1.5, length.out = 100)
+
+ann <- data.frame(condition = colData(dds_heat)[,"condition"],
+  stringsAsFactors = FALSE)
+
+colAnn <- HeatmapAnnotation(
+  df = ann,
+  which = 'col', # 'col' (samples) or 'row' (gene) annotation?
+  na_col = 'white', # default colour for any NA values in the annotation data-frame, 'ann'
+  # col = colours,
+  annotation_height = 0.6,
+  annotation_width = unit(1, 'cm'),
+  gap = unit(1, 'mm'))
+
+boxplotCol <- HeatmapAnnotation(
+  boxplot = anno_boxplot(
+    heat,
+    border = FALSE,
+    gp = gpar(fill = '#CCCCCC'),
+    pch = '.',
+    size = unit(2, 'mm'),
+    axis = TRUE,
+    axis_param = list(
+      gp = gpar(fontsize = 12),
+      side = 'left')),
+  annotation_width = unit(c(2.0), 'cm'),
+  which = 'col')
+
+boxplotRow <- HeatmapAnnotation(
+  boxplot = row_anno_boxplot(
+    heat,
+    border = FALSE,
+    gp = gpar(fill = '#CCCCCC'),
+    pch = '.',
+    size = unit(2, 'mm'),
+    axis = TRUE,
+    axis_param = list(
+      gp = gpar(fontsize = 12),
+      side = 'top')),
+  annotation_width = unit(c(2.0), 'cm'),
+  which = 'row')
+
+genelabels <- rowAnnotation(
+  Genes = anno_mark(
+    at = pos,
+    labels = genes,
+    labels_gp = gpar(fontsize = 15, fontface = 'bold'),
+    padding = 0.75),
+  width = unit(2.0, 'cm') +
+
+    max_text_width(
+      rownames(heat)[seq(1, nrow(heat), 40)],
+      gp = gpar(fontsize = 10,  fontface = 'bold')))
+
+# Clusters
+# pamClusters <- cluster::pam(heat, k = 4) # pre-select k = 4 centers
+pamClusters <- cluster::pam(heat, k= 8) # pre-select k = 4 centers
+
+pamClusters$clustering <- paste0('Cluster ', pamClusters$clustering)
+pamClusters$clustering <- factor(pamClusters$clustering,
+                                 levels = c('Cluster 1', 'Cluster 2', 'Cluster 3', 'Cluster 4','Cluster 5', 'Cluster 6'))
+
+hmap <- Heatmap(heat,
+
+                # split the genes / rows according to the PAM clusters
+                split = pamClusters$clustering,
+                cluster_row_slices = TRUE,
+
+                name = 'Gene\nZ-\nscore',
+
+                col = colorRamp2(myBreaks, myCol),
+
+                # parameters for the colour-bar that represents gradient of expression
+                heatmap_legend_param = list(
+                  color_bar = 'continuous',
+                  legend_direction = 'vertical',
+                  legend_width = unit(8, 'cm'),
+                  legend_height = unit(5.0, 'cm'),
+                  title_position = 'topcenter',
+                  title_gp=gpar(fontsize = 12, fontface = 'bold'),
+                  labels_gp=gpar(fontsize = 12, fontface = 'bold')),
+
+                # row (gene) parameters
+                cluster_rows = TRUE,
+                show_row_dend = TRUE,
+                row_title = 'Statistically significant genes',
+                row_title_side = 'left',
+                row_title_gp = gpar(fontsize = 12,  fontface = 'bold'),
+                row_title_rot = 90,
+                show_row_names = FALSE,
+                row_names_gp = gpar(fontsize = 10, fontface = 'bold'),
+                row_names_side = 'left',
+                row_dend_width = unit(25,'mm'),
+
+                # column (sample) parameters
+                cluster_columns = TRUE,
+                show_column_dend = TRUE,
+                column_title = '',
+                column_title_side = 'bottom',
+                column_title_gp = gpar(fontsize = 12, fontface = 'bold'),
+                column_title_rot = 0,
+                show_column_names = TRUE,
+                column_names_gp = gpar(fontsize = 10, fontface = 'bold'),
+                column_names_max_height = unit(10, 'cm'),
+                column_dend_height = unit(25,'mm'),
+
+                # cluster methods for rows and columns
+                clustering_distance_columns = function(x) as.dist(1 - cor(t(x))),
+                clustering_method_columns = 'ward.D2',
+                clustering_distance_rows = function(x) as.dist(1 - cor(t(x))),
+                clustering_method_rows = 'ward.D2',
+
+                # specify top and bottom annotations
+                top_annotation = colAnn)
+                # bottom_annotation = boxplotCol)
+
+pdf("Heatmap2.pdf", height = 18, width = 6)
+draw(hmap + genelabels,
+     heatmap_legend_side = 'left',
+     annotation_legend_side = 'right',
+     row_sub_title_side = 'left')
+dev.off()
+
+```
+
+# Export
+
+```{r}
+
+library(writexl)
+
+# combine results
+
+res <- res_WT_BL_vs.pcry_BL
+res1 <- res_WT_D_vs.pcry_D
+res2 <- res_WT_BL_vs.pcry_BL
+res3 <- res_WT_R_vs.pcry_R
+res4 <- res_WT_D_vs.WT_BL
+res5 <- res_WT_D_vs.WT_R
+
+colnames(anno)
+
+res_exp <- data.frame(
+  # general
+  "gene_id" = rownames(dds),
+  "gene_name" = anno[rownames(dds),"geneSymbol"],
+  "Alias" = anno[rownames(dds),"prev.symbols"],
+  "id.symbol" = anno[rownames(dds),"id.symbol"],
+ "Description" = anno[rownames(dds),"Description"],
+ "Comments" = anno[rownames(dds),"Comments"],
+ "TargetP" = anno[rownames(dds),"TargetP"],
+ "Predalgo" = anno[rownames(dds),"Predalgo"],
+ "baseMean" = res$baseMean,
+ # results 1
+ "L2FC.WT_D_vs.pcry_D" = res1$log2FoldChange,
+ "pvalue.WT_D_vs.pcry_D" = res1$pvalue,
+ "padj.WT_D_vs.pcry_D" = res1$padj,
+  # results 2
+ "L2FC.WT_BL_vs.pcry_BL" = res2$log2FoldChange,
+ "pvalue.WT_BL_vs.pcry_BL" = res2$pvalue,
+ "padj.WT_BL_vs.pcry_BL" = res2$padj,
+  # results 3
+ "L2FC.WT_R_vs.pcry_R" = res3$log2FoldChange,
+ "pvalue.WT_R_vs.pcry_R" = res3$pvalue,
+ "padj.WT_R_vs.pcry_R" = res3$padj,
+  # results 4
+ "L2FC.WT_D_vs.WT_BL" = res4$log2FoldChange,
+ "pvalue.WT_D_vs.WT_BL" = res4$pvalue,
+ "padj.WT_D_vs.WT_BL" = res4$padj,
+  # results 5
+ "L2FC.WT_D_vs.WT_R" = res5$log2FoldChange,
+ "pvalue.WT_D_vs.WT_R" = res5$pvalue,
+ "padj.WT_D_vs.WT_R" = res5$padj,
+  counts(dds, normalized=TRUE))
+res_exp
+
+outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
+
+write_xlsx(data.frame(res_exp),
+           paste(outdir,"2023_08 P3044 results.xlsx",sep="/"))
+
+# Counts only
+counts <- round(counts(dds, normalized=TRUE))
+colnames(counts)
+colData(dds)$sample.name <- paste(colData(dds)$condition,colData(dds)$repetition,sep="_")
+colnames(counts) == colData(dds)$sample.id
+colnames(counts) <- colData(dds)$sample.name
+
+write.csv(counts, file.path(outdir,"Counts.csv"))
+
+Exp_design <- data.frame(names = colData(dds)$sample.name,
+                         genotype = colData(dds)$genotype,
+                         treatment = colData(dds)$treatment,
+                         condition = colData(dds)$condition) %>% t()
+
+Exp_design <- matrix(colData(dds)$sample.name,
+                     colData(dds)$genotype,
+                     colData(dds)$treatment,
+                    colData(dds)$condition)
+
+rownames(Exp_design)[1] <- {1}
+
+write.table(Exp_design,file.path(outdir,"Exp_design.csv"), sep=",",  col.names=FALSE)
 

### 2024-04-22 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
vulcano start

@@
 gitdir <- scriptdir %>% dirname() %>% dirname()
-dir <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq aCRY"
+
+if(Sys.info()["sysname"]== "Linux"){
+  s <- "/mnt/s"
+  dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")} else {
+    if(Sys.info()["sysname"]== "Darwin"){
+       s <- "/mnt/s"
+       dir <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq aCRY"} else {
+       s <- "S:" 
+       dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")}}
 
@@
 # 3. Results
-
-## Plot Counts
-```{r}
-
-# TF genes
-
-goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
-anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
-anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
-anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
-summary(anno_tf$gene_id %in% rownames(dds))
-anno_tf[!anno_tf$gene_id %in% rownames(dds),]
-anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]
-
-# Combined counts table TFs
-
-goi <- anno_tf
-l <- nrow(goi)
-all_counts <- {}
-for (i in 1:l){
-  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
-  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
-  d$sample <- rownames(d)
-  rownames(d) <- {}
-  all_counts <- bind_rows(all_counts,d)
-  }
-
-all_counts$Gene
-levels(all_counts$condition)
-all_counts$Gene <- factor(all_counts$Gene)
-levels(all_counts$Gene)
-
-all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
-levels(all_counts$Gene)
-
-# colours
-Greys <- brewer.pal(9, name="Greys")[c(7,5)]
-Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
-colours <- c(Greys,Paired)
-group.colors <- colours
-group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
-max_val <- 1.0*max(all_counts$count)
-
-# Plot
-gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
-  geom_boxplot(fatten = 1) +
-  scale_fill_manual(values = group.colors) +
-  scale_y_continuous(trans = "log2")
-gcounts
-ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
-  width = 12,
-  height = 8)
-
-## Phot genes all-in-one
-goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",
-
-anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
-rownames(anno_phot) <- anno_phot$geneSymbol
-anno_phot <- anno_phot[goi_phot,]
-
-# Combined counts table TFs
-
-goi <- anno_phot
-l <- nrow(goi)
-all_counts <- {}
-for (i in 1:l){
-  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
-  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
-  d$sample <- rownames(d)
-  rownames(d) <- {}
-  all_counts <- bind_rows(all_counts,d)
-  }
-
-all_counts$Gene
-levels(all_counts$condition)
-levels(all_counts$Gene)
-
-all_counts$Gene <- factor(all_counts$Gene)
-levels(all_counts$Gene)
-
-all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
-levels(all_counts$Gene)
-
-# colours
-Greys <- brewer.pal(9, name="Greys")[c(7,5)]
-Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
-colours <- c(Greys,Paired)
-group.colors <- colours
-group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
-max_val <- 1.0*max(all_counts$count)
-
-# Plot
-gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
-  geom_boxplot(fatten = 1) +
-  scale_fill_manual(values = group.colors) +
-  scale_y_continuous(trans = "log2")
-gcounts
-ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
-  width = 12,
-  height = 8)
-
-```
-
-# ---------------------------
-# COPY & PASTE from pCRY
 ## Make results
 ```{r, eval=FALSE, include=FALSE, echo=FALSE}
-outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
-setwd(outdir)
-
-dds <- readRDS(paste(outdir,"dds_23design.RDS",sep="/"))
-anno <- readRDS(file=paste(outdir,"anno.RDS", sep="/"))
 
@@
 
-### combine L2FC with padj
-```{r}
-
-res3_shrink
-subset(mcols(dds), id.symbol =="SELU1") %>% rownames()
-res3_shrink["Cre06.g263550",]
-
-# DataFrame with 1 row and 5 columns
-#                baseMean log2FoldChange     lfcSE      pvalue        padj
-#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
-# Cre06.g263550   3566.14        2.90397  0.167321 1.72033e-71 1.97322e-68
-
-plot(res3_shrink$log2FoldChange,-log(res3_shrink$padj))
-res3_shrink$logpadj <- -log(res3_shrink$padj, base=10)
-
--log(1.38379e-06,base = 10)
-
-x=-2.5
-60/(3^2)
-
-y=-8*x^2 + 50
-y
-
-res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < res3_shrink$logpadj,TRUE,FALSE)
-summary(res3_shrink$ishs)
-subset(res3_shrink,ishs==TRUE & logpadj< 50 & abs(log2FoldChange)<2.5)
-subset(res3_shrink,ishs==TRUE & logpadj< 50)
-
-# for log2 -/+ 5 and padj 100 
-res1_shrink$ishs <- ifelse((-20/3*(res1_shrink$log2FoldChange^2) + 60) < -log(res1_shrink$padj, base=10),TRUE,FALSE)
-res2_shrink$ishs <- ifelse((-20/3*(res2_shrink$log2FoldChange^2) + 60) < -log(res2_shrink$padj, base=10),TRUE,FALSE)
-res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < -log(res3_shrink$padj, base=10),TRUE,FALSE)
-res4_shrink$ishs <- ifelse((-20/3*(res4_shrink$log2FoldChange^2) + 60) < -log(res4_shrink$padj, base=10),TRUE,FALSE)
-res5_shrink$ishs <- ifelse((-20/3*(res5_shrink$log2FoldChange^2) + 60) < -log(res5_shrink$padj, base=10),TRUE,FALSE)
-res6_shrink$ishs <- ifelse((-20/3*(res6_shrink$log2FoldChange^2) + 60) < -log(res6_shrink$padj, base=10),TRUE,FALSE)
-res7_shrink$ishs <- ifelse((-20/3*(res7_shrink$log2FoldChange^2) + 60) < -log(res7_shrink$padj, base=10),TRUE,FALSE)
-
-```
-
-```{r}
-outdir
-figures <- paste(outdir,"figures",sep="/")
-dir.create(figures)
-
-```
-
-## Colours
-```{r}
-Greys <- brewer.pal(9, name="Greys")[c(7,5)]
-Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
-colours <- c(Greys,Paired)
-```
-
-## PCA (Panel B)
-```{r}
-library("PCAtools")
-library(ggalt)
-
-vsd <- vst(dds, blind=FALSE)
-length(vsd)
-
-dubs <- duplicated(mcols(vsd)$id.symbol)
-length(dubs)
-
-unique <- unique(mcols(vsd)$id.symbol)
-length(unique)
-
-vsd2 <- vsd[!duplicated(mcols(vsd)$id.symbol)]
-rownames(vsd2) <- mcols(vsd2)$id.symbol
-
-# colours
-Greys <- brewer.pal(9, name="Greys")[c(7,5)]
-Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
-colours <- c(Greys,Paired)
-group.colors <- colours
-group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
-names(group.colors) <- levels(colData(dds)$condition)
-group.colors
-
-plotPCA(vsd2, intgroup="condition")
-vst <- assay(vsd2)
-p <- pca(vst, metadata = colData(dds), removeVar = 0.1)
-screeplot(p, axisLabSize = 18, titleLabSize = 22)
-biplot(p, colby = "condition")
-biplot(p, showLoadings = TRUE,labSize = 5, pointSize = 5, sizeLoadingsNames = 5)
-
-biplot(p,x= "PC1", y="PC3",
-       lab = p$metadata$sample,
-    colby = 'condition', colkey = group.colors,
-    encircle = TRUE,
-    encircleFill = TRUE,
-    legendPosition = 'right',
-    labSize = 4)
-ggsave(paste(figures,"2023_12_PCA1-3.png",sep="/"),
-  width = 6,
-  height = 6)
-
-biplot(p,x= "PC3", y="PC4",
-       lab = p$metadata$sample,
-    colby = 'condition', colkey = group.colors,
-    encircle = TRUE,
-    encircleFill = TRUE,
-    legendPosition = 'right')
-ggsave(paste(figures,"2023_12_PCA3-4.pdf",sep="/"),
-  width = 12,
-  height = 12)
-
-pairsplot(p,colby = 'condition', colkey = group.colors)
-ggsave(paste(figures,"2023_12_pairsplot.pdf",sep="/"),
-  width = 12,
-  height = 12)
-
-plotloadings(p,
-    rangeRetain = 0.01,
-    labSize = 2.0,
-    title = 'Loadings plot',
-    subtitle = 'PC1, PC2, PC3, PC4, PC5',
-    caption = 'Top 1% variables',
-    drawConnectors = TRUE)
-ggsave(paste(figures,"2023_12_loadings.pdf",sep="/"),
-  width = 12,
-  height = 12)
-
-# Determine optimum number of PCs to retain
-elbow <- findElbowPoint(p$variance)
-elbow
-
-anno["Cre01.g016600",]
-
- biplot(p,
-    lab = p$metadata$condition,
-    colby = 'condition', colkey = group.colors,
-    hline = 0, vline = 0,
-    encircle = TRUE, encircleFill = TRUE,
-    legendLabSize = 16, legendIconSize = 8.0,
-    legendPosition = 'right')
-
-pairsplot(p,
-    components = getComponents(p, c(1:6)),
-    triangle = TRUE, trianglelabSize = 12,
-    hline = 0, vline = 0,
-    pointSize = 2,
-    gridlines.major = FALSE, gridlines.minor = FALSE,
-    colby = 'condition', colkey = group.colors,
-    title = 'Pairs plot', plotaxes = FALSE,
-    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))
-
-```
-
-## Counts
+## Plot Counts
 ```{r}
@@
     y = 'pvalue')
+
+```
+
+# ---------------------------
+# COPY & PASTE from pCRY
+## Make results
+```{r, eval=FALSE, include=FALSE, echo=FALSE}
+outdir <- "C:/Users/kelterbs/OneDrive - Charité - Universitätsmedizin Berlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044"
+setwd(outdir)
+
+dds <- readRDS(paste(outdir,"dds_23design.RDS",sep="/"))
+anno <- readRDS(file=paste(outdir,"anno.RDS", sep="/"))
+
+design(dds)
+resultsNames(dds)
+cbind(resultsNames(dds),c(rep(0,6)))
+
+# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))
+
+# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
+res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
+res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")
+
+res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))
+
+res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))
+
+# Unterschiede zwischen versch. Belichtungen über beide Genotypen
+# D vs. (BL+R)
+res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))
+
+# D vs. BL
+res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))
+
+# D vs. R
+res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))
+
+# WT vs. pCRY
+res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))
+
+## Make top lists:
+
+topgenes <- function(res){
+a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
+a$id.symbol <- anno[rownames(a),"id.symbol"]
+a$symbol <- anno[rownames(a),"geneSymbol"]
+a <- a[order(a$baseMean, decreasing = T),]
+  a$rank.bm <- seq(1:length(rownames(a)))
+a <- a[order(a$padj, decreasing = F),]
+  a$rank.padj <- seq(1:length(rownames(a)))
+a <- a[order(abs(a$log2FoldChange), decreasing = T),]
+  a$rank.l2FC <- seq(1:length(rownames(a)))
+a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
+  a <- a[order(a$rank.sum),]
+print(a)
+}
+
+top.res1 <- topgenes(res=res1)
+summary(res1)
+top.res2 <- topgenes(res=res2)
+summary(res2)
+top.res3 <- topgenes(res=res3)
+summary(res3)
+top.res4 <- topgenes(res=res4)
+summary(res4)
+top.res5 <- topgenes(res=res5)
+summary(res5)
+top.res6 <- topgenes(res=res6)
+summary(res6)
+top.res7 <- topgenes(res=res7)
+summary(res7)
+
+top.all <- rbind(top.res1,
+         top.res2,
+         top.res3,
+         top.res4,
+         top.res5,
+         top.res6,
+         top.res7)
+top.all <- top.all[order(top.all$rank.sum),]
+head(top.all, n=185) %>% rownames() %>% unique() %>% length()
+top.all <- top.all[unique(rownames(top.all)),]
+
+# top.all <- head(top.all, n=100)
+
+## Shrinkage
+res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
+res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
+res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
+res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
+res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
+res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
+res7_shrink <- lfcShrink(dds, res=res7, type="ashr")
+
+```
+
+### combine L2FC with padj
+```{r}
+
+res3_shrink
+subset(mcols(dds), id.symbol =="SELU1") %>% rownames()
+res3_shrink["Cre06.g263550",]
+
+# DataFrame with 1 row and 5 columns
+#                baseMean log2FoldChange     lfcSE      pvalue        padj
+#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
+# Cre06.g263550   3566.14        2.90397  0.167321 1.72033e-71 1.97322e-68
+
+plot(res3_shrink$log2FoldChange,-log(res3_shrink$padj))
+res3_shrink$logpadj <- -log(res3_shrink$padj, base=10)
+
+-log(1.38379e-06,base = 10)
+
+x=-2.5
+60/(3^2)
+
+y=-8*x^2 + 50
+y
+
+res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < res3_shrink$logpadj,TRUE,FALSE)
+summary(res3_shrink$ishs)
+subset(res3_shrink,ishs==TRUE & logpadj< 50 & abs(log2FoldChange)<2.5)
+subset(res3_shrink,ishs==TRUE & logpadj< 50)
+
+# for log2 -/+ 5 and padj 100 
+res1_shrink$ishs <- ifelse((-20/3*(res1_shrink$log2FoldChange^2) + 60) < -log(res1_shrink$padj, base=10),TRUE,FALSE)
+res2_shrink$ishs <- ifelse((-20/3*(res2_shrink$log2FoldChange^2) + 60) < -log(res2_shrink$padj, base=10),TRUE,FALSE)
+res3_shrink$ishs <- ifelse((-20/3*(res3_shrink$log2FoldChange^2) + 60) < -log(res3_shrink$padj, base=10),TRUE,FALSE)
+res4_shrink$ishs <- ifelse((-20/3*(res4_shrink$log2FoldChange^2) + 60) < -log(res4_shrink$padj, base=10),TRUE,FALSE)
+res5_shrink$ishs <- ifelse((-20/3*(res5_shrink$log2FoldChange^2) + 60) < -log(res5_shrink$padj, base=10),TRUE,FALSE)
+res6_shrink$ishs <- ifelse((-20/3*(res6_shrink$log2FoldChange^2) + 60) < -log(res6_shrink$padj, base=10),TRUE,FALSE)
+res7_shrink$ishs <- ifelse((-20/3*(res7_shrink$log2FoldChange^2) + 60) < -log(res7_shrink$padj, base=10),TRUE,FALSE)
+
+```
+
+```{r}
+outdir
+figures <- paste(outdir,"figures",sep="/")
+dir.create(figures)
+
+```
+
+## Colours
+```{r}
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+```
+
+## PCA (Panel B)
+```{r}
+library("PCAtools")
+library(ggalt)
+
+vsd <- vst(dds, blind=FALSE)
+length(vsd)
+
+dubs <- duplicated(mcols(vsd)$id.symbol)
+length(dubs)
+
+unique <- unique(mcols(vsd)$id.symbol)
+length(unique)
+
+vsd2 <- vsd[!duplicated(mcols(vsd)$id.symbol)]
+rownames(vsd2) <- mcols(vsd2)$id.symbol
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+names(group.colors) <- levels(colData(dds)$condition)
+group.colors
+
+plotPCA(vsd2, intgroup="condition")
+vst <- assay(vsd2)
+p <- pca(vst, metadata = colData(dds), removeVar = 0.1)
+screeplot(p, axisLabSize = 18, titleLabSize = 22)
+biplot(p, colby = "condition")
+biplot(p, showLoadings = TRUE,labSize = 5, pointSize = 5, sizeLoadingsNames = 5)
+
+biplot(p,x= "PC1", y="PC3",
+       lab = p$metadata$sample,
+    colby = 'condition', colkey = group.colors,
+    encircle = TRUE,
+    encircleFill = TRUE,
+    legendPosition = 'right',
+    labSize = 4)
+ggsave(paste(figures,"2023_12_PCA1-3.png",sep="/"),
+  width = 6,
+  height = 6)
+
+biplot(p,x= "PC3", y="PC4",
+       lab = p$metadata$sample,
+    colby = 'condition', colkey = group.colors,
+    encircle = TRUE,
+    encircleFill = TRUE,
+    legendPosition = 'right')
+ggsave(paste(figures,"2023_12_PCA3-4.pdf",sep="/"),
+  width = 12,
+  height = 12)
+
+pairsplot(p,colby = 'condition', colkey = group.colors)
+ggsave(paste(figures,"2023_12_pairsplot.pdf",sep="/"),
+  width = 12,
+  height = 12)
+
+plotloadings(p,
+    rangeRetain = 0.01,
+    labSize = 2.0,
+    title = 'Loadings plot',
+    subtitle = 'PC1, PC2, PC3, PC4, PC5',
+    caption = 'Top 1% variables',
+    drawConnectors = TRUE)
+ggsave(paste(figures,"2023_12_loadings.pdf",sep="/"),
+  width = 12,
+  height = 12)
+
+# Determine optimum number of PCs to retain
+elbow <- findElbowPoint(p$variance)
+elbow
+
+anno["Cre01.g016600",]
+
+ biplot(p,
+    lab = p$metadata$condition,
+    colby = 'condition', colkey = group.colors,
+    hline = 0, vline = 0,
+    encircle = TRUE, encircleFill = TRUE,
+    legendLabSize = 16, legendIconSize = 8.0,
+    legendPosition = 'right')
+
+pairsplot(p,
+    components = getComponents(p, c(1:6)),
+    triangle = TRUE, trianglelabSize = 12,
+    hline = 0, vline = 0,
+    pointSize = 2,
+    gridlines.major = FALSE, gridlines.minor = FALSE,
+    colby = 'condition', colkey = group.colors,
+    title = 'Pairs plot', plotaxes = FALSE,
+    margingaps = unit(c(-0.01, -0.01, -0.01, -0.01), 'cm'))
+
+```
+
+## Counts
+```{r}
+
+# TF genes
+
+goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
+anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
+anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
+anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
+summary(anno_tf$gene_id %in% rownames(dds))
+anno_tf[!anno_tf$gene_id %in% rownames(dds),]
+anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]
+
+# Combined counts table TFs
+
+goi <- anno_tf
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+all_counts$Gene <- factor(all_counts$Gene)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
+levels(all_counts$Gene)
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+max_val <- 1.0*max(all_counts$count)
+
+# Plot
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  scale_y_continuous(trans = "log2")
+gcounts
+ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
+  width = 12,
+  height = 8)
+
+## Phot genes all-in-one
+goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",
+
+anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
+rownames(anno_phot) <- anno_phot$geneSymbol
+anno_phot <- anno_phot[goi_phot,]
+
+# Combined counts table TFs
+
+goi <- anno_phot
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
+levels(all_counts$Gene)
+
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+max_val <- 1.0*max(all_counts$count)
+
+# Plot
+gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  scale_y_continuous(trans = "log2")
+gcounts
+ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
+  width = 12,
+  height = 8)
 

### 2024-04-22 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
local copy macbook

--- 2024-04-21 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
@@
 knit: (function(input_file, encoding) {
+    rmarkdown::render(input_file,output_file= 'Readme.md')
         })
-    rmarkdown::render(input_file,output_file= 'Readme.md')
 ---
@@
 # 3. Results
+## Colors
+```{r}
+# colours
+Greys <- brewer.pal(9, name="Greys")[c(7,5)]
+Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
+colours <- c(Greys,Paired)
+group.colors <- colours
+group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")
+
+anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
+                    genotype = c("green4","orchid","tan2"))
+
+group.colours2 <- c("#00B500","#007E56","#805B00","#FF00AB","#8024AB","#FF0056","tan2","brown","orangered")
+
+```
 
 ## Plot Counts
-```{r}
+```{r plot_counts}
 
@@
 for (i in 1:l){
-  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup="condition", col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
   d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
@@
 
-# colours
-Greys <- brewer.pal(9, name="Greys")[c(7,5)]
-Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
-colours <- c(Greys,Paired)
-group.colors <- colours
-group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
 max_val <- 1.0*max(all_counts$count)
@@
 # Plot
-gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
   geom_boxplot(fatten = 1) +
@@
   scale_y_continuous(trans = "log2")
-gcounts
-ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
-  width = 12,
-  height = 8)
+gcounts1
+
+gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = anno_colors$genotype) +
+  scale_y_continuous(trans = "log2")
+gcounts2
+
+gcounts1 / gcounts2
+
+# ggsave(paste(figures,"2023_12_counts_tfs_aio_log2.pdf",sep="/"), plot = gcounts,
+# width = 12,
+# height = 8)
 
@@
 
-# colours
-Greys <- brewer.pal(9, name="Greys")[c(7,5)]
-Paired <- brewer.pal(12, name="Paired")[c(2,1,6,5)]
-colours <- c(Greys,Paired)
-group.colors <- colours
-group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
+# Plot
 max_val <- 1.0*max(all_counts$count)
-
-# Plot
-gcounts <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
   geom_boxplot(fatten = 1) +
@@
   scale_y_continuous(trans = "log2")
-gcounts
-ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
-  width = 12,
-  height = 8)
+gcounts1
+
+gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = anno_colors$genotype) +
+  scale_y_continuous(trans = "log2")
+gcounts2
+
+gcounts3 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colours2) +
+  scale_y_continuous(trans = "log2")
+gcounts3
+
+gcounts1 / gcounts2 / gcounts3
+
+# ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
+#   width = 12,
+#   height = 8)
 
@@
 ### combine L2FC with padj
-```{r}
+```{r, eval=FALSE}
 
@@
 
-```{r}
+```{r, eval=FALSE}
 outdir
@@
 ## Colours
-```{r}
+```{r, eval=FALSE}
 Greys <- brewer.pal(9, name="Greys")[c(7,5)]
@@
 ## PCA (Panel B)
-```{r}
+```{r, eval=FALSE}
 library("PCAtools")
@@
 ## Counts
-```{r}
+```{r, eval=FALSE}
 
@@
 ### dark
-```{r}
+```{r, eval=FALSE}
 
@@
 ### blue
-```{r}
+```{r, eval=FALSE}
 
@@
 ### red
-```{r}
+```{r, eval=FALSE}
 
@@
 ### all
-```{r}
+```{r, eval=FALSE}
 
@@
 ### plot heatmap
-```{r}
+```{r, eval=FALSE}
 library("pheatmap")
@@
 ### Complex Heatmap
-```{r}
+```{r, eval=FALSE}
 
@@
 
-```{r}
+```{r, eval=FALSE}
 

### 2024-04-23 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
merge files

--- 2024-04-22 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
@@
 knit: (function(input_file, encoding) {
+    rmarkdown::render(input_file,output_file= 'Readme.md')
       })
-    rmarkdown::render(input_file,output_file= 'Readme.md')
 ---

### 2024-04-23 salmon_index.R

@@ -1,0 +1,48 @@
+
+# get Chlamydomonas genomic data
+# https://data.jgi.doe.gov/refine-download/phytozome?q=Chlamydomonas+reinhardtii+CC-4532+v6.1
+# (Create Phytozome Account)
+
+## Files
+# Genome (DNA):            CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz
+# Transcriptome (RNA):     CreinhardtiiCC_4532_707_v6.1.transcript.fa.gz
+# Gene Info:               CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz
+# Master Annotation: 	     CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv
+
+genomicdir=""
+genome=""
+transcripts=""
+genes=""
+annotation=""
+
+# create list of chromosomes
+cd $genomicdir
+grep "^>" <(gunzip -c $genome) | cut -d " " -f 1 > decoys_chlamy_v6.1.txt
+
+# to modify, copy to linux partition
+mkdir ~/decoys
+cp decoys_chlamy_v6.1.txt ~/decoys
+
+# remove ">"
+sed -i.bak -e 's/>//g' ~/decoys/decoys_chlamy_v6.1.txt
+
+# Optional: check txt file
+vim ~/decoys/decoys_chlamy_v6.1.txt
+# to exit: -press 'ESC', type ":q!", press 'ENTER'
+
+# copy back
+cp ~/decoys/decoys_chlamy_v6.1.txt $genomicdir
+
+# combine cdna ncdna and dna in one file
+cd $genomicdir
+
+cat $genome $transcripts > gentrome_chlamy_v6.1.fa.gz
+
+# make index with salmon
+# is salmon activated?
+conda activate salmon
+cd $gdir
+indexname=chlamy_v6.1_index
+echo $indexname
+
+salmon index -t gentrome_chlamy_v6.1.fa.gz -d decoys_chlamy_v6.1.txt -p 10 -i $indexname

### 2024-04-23 2_B_dive_cia5/2_dive_cia5.Rmd
rename folder

--- 2024-04-16 2_A_Results_cia5/2_results_cia5.Rmd
@@
 
+dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3043",sep="/")
+gitdir <- paste(dir,"git_Chlamy_RNAseq_plap6",sep="/")
+datadir <- paste(dir,"data",sep="/")
+outdir <- gitdir
+pubdir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3043/git_Chlamy_RNAseq_plap6/pub_figures",sep="/")
+
 fastqdir <- paste(dir,"fastq",sep="/")
@@
 ```{r load_data}
-load(file=paste(dir,"dds_cia5.RDS", sep="/"))
-load(paste(dir,"anno.RDS",sep="/"))
+# load(paste(dir,"dds_cia5.RDS", sep="/"))
+# load(paste(dir,"anno.RDS",sep="/"))
+
+load("~/S/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/dds_cia5.RDS")
+load("~/S/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/anno.RDS")
 
@@
    scale_y_continuous(limits = c(60,150),trans = "log2") +
-  scale_fill_manual(values = c("grey30","salmon")) + 
+  scale_fill_manual(values = c("grey30","mediumpurple2")) + 
     labs(title = paste("PLAP6 (",PLAP6,")",sep=""))
@@
    scale_y_continuous(trans = "log2") +
-  scale_fill_manual(values = c("grey30","salmon")) + 
+  scale_fill_manual(values = c("grey30","mediumpurple2")) + 
 #  scale_color_manual(values = c("grey30","salmon")) + 
@@
 g2
-g1+g2 + plot_layout(guides = "collect", axes="collect", axis_titles="collect") & theme(legend.position = 'right')
+p <- g1+g2 + plot_layout(guides = "collect", axes="collect", axis_titles="collect") & theme(legend.position = 'right')
+
+```
+
+###### -export
+```{r export_plap6, eval=FALSE}
+ggexport(g1, filename = paste(pubdir,"Counts_cia5_PLAP6.pdf",sep="/"),width = 8.2, height = 4.7)
+ggsave(g1, filename = paste(pubdir,"Counts_cia5_PLAP6.png",sep="/"),width = 8.2, height = 4.7)
+
+ggexport(g2, filename = paste(pubdir,"Counts_cia5_rbcl.pdf",sep="/"),width = 8.2, height = 4.7)
+ggsave(g2, filename = paste(pubdir,"Counts_cia5_rbcl.png",sep="/"),width = 8.2, height = 4.7)
+
+ggexport(p, filename = paste(pubdir,"Counts_cia5_PLAP6+rbcl.pdf",sep="/"),width = 8.2, height = 4.7)
+ggsave(p, filename = paste(pubdir,"Counts_cia5_PLAP6+rbcl.png",sep="/"),width = 8.2, height = 4.7)
 

### 2024-04-26 2_A_dive_aCRY_pCRY (combined)/2_A_dive_aCRY_pCRY.Rmd
acry pcry separate

--- 2024-04-22 2_A_dive_aCRY_pCRY/2_A_dive_aCRY_pCRY.Rmd
@@
 ```{r r_libraries}
+library(patchwork)
 library(PCAtools)
@@
 library(viridis)
-library(patchwork)
+
 library(ggpubr)
@@
 # Seq 1
-tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm"
+ifelse(Sys.info()["sysname"]== "Linux",
+       tximeta_pcry_file <- dir %>% dirname() %>% paste(.,"P3044/tximeta.txm",sep="/"),
+    ifelse(Sys.info()["sysname"]== "Darwin",
+       tximeta_pcry_file <- "/Users/simonkelterborn/Library/CloudStorage/OneDrive-Charité-UniversitätsmedizinBerlin/NGS (RNA-Seq & CRISPR)/Chlamy RNA-Seq P3044/tximeta.txm",
+       tximeta_pcry_file <- "S:/AG/AG-Scholz-NGS/Daten/Simon/P3044/tximeta.txm"))
+
 load(file=tximeta_pcry_file)
@@
 
-# colData(dds) %>% head() %>% kable()
-# add metadata
-# colData(dds)$strain <- factor(colData(dds)$strain)
-
-# ## combine technical replicates #####\n
-# colData(dds)
-# dds <- collapseReplicates(dds, dds$sampleid,dds$lane)
-# colData(dds)
-# dds$runsCollapsed
-# colnames(dds)
-# matchFirstLevel <- dds$sampleid == levels(factor(dds$sampleid))[1]
-# all(rowSums(counts(dds[,matchFirstLevel])) == counts(dds[,1]))
+# remove outlier
+# outliner <- c("WTdark1")
+# colData(dds)[outliner,]
+# keep <- colnames(dds) %>% .[!str_detect(.,outliner)]
+# dds <- dds[,keep]
 
@@
 
-# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
+# save(dds, file = paste(dir,"dds_acry_pcry_all.RDS",sep="/"), compress = FALSE)
 # save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
@@
 ntd <- normTransform(dds)
-rld <- rlog(dds, blind=FALSE) #regularized logarithm
+# rld <- rlog(dds, blind=FALSE) #regularized logarithm
+# save(rld, file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"), compress = FALSE)
 ```
@@
 ```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
-# load(file=paste(data,"rlog.rld", sep="/"))
+load(file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"))
 meanSdPlot(counts(dds, normalized =TRUE))
@@
 meanSdPlot(assay(rld))
+
+# (m1+m2) / (m3)
+# patchwork::wrap_plots(m1)
 
@@
 anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
-anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
+anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
                     genotype = c("green4","orchid","tan2"))
-
+#                    "ivory","yellow1","gold"
 names(anno_colors$treatment) <- levels(anno_col$treatment)
@@
 
+# -> how to handel batch effects: https://www.biostars.org/p/266507/
+
 ```
@@
 # find explaining PCs
-horn <- parallelPCA(vst_dat)
-# horn$n = 2
+# horn <- parallelPCA(vst_dat)
+# horn$n = 5 # aCRY&pCRY
 # warnings()
-
-# horn <- list()
-# horn$n <- 2
+# 
+horn <- list()
+horn$n <- 5
 
@@
 group.colors <- c("#525252", "#969696", "#1F78B4", "#A6CEE3", "#E31A1C", "#FB9A99")
-group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")
+# group.colors <- c("grey10","grey40","grey80","royalblue4","royalblue2", "lightblue2",  "darkred","brown2","salmon1")
 
@@
 group.colours2 <- c("#00B500","#007E56","#805B00","#FF00AB","#8024AB","#FF0056","tan2","brown","orangered")
-
 ```
@@
 #   height = 8)
+
+```
+
+## Make results
+```{r, eval=FALSE, include=FALSE, echo=FALSE}
+
+design(dds)
+resultsNames(dds)
+cbind(resultsNames(dds),c(rep(0,resultsNames(dds) %>% length())))
+
+# res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))
+
+# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
+res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
+res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")
+
+res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))
+
+res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))
+
+# Unterschiede zwischen versch. Belichtungen über beide Genotypen
+# D vs. (BL+R)
+res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))
+
+# D vs. BL
+res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))
+
+# D vs. R
+res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))
+
+# WT vs. pCRY
+res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))
+
+## Make top lists:
+
+topgenes <- function(res){
+a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
+a$id.symbol <- anno[rownames(a),"id.symbol"]
+a$symbol <- anno[rownames(a),"geneSymbol"]
+a <- a[order(a$baseMean, decreasing = T),]
+  a$rank.bm <- seq(1:length(rownames(a)))
+a <- a[order(a$padj, decreasing = F),]
+  a$rank.padj <- seq(1:length(rownames(a)))
+a <- a[order(abs(a$log2FoldChange), decreasing = T),]
+  a$rank.l2FC <- seq(1:length(rownames(a)))
+a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
+  a <- a[order(a$rank.sum),]
+print(a)
+}
+
+top.res1 <- topgenes(res=res1)
+summary(res1)
+top.res2 <- topgenes(res=res2)
+summary(res2)
+top.res3 <- topgenes(res=res3)
+summary(res3)
+top.res4 <- topgenes(res=res4)
+summary(res4)
+top.res5 <- topgenes(res=res5)
+summary(res5)
+top.res6 <- topgenes(res=res6)
+summary(res6)
+top.res7 <- topgenes(res=res7)
+summary(res7)
+
+top.all <- rbind(top.res1,
+         top.res2,
+         top.res3,
+         top.res4,
+         top.res5,
+         top.res6,
+         top.res7)
+top.all <- top.all[order(top.all$rank.sum),]
+head(top.all, n=185) %>% rownames() %>% unique() %>% length()
+top.all <- top.all[unique(rownames(top.all)),]
+
+# top.all <- head(top.all, n=100)
+
+## Shrinkage
+res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
+res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
+res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
+res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
+res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
+res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
+res7_shrink <- lfcShrink(dds, res=res7, type="ashr")
+
+```
+
+## Volcanos
+### dark
+```{r, eval=FALSE}
+
+# dark
+res <- res1_shrink
+n <- 'WT vs. pCRY in dark'
+nf <- 'figures/Volcano_dark.png'
+l <- nrow(top.res1)
+pcol <- "black"
+lcol <- colours[1+1]
+
+rownames(res) <- mcols(dds)$id.symbol
+res <- res[!is.na(res$padj),]
+res <- res[!is.na(res$log2FoldChange),]
+
+shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
+                ifelse(res$padj < 1e-100,6,16))
+summary(is.na(shape))
+# shape[is.na(shape)] <- 2
+names(shape)[shape == 6] <- 'outliner'
+names(shape)[shape == 16] <- 'in range'
+
+res$log2FoldChange[res$log2FoldChange > 5] <- 5
+res$log2FoldChange[res$log2FoldChange < -5] <- -5
+res$padj[res$padj < 1e-100] <- 1e-100
+summary(res$padj < 1e-100)
+
+keyvals <- ifelse(
+    res$ishs == TRUE, pcol,
+      'grey')
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == pcol] <- 'highly significant'
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+evd <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = rownames(res),
+    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
+    # selectLab = anno_tf$geneSymbol,
+    # boxedLabels = TRUE,
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 10,
+    colCustom = keyvals,
+    xlim = c(-5.1, 5.1),
+    ylim = c(0, 101),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("DE genes:",l),
+    # sub = "SVF",
+    pCutoff = 10^(-60),
+    FCcutoff = 3,
+   shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
+                   'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+evd
+
+```
+
+### blue
+```{r eval=FALSE}
+
+# blue
+res <- res2_shrink
+n <- 'WT vs. pCRY in blue'
+nf <- 'figures/Volcano_blue.png'
+l <- nrow(top.res2)
+pcol <- colours[3]
+lcol <- colours[3+1]
+
+rownames(res) <- mcols(dds)$id.symbol
+res <- res[!is.na(res$padj),]
+res <- res[!is.na(res$log2FoldChange),]
+
+shape <- ifelse(abs(res$log2FoldChange) > 5, 17,
+                ifelse(res$padj < 1e-100,17,16))
+summary(is.na(shape))
+# shape[is.na(shape)] <- 2
+names(shape)[shape == 6] <- 'outliner'
+names(shape)[shape == 16] <- 'in range'
+
+res$log2FoldChange[res$log2FoldChange > 5] <- 5
+res$log2FoldChange[res$log2FoldChange < -5] <- -5
+res$padj[res$padj < 1e-100] <- 1e-100
+summary(res$padj < 1e-100)
+
+keyvals <- ifelse(
+    res$ishs == TRUE, pcol,
+      'grey')
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == pcol] <- 'highly significant'
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+evb <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = rownames(res),
+    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
+    # selectLab = anno_tf$geneSymbol,
+    # boxedLabels = TRUE,
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 10,
+    colCustom = keyvals,
+    xlim = c(-5.1, 5.1),
+    ylim = c(0, 101),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("DE genes:",l),
+    # sub = "SVF",
+    pCutoff = 10^(-50),
+    FCcutoff = 2.5,
+   shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
+                   'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+evb
+
+```
+
+### red
+```{r}
+
+# red
+res <- res3_shrink
+n <- 'WT vs. pCRY in red'
+nf <- 'figures/Volcano_r.png'
+l <- nrow(top.res3)
+pcol <- colours[5]
+lcol <- colours[5+1]
+
+rownames(res) <- mcols(dds)$id.symbol
+res <- res[!is.na(res$padj),]
+res <- res[!is.na(res$log2FoldChange),]
+
+shape <- ifelse(abs(res$log2FoldChange) > 5, 6,
+                ifelse(res$padj < 1e-100,6,16))
+summary(is.na(shape))
+# shape[is.na(shape)] <- 2
+names(shape)[shape == 6] <- 'outliner'
+names(shape)[shape == 16] <- 'in range'
+
+res$log2FoldChange[res$log2FoldChange > 5] <- 5
+res$log2FoldChange[res$log2FoldChange < -5] <- -5
+res$padj[res$padj < 1e-100] <- 1e-100
+summary(res$padj < 1e-100)
+
+keyvals <- ifelse(
+    res$ishs == TRUE, pcol,
+      'grey')
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == pcol] <- 'highly significant'
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+evr <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = rownames(res),
+    selectLab  = rownames(res)[which(names(keyvals) == 'highly significant')],
+    # selectLab = anno_tf$geneSymbol,
+    # boxedLabels = TRUE,
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 10,
+    colCustom = keyvals,
+    xlim = c(-5.1, 5.1),
+    ylim = c(0, 101),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("DE genes:",l),
+    # sub = "SVF",
+    pCutoff = 10^(-50),
+    FCcutoff = 2.5,
+   shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2.5','p-adj < 0.05',
+                   'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+evr
+
+#ggsave(file="figures/Vulcano_rtest.pdf",
+#       width = 12,
+#       height = 12)
+
+```
+
+### all
+```{r}
+
+ga <- ggarrange(evb,
+                evd+rremove("ylab"),
+                evr+rremove("ylab"),
+                ncol = 3, nrow = 1,common.legend = TRUE, legend = "bottom",
+                widths = c(1.05,1,1))
+
+ggsave(ga, file="figures/Vulcano_all.pdf",
+       width = 12,
+       height = 8)
+
+# + ggplot2::coord_cartesian(xlim=c(-10, 10))
+# ggplot2::coord_cartesian(xlim=c(-5, 5)
+
+# with ggplot
+ggplot(as.data.frame(resLFC_WT_BL_vs.pcry_BL),aes(log2FoldChange,-log(padj))) +
+  geom_point() +
+  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rownames(top_WT_BL_vs.pcry_BL),]), colour = "blue") + 
+  geom_point(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), colour = "red") +
+  geom_text_repel(data = as.data.frame(resLFC_WT_BL_vs.pcry_BL[rocs,]), aes(label=anno[rocs,"id.symbol"]),colour = "red",hjust=-0.5, vjust=-1) +
+  xlim(-5,5)
+getwd()
+ggsave("graphs/Vulcano ROCs-2.pdf",
+       width = 10,
+       height = 6)
+
+## enhanced vulcano
+library(airway)
+  library(magrittr)
+
+ library('DESeq2')
+
+  dds <- DESeqDataSet(airway, design = ~ cell + dex)
+  dds <- DESeq(dds, betaPrior=FALSE)
+  res <- results(dds,
+    contrast = c('dex','trt','untrt'))
+  res <- lfcShrink(dds,
+    contrast = c('dex','trt','untrt'), res=res, type = 'normal')
+ EnhancedVolcano(res,
+    lab = rownames(res),
+    x = 'log2FoldChange',
+    y = 'pvalue')
 

### 2024-04-26 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
acry pcry separate

--- 2024-04-26 2_A_dive_aCRY_pCRY (combined)/2_A_dive_aCRY_pCRY.Rmd
@@
 ---
-title: "RNAseq aCRY pCRY data dive"
+title: "RNAseq aCRY pCRY data dive (separate)"
 author: "Kelterborn"
@@
     toc: true
+    toc_depth: 2
 always_allow_html: true
@@
 
-# Seq 1
+# Seq 1 (pcry)
 ifelse(Sys.info()["sysname"]== "Linux",
@@
 load(file=tximeta_pcry_file)
+
+colData(gse)$treatment <- colData(gse)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
+  str_replace(pattern="BL", replacement = "blue") %>%
+  str_replace(pattern="R", replacement = "red") %>% factor(levels = c("dark","blue","red"))
+
+colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","pcry_dark","WT_blue","pcry_blue","WT_red","pcry_red"))
+
+names(colData(gse)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
+colData(gse)$experiment <- "pcry"
 gse_pcry <- gse
@@
 
-# combine
-( mcols(gse_acry) == mcols(gse_pcry) ) %>% summary()
-# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
-names(colData(gse_pcry)) <- c("names","clientId","filename","clientName","genotype","treatment","condition","replicate","lane","RNA_conc","mappingrates")
-colData(gse_acry) <- colData(gse_acry)[,c(1:4,7:12)]
-colData(gse_pcry)$experiment <- "pcry"
-colData(gse_pcry) <- colData(gse_pcry)[,c(1,3,2,4:6,8,12,7,11)]  
-# (( colData(gse_acry) %>% names() ) == ( colData(gse_pcry) %>% names() ) ) %>% summary()
-
-# now same metadata
-gse <- cbind(gse_acry,gse_pcry)
-
-colData(gse)$genotype
-
-colData(gse)$treatment <- colData(gse)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
+colData(gse_acry)$treatment <- colData(gse_acry)$treatment %>% str_replace(pattern="D", replacement = "dark") %>%
   str_replace(pattern="BL", replacement = "blue") %>%
@@
 
-colData(gse)$condition <- paste(colData(gse)$genotype,colData(gse)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","pcry_dark","WT_blue","acry_blue","pcry_blue","WT_red","acry_red","pcry_red"))
-
-colData(gse)$experiment
+colData(gse_acry)$condition <- paste(colData(gse_acry)$genotype,colData(gse_acry)$treatment,sep="_") %>% factor(levels=c("WT_dark","acry_dark","WT_blue","acry_blue","WT_red","acry_red"))
 
@@
 ```{r dds_acry_pcry}
+
 # design <- ~condition
+exp <- c("acry","pcry")
+gse_list <- list(gse_acry,gse_pcry)
+names(gse_list) <- exp
+dds_list <- list()
+i <- 1
+for (e in names(gse_list)){
+ gse <- gse_list[[e]]
  design <- ~genotype+treatment+genotype:treatment
-
 dds <- DESeqDataSet(gse, design=design)
@@
 
-plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$experiment)
-
-head(mcols(dds)$geneSymbol)
+plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$condition)
+
+assign(paste("dds",e,sep="_"),dds)
+dds_list[[e]] <- dds
+}
 
@@
 
-# save(dds, file = paste(dir,"dds_acry_pcry_all.RDS",sep="/"), compress = FALSE)
+# save(dds_list, file = paste(dir,"dds_list.RDS",sep="/"), compress = FALSE)
+# save(dds, file = paste(dir,"dds_acry_pcry.RDS",sep="/"), compress = FALSE)
 # save(dds, file = paste(dir,"dds_acry.RDS",sep="/"), compress = FALSE)
@@
 ```{r pre_trans, eval=TRUE, include=FALSE, echo=FALSE}
-vsd <- vst(dds, blind=FALSE) #Variance stabilized transformation
-ntd <- normTransform(dds)
-# rld <- rlog(dds, blind=FALSE) #regularized logarithm
-# save(rld, file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"), compress = FALSE)
+
+vsd_list <- lapply(dds_list,vst,blind=FALSE)
+ntd_list <- lapply(dds_list,normTransform)
+rld_list <- lapply(dds_list,rlog, blind=FALSE)
+# save(rld_list, file = paste(dir,"rlog_list.RDS",sep="/"), compress = FALSE)
+
 ```
@@
 ```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
-load(file = paste(dir,"rlog_aCRY_pCRY.RDS",sep="/"))
-meanSdPlot(counts(dds, normalized =TRUE))
-meanSdPlot(assay(ntd))
-meanSdPlot(assay(vsd))
-meanSdPlot(assay(rld))
-
-# (m1+m2) / (m3)
-# patchwork::wrap_plots(m1)
+# load(file = paste(dir,"rlog_list.RDS",sep="/"))
+msdp1 <- lapply(lapply(dds_list,assay,normalized =TRUE),meanSdPlot)
+msdp2 <- lapply(lapply(vsd_list,assay,normalized =TRUE),meanSdPlot)
+msdp3 <- lapply(lapply(ntd_list,assay,normalized =TRUE),meanSdPlot)
+msdp4 <- lapply(lapply(rld_list,assay,normalized =TRUE),meanSdPlot)
 
@@
 ```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
+vsd <- vsd_list[[1]]
+for (vsd in vsd_list){
 sampleDists <- dist(t(assay(vsd)))
@@
 anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
-                    genotype = c("green4","orchid","tan2"))
-#                    "ivory","yellow1","gold"
+                    genotype = c("green4","orchid"))
 names(anno_colors$treatment) <- levels(anno_col$treatment)
@@
          cutree_cols = 6,
-         fontsize_row = 5)
-
+         fontsize_row = 5) %>% print()
+}
 # -> how to handel batch effects: https://www.biostars.org/p/266507/
@@
 # load(file=paste(data,"deseq2.dds", sep="/"))
-vst_dat <- assay(vst(dds))
+
+for (vsd in vsd_list){
 
@@
 
-g2 <- ggplot(pcaData, aes(PC1, PC2, color=experiment, shape=genotype)) +
+g2 <- ggplot(pcaData, aes(PC1, PC2, color=genotype, shape=genotype)) +
   geom_point(size=5, alpha=0.7) +
-  labs(title = "experiment") +
+  labs(title = "genotype") +
   xlab(paste0("PC1: ",percentVar[1],"% variance")) +
@@
 
-g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype)) +
+g3 <- ggplot(pcaData, aes(PC1, PC2, color=condition, shape=genotype, label=clientName)) +
   geom_point(size=5, alpha=0.7) +
+  geom_text_repel() +
   labs(title = "condition") +
@@
 
-((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect")
+gp <- ((g1+g2) / (g3)) + plot_layout(guides = "collect", axis_titles="collect") & plot_annotation(title = colData(vsd)$experiment[1])
+gp %>% print()
+}
 
@@
 ```{r pca_advanced, warning = FALSE, fig.show="hold", out.width="80%", fig.height=6, eval=TRUE, include=TRUE, echo=FALSE}
-# transform data
-# calculate PCA (all data)
-## https://www.bioconductor.org/packages/devel/bioc/vignettes/PCAtools/inst/doc/PCAtools.html#modify-bi-plots
-
-# p <- pca(vst_dat, metadata = colData(dds))
-# vst <- assay(vst(dds))
+
+for (dds in dds_list) {
+vst_dat <- assay(vst(dds))
 p <- pca(vst_dat, metadata = colData(dds), removeVar = 0.1)
@@
 # check different PCAs
-p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 15,trianglelabSize = 6,
+p1 <- pairsplot(p,colby = 'treatment', colkey = anno_colors$treatment,title = 'treatment',titleLabSize = 8,trianglelabSize = 6,
                 hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
@@
 
-p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 15,trianglelabSize = 6,
+p2 <- pairsplot(p,colby = 'genotype', colkey = anno_colors$genotype,title = 'genotype',titleLabSize = 8,trianglelabSize = 6,
                 hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
@@
 
-p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 15,trianglelabSize = 6,
+p4 <- pairsplot(p,colby = 'condition', colkey = viridis(length(levels(colData(dds)$condition))),title = 'condition',titleLabSize = 8,trianglelabSize = 6,
                 hline = 0, vline = 0,gridlines.major = FALSE, gridlines.minor = FALSE,
@@
 
-(p1+p2) / (p4)
-
+pairs <- (p1+p2) / (p4) + plot_annotation(title = colData(dds)$experiment[1])
+pairs %>% print()
 # Determine optimum number of PCs to retain
@@
 # find explaining PCs
-# horn <- parallelPCA(vst_dat)
-# horn$n = 5 # aCRY&pCRY
+horn <- parallelPCA(vst_dat)
+# horn$n = 4 #aCRY
 # warnings()
-# 
-horn <- list()
-horn$n <- 5
-
-screeplot(p,
+
+# horn <- list()
+# horn$n <- 4
+
+s <- screeplot(p,
     components = getComponents(p, 1:20),
@@
     geom_label(aes(x = elbow + 1, y = 50,
-      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8))
+      label = paste('Elbow',elbow,sep="="), vjust = -1, size = 8)) + plot_annotation(title = colData(dds)$experiment[1])
+
+s %>% print()
 
 bi <- biplot(p,x="PC3",y="PC1",
-    lab = p$metadata$condition,
+    lab = p$metadata$clientName,
     colby = 'condition',colkey = viridis(length(levels(colData(dds)$condition))),
@@
 )
-pairs + bi + 
-  plot_layout(design = layout)
+pairs_bi <- pairs + bi + 
+  plot_layout(design = layout) & plot_annotation(title = colData(dds)$experiment[1])
+pairs_bi %>% print()
 
@@
 
-plotloadings(p,
+load <- plotloadings(p,
+                     axisLabSize = 8,
              components = getComponents(p)[1:7],
              col = viridis(3),
-             labSize = 3)
+             labSize = 3) + 
+  plot_annotation(title = colData(dds)$experiment[1]) &
+  theme(legend.position = 'none', 
+        axis.title=element_text(size=8),
+        axis.text=element_text(size=6))
+load %>% print()
+}
 
@@
 ```{r plot_counts}
-
+dds <- dds_list[[1]]
+for (dds in dds_list){
 # TF genes
@@
 # Plot
-gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+gcounts_tf1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
   geom_boxplot(fatten = 1) +
   scale_fill_manual(values = group.colors) +
-  scale_y_continuous(trans = "log2")
-gcounts1
-
-gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
-  geom_boxplot(fatten = 1) +
-  scale_fill_manual(values = anno_colors$genotype) +
-  scale_y_continuous(trans = "log2")
-gcounts2
-
-gcounts1 / gcounts2
+  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
+gcounts_tf1 %>% print()
 
@@
 max_val <- 1.0*max(all_counts$count)
-gcounts1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+gcounts_phot1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
   geom_boxplot(fatten = 1) +
   scale_fill_manual(values = group.colors) +
-  scale_y_continuous(trans = "log2")
-gcounts1
-
-gcounts2 <- ggplot(all_counts, aes(x = Gene, y = count, fill=genotype)) +
-  geom_boxplot(fatten = 1) +
-  scale_fill_manual(values = anno_colors$genotype) +
-  scale_y_continuous(trans = "log2")
-gcounts2
-
-gcounts3 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
-  geom_boxplot(fatten = 1) +
-  scale_fill_manual(values = group.colours2) +
-  scale_y_continuous(trans = "log2")
-gcounts3
-
-gcounts1 / gcounts2 / gcounts3
+  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
+gcounts_phot1 %>% print()
+
+print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))
+
+}
 
@@
 ```{r, eval=FALSE, include=FALSE, echo=FALSE}
+
+res_list_list <- list()
+for (dds in dds_list) {
 

### 2024-04-26 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
figure fix

@@
 ## -R_libraries
-```{r r_libraries}
+```{r r_libraries, imnclude=FALSE}
 library(patchwork)
@@
 ## -R_folders
-```{r r_folders}
+```{r r_folders, include = FALSE}
 
@@
 ## - Tximeta files
-```{r get_tximeta_files}
+```{r get_tximeta_files, include=FALSE}
 
@@
 ## Run Deseq2
-```{r dds_acry_pcry}
+```{r dds_acry_pcry, echo=FALSE}
 
@@
 par(mfrow=c(1,2))
-hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
+hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10), main = colData(dds)$experiment[1]) 
 hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
@@
 
-plotDispEsts(dds)
+plotDispEsts(dds, main = colData(dds)$experiment[1])
 resultsNames(dds)
-DESeq2::plotMA(dds)
-
-plotCounts(dds, gene = "Cre01.g000150", intgroup = "condition", col=colData(dds)$condition)
+DESeq2::plotMA(dds, main=colData(dds)$experiment[1])
+
+ROC40 <- "Cre06.g275350"
+plotCounts(dds, gene = ROC40, intgroup = "condition", col=colData(dds)$treatment, main = paste("ROC40 in Δ",colData(dds)$experiment[1],sep = ""))
 
@@
 
-```{r pre_trans_fig, figures-side, fig.show="hold", out.width="25%"}
+```{r pre_trans_fig, figures-side, fig.show="hold", out.width="50%"}
 # load(file = paste(dir,"rlog_list.RDS",sep="/"))
@@
 ### - Check sample distance
-```{r pre_sample_dist, fig.height=6, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
+```{r pre_sample_dist, fig.height=4, out.width="100%", eval=TRUE, include=TRUE, echo=FALSE}
 vsd <- vsd_list[[1]]
@@
 ### - Perform principal component analysis
-```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=7, eval=TRUE, include=TRUE, echo=FALSE}
+```{r pca, warning = FALSE, fig.show="hold", out.width="80%", fig.height=5, eval=TRUE, include=TRUE, echo=FALSE}
 # transform data
@@
 ## Colors
-```{r}
+```{r colors, echo=FALSE}
 # colours
@@
 ## Plot Counts
-```{r plot_counts}
+```{r plot_counts, echo=FALSE}
 dds <- dds_list[[1]]
@@
 
-print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))
+# print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))
+
+# single gene
+    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
+
+gcounts <- ggplot(d, aes(x = treatment, y = count, fill=condition, color=condition)) +
+    geom_boxplot(color="black") +
+    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
+  geom_text_repel(aes(label=clientName)) +
+    scale_fill_manual(values=group.colors) +
+    scale_color_manual(values=group.colors) +
+    scale_y_continuous(trans = "log2") +
+    labs(title = paste("ROC40"," in Δ",colData(dds)$experiment[1],sep = ""))
+gcounts %>% print()
 
@@
 
-# ---------------------------
-# COPY & PASTE from pCRY
-## Make results
+#### ---------------------------
+#### COPY & PASTE from pCRY
+#### Make results
 ```{r, eval=FALSE, include=FALSE, echo=FALSE}
@@
 
-## Colours
+### Colours
 ```{r, eval=FALSE}
@@
 
-## PCA (Panel B)
+### PCA (Panel B)
 ```{r, eval=FALSE}
@@
 
-## Counts
+### Counts
 ```{r, eval=FALSE}
@@
 
-## Volcanos
+### Volcanos
 ### dark
@@
 
-## Heatmap
+### Heatmap
 

### 2024-04-26 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
figure fixes

@@
 
-BiocManager::install("ggalt") # , type="source"
+BiocManager::install()
 
@@
 ## -R_libraries
-```{r r_libraries, imnclude=FALSE}
+```{r r_libraries, echo=FALSE}
 library(patchwork)
@@
 colors <- colorRampPalette( rev(brewer.pal(9, "Blues")) )(255)
-anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype","experiment")])
+anno_col <- as.data.frame(colData(vsd)[,c("treatment","genotype")])
 anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"),
-                    genotype = c("green4","orchid"))
+                    genotype = c("green4","tan2"))
 names(anno_colors$treatment) <- levels(anno_col$treatment)
@@
 anno_colors <- list(treatment = c("grey30","skyblue1","lightcoral"), #,"ivory","yellow1","gold"
-                    genotype = c("green4","orchid","tan2"))
+                    genotype = c("green4","tan2","orchid"))
 
@@
 ## Plot Counts
-```{r plot_counts, echo=FALSE}
+```{r plot_counts, echo=FALSE, fig.width=12}
 dds <- dds_list[[1]]

### 2024-05-16 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
update
0_tutorial

@@
 
+## -R functions
+```{r r_functions}
+getresults_SK <- function(contrast,nres){
+r1 <- results(dds, contrast = contrast)
+r1$symbol <- mcols(dds)$symbol
+assign(paste("res",nres,sep="."),r1)
+r1 <- list(r1)
+names(r1) <- nres
+r1
+}
+```
+
 # 1. Load aCRY & pCRY
@@
 # remove outlier
-# outliner <- c("WTdark1")
-# colData(dds)[outliner,]
-# keep <- colnames(dds) %>% .[!str_detect(.,outliner)]
-# dds <- dds[,keep]
+outliner <- c("WTdark3")
+colData(dds)[outliner,]
+keep <- colnames(dds) %>% .[!str_detect(.,outliner)]
+dds <- dds[,keep]
 
@@
 
-res_list_list <- list()
-for (dds in dds_list) {
-
 design(dds)
@@
 
+#    [,1]                       [,2]
+# [1,] "Intercept"                "0" 
+# [2,] "treatment_BL_vs_D"        "0" 
+# [3,] "treatment_R_vs_D"         "0" 
+# [4,] "genotype_pcry_vs_WT"      "0" 
+# [5,] "treatmentBL.genotypepcry" "0" 
+# [6,] "treatmentR.genotypepcry"  "0"
+
+# assign(paste("res",n,sep="_"),list())
+
+res_list <- list()
+for (i in 1:2) {
+dds <- dds_list[i]
+nlist <- names(dds_list[i])
+c <- resultsNames(dds)
+res_l <- list()
+
 # res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))
 
-# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
-res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
-res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")
-
-res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))
-
-res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))
-
+# Belichtungseffect für jeden Genotyp
+nres <- "WT_BL.vs.D"
+contrast <- c("treatment_BL_vs_D")
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"_BL.vs.D",sep="")
+contrast <- list(c("treatment_BL_vs_D",c[5]))
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"_R.vs.D",sep="")
+contrast <- list(c("treatment_R_vs_D",c[6])))
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+# Unterschiede zwischen WT und pCRY unter gleicher Belichtung
+nres <- paste(nlist,"D.vs.WT_D",sep="_")
+contrast <- c(0,0,0,1,0,0)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"BL.vs.WT_BL",sep="_")
+contrast <- c(0,1,0,0,1,0)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"R.vs.WT_R",sep="_")
+contrast <- c(0,0,1,0,0,1)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+# Unterschiede im Belichtungseffekt
+nres <- paste(nlist,"BLvD.vs.WT_BLvD",sep="_")
+contrast <- c(0,0,0,0,1,0)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"RvD.vs.WT_RvD",sep="_")
+contrast <- c(0,0,0,0,0,1)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+## Exotische Kontraste
 # Unterschiede zwischen versch. Belichtungen über beide Genotypen
-# D vs. (BL+R)
-res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))
-
-# D vs. BL
-res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))
-
-# D vs. R
-res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))
+nres <- "BL+R.vs.D"
+contrast <- c(0,0,2/6,2/6,1/3,1/3)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- "BL.vs.D"
+contrast <- c(0,0,1,0,0.5,0)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- "R.vs.D"
+contrast <- c(0,0,0,1,0,0.5)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
 
 # WT vs. pCRY
-res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))
+nres <- paste(nlist,".vs.WT",sep="")
+contrast <- c(0,3,0,0,1,1)
+res_l[[nres]] <- getresults_SK(contrast = contrast, nres = nres)
 

### 2024-05-16 0_Tutorial/m.table.R

@@ -1,0 +1,24 @@
+samplelist <- {}
+mappingrates <- {}
+for (i in list.files(path = quantdir)){
+  print(i)
+  si <- paste("",str_sub(i,11,-7),sep = "")
+  si
+  samplelist <- c(samplelist,si)
+  f <- readLines(paste(quantdir,i,"logs/salmon_quant.log", sep="/"))
+  line <- grep("Mapping rate = ",f,value=TRUE)
+  sl <- str_length(line)
+  sl
+  notime <- substring(line,30,sl)
+  notime
+  manual <- substring(line,sl-7,sl-1)
+  val <- as.numeric(str_extract(notime,"[0-9.]+"))
+  val
+  valr<-round(val, digits=2)
+  print(paste("Mapping rate of ",si," is: ",valr," %"))
+  mappingrates <- c(mappingrates,valr)
+}
+sample.table$mappingrates <- mappingrates
+# Make table
+
+m.table <- data.frame(sample.table,mappingrates)

### 2024-05-16 0_Tutorial/1_data_processing_Tutorial.Rmd
Add files via upload

--- 2024-04-17 1_data_processing/1_data_processing.Rmd
@@
 # 0. Prepare System
+## Download RStudio & R
+https://posit.co/download/rstudio-desktop/
+
+##  -install BiocManager
+```{r}
+if (!require("BiocManager", quietly = TRUE))
+    install.packages("BiocManager")
+BiocManager::install(version = "3.18")
+```
+
+## -update Packages
+```{r}
 BiocManager::install()
+```
+
+## -install new packages
+```{r}
+BiocManager::install("PCAtools")
+
+```
 
@@
 library(AnnotationHub)
-library(DESeq2)
+library(DESeq2) # Main RNA-Seq Package
 library(EnhancedVolcano)
@@
 
+library(readxl)
+
+library(sessioninfo)
+library(data.table)
+library(plyr)
+library(tidyverse)
+library(tximeta)
+library(tximport)
+library(curl)
+
 # library(wget)
@@
 ```{r r_folders}
-ifelse(Sys.info()["sysname"]== "Linux",
-       s <- "/mnt/s",
-       s <- "S:")
-dir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY",sep="/")
-list.files(dir) %>% head()
-gitdir <- paste(dir,"git_Chlamy_RNAseq_aCRY",sep="/")
-list.files(gitdir) %>% head()
-
-fastqdir <- paste(dir,"fastq",sep="/")
-list.files(fastqdir) %>% tail()
-fastqdir2 <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/",sep="/")
-list.files(fastqdir2) %>% head()
-
-indexDir <- paste(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1",sep="/")
-
-fastaPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz")
-head(readLines(fastaPath,n=10))
-
-gtfPath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz")
-readLines(gtfPath,n=10)
-
-quantdir <- paste(dir,"quants",sep="/")
+
+# Server
+server <- "/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq"
+list.files(server)
+
+# for test folder
+dir <- "/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/test" # Working directory
+quantdir <- "/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/test/quants" # quant is the output of the mapping
+figuredir <- "/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/test/figures"
+
+```
+
+## -Connect to Linux machine (Enno))
+```{r, eval=FALSE}
+# run in terminal (mac)
+# run in cmd.exe (Windows)
+
+ssh -Y student@141.20.64.235
+# Password: StudentExpBP
+
+# activate salmon
+conda activate /home/student/conda_salmon
+
+# select server
+cd /ExpBP/Chlamyland/RNA_seq
+cd test
 
@@
 ```{bash linux_folders, eval=FALSE}
-fastqdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/fastq" #aCRY;
-fastqdir2="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/P3044/fastq/" #pCRY;
-
-dir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY";
-
-quantdir="$dir/quants";
-
-indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";
+
+# fastqdir=
+# fastqdir2=
+
+# dir="
+
+# quantdir="$dir/quants";
+
+# indexdir="/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/chlamy_index_v6.1";
 
@@
 
+## -Download via command line in linux
 ```{bash download data, eval=FALSE}
@@
 
-## 1c. Mapping
+## Salmon Mapping
+### Make index
+```{bash make_index}
+
+# Instructions
+# https://salmon.readthedocs.io/en/latest/salmon.html
+
+cd /ExpBP/Chlamyland/RNA_seq
+cd test
+
+salmon index -t gentrome_chlamy.v6.1.fa.gz -i Chlamy_index_Olga  -k 31 -p 10
+
+# --decoys decoys_chlamy_v6.1.txt
+
+```
+
+### 1c. Mapping / Quantifying
 ```{bash mapping, eval=FALSE}
-mamba activate salmon;
-
-fastqfiles=$(ls $fastqdir/*_1*.fq.gz);
-fastqfiles2=$(ls $fastqdir2/*R1*.fastq.gz);
-
-# fastqfiles_all="${fastqfiles} ${fastqfiles2}";
-fastqfiles_all=$fastqfiles;
+
+conda activate /home/student/conda_salmon
+cd /ExpBP/Chlamyland/RNA_seq
+cd test
+
+# test folder
+quantdir="/ExpBP/Chlamyland/RNA_seq/test/quants"
+indexdir="/ExpBP/Chlamyland/RNA_seq/test/Chlamy_index_Olga"
+fasta="/ExpBP/Chlamyland/RNA_seq/test/fastq"
+
+# "Real" data Chlamy_RNASeq_aCRY
+quantdir="/ExpBP/Chlamyland/RNA_seq/Chlamy_RNASeq_aCRY/quants_v2"
+indexdir="/ExpBP/Chlamyland/RNA_seq/test/Chlamy_index_Olga"
+fasta="/ExpBP/Chlamyland/RNA_seq/Chlamy_RNASeq_aCRY/fastq"
+
+# go to fastq folder
+fastqfiles=$(\ls *_1.fq.gz)
 
 echo "${fastqfiles:0:100}";
-echo "${fastqfiles2:0:100}";
-echo "${fastqfiles_all:0:100}";
-
+
+cd $fasta
 # files: Unknown_BU327-002T0001_1.fq.gz;
-for fn in $fastqfiles_all;
+for fn in $fastqfiles;
 do
 bfn=$(echo "`basename ${fn}`");
-samp=${bfn%"_1.fq.gz"};
-R1=$fn;
+dfn=$(echo "`dirname ${fn}`");
+
+samp=$(echo "$bfn" | sed 's/_1.fq.gz//');
+R1=$(echo $fasta"/"$bfn);
 R2=$(echo "$R1" | sed 's/_1/_2/');
 echo "Processing Sample: $samp";
-test -f $R1 && echo "--> File: $R1 exists"
-test -f $R2 && echo "--> File: $R2 exists"
+echo $R1
+echo $R2
+
 outquant=${quantdir}/${samp}_quant
@@
   -2 $R2 \
-  -p 25 --validateMappings --gcBias -o $outquant
+  -p 15 --validateMappings --gcBias -o $outquant
 
 done
+
+# code end
+
+# test if file is there
+if [ -e $R1 ]
+then
+    echo "ok"
+else
+    echo "nok"
+fi
 
@@
 
-quantdir <- paste(dir,"quants",sep="/")
+quantdir # defined above
 
@@
 files <- as.factor(str_remove(f,pattern ="_quant"))
+
 # short form
@@
 sample
+
 # condition
-library(readxl)
-Samplepath <- filePath(path=dir, file="sample_table.xlsx")
+Samplepath <- filePath(path="/Users/olga/Desktop/RNA-Seq Simon/Chlamy/Server/aCRY/", file="sample_table.xlsx")
 Samplefile <- read_xlsx(Samplepath)
@@
 
-m.table <- data.frame(sample.table,mappingrates)
+m.table <- data.frame(sample.table,mappingrates) # here used earlier only for columns [1:2,], otherwise for the whole table use sample.table
 ```
@@
 
-xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename, horiz=T, las=1, xlim=c(0,100))
+xx <- barplot(height=mappingrates, cex.names=0.5, names=samplename[2], horiz=T, las=1, xlim=c(0,100))
 text(x = mappingrates, y = xx, label = mappingrates, pos = 4, cex = 0.5, col = "red")
@@
 ## 2c. Tximeta
+combine counts from different transcripts (splice variants) to gene level
 
 ```{r tximeta}
-library(sessioninfo)
-library(data.table)
-library(plyr)
-library(tidyverse)
-library(tximeta)
-library(tximport)
-library(curl)
-
-getwd()
+
 example.quant <- paste(quantdir,list.files(quantdir)[1],"quant.sf",sep="/")
@@
 head(example.table)
+dim(example.table)
 basename(example.quant)
@@
 files={}
-for (i in list.dirs(path = dirname(dirname(example.quant)), full.names = TRUE, recursive = FALSE)) {
+for (i in list.dirs(path = quantdir, full.names = TRUE, recursive = FALSE)) {
 files <- c(files,paste(i,"/quant.sf",sep=""))
-# print(basename(i))
-# print(head(read.table(files[length(files)], header=T)))
-# print (files[length(files)])
 }
 head(files)
+length(files)
 
@@
 coldata <- data.frame(files, names = run, stringsAsFactors=FALSE)
-coldata <- data.frame(coldata,sample.table)
+coldata <- data.frame(coldata,sample.table[1:2,])
 coldata %>% head() %>% kable()
@@
 ## Chlamy
+
+indexDir="/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/test/Chlamy_index_Olga" # salmon index
+fastaPath="/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.0.hardmasked.fa.gz" # genomic fasta
+gtfPath="/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.gene.gff3.gz"
+
 file.exists(indexDir, fastaPath, gtfPath)
@@
 
+# Column infos = Sample infos -> colData(se)
 colData(se) %>% head() %>% kable()
-# meta infos sind da (genotype, treatment,...)
+
+# Row Infos = Gene Infos -> mcols(se)
+mcols(se) %>% head() %>% kable()
 
@@
 
-rowRanges(se) %>% kable()
-# genome info
-seqinfo(se)
-# ?
+# Gene length
+rowRanges(se) %>% head() %>% kable()
+
 head(assays(se)[["counts"]]) %>% kable()
@@
 
-TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, column = "TXNAME", keytype = "GENEID", multiVals="first"))
+TXNAME <- as.character(mapIds(edb,keys = mcols(gse)$gene_id, keytype = "GENEID", column = "TXNAME", multiVals="first"))
 # select()' returned 1:many mapping between keys and columns = 1 gene has many transcripts...
@@
 getwd()
-# save(gse,file=paste(dir,"tximeta.RDS", sep="/"))
+save(gse,file=paste(dir,"tximeta_test.RDS", sep="/"))
 # gse <- 1
@@
 # more gene info from master_annotation_table
-annopath <- file.path(s,"AG/AG-Scholz-NGS/Daten/Simon/linux-ngs/salmon/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")
+annopath <- file.path("/Volumes/biopro3/ExpBP/Chlamyland/RNA_seq/salmon_index_chlamy/Phytozone_v6.1/CreinhardtiiCC_4532_707_v6.1.master_annotation_table.tsv")
 
 anno <- read.delim2(annopath, header = T, sep = "\t",)
-head(anno)[1:5,1:9] %>% kable()
+head(anno)[1:5,] %>% kable()
 
@@
 
-save(gse,file=paste(dir,"tximeta.RDS",sep="/"))
+save(gse,file=paste(dir,"tximeta_test.RDS",sep="/"))
 ```
@@
 # design <- ~genotype+treatment+genotype:treatment
-
-dds <- DESeqDataSet(gse_acry, design=design)
+colData(gse)
+dds <- DESeqDataSet(gse, design=~ 1)
+dds <- DESeqDataSet(gse, design=design)
+
 colData(dds) %>% head() %>% kable()
@@
 par(mfrow=c(1,2))
-hist(log(counts(dds)), breaks=100, ylim = c(0,11000), xlim = c(0,10))
-hist(counts(dds), breaks=1000000, ylim = c(0,100000), xlim = c(0,10))
+hist(log(counts(dds)), breaks=100)
+hist(counts(dds), breaks=1000000)
 # -> dip at log(counts(dds)=2-3)
@@
 
-colData(dds) %>% head() %>% kable()
 # add metadata

### 2024-05-17 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
Vulcanos

--- 2024-04-26 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
@@
 
+## -R functions
+```{r f_get_results}
+getresults_SK <- function(contrast,nres){
+r1 <- results(dds, contrast = contrast)
+r1$symbol <- mcols(dds)$symbol
+assign(paste("res",nres,sep="."),r1)
+r1 <- list(r1)
+names(r1) <- nres
+r1
+}
+```
+
+```{r f_topgenes}
+topgenes_f <- function(res,p=0.5,bM=20,l2FC=1){
+a <- subset(res, padj < p & baseMean > bM & abs(log2FoldChange) > l2FC)
+a <- a[order(a$baseMean, decreasing = T),]
+  a$rank.bm <- seq(1:length(rownames(a)))
+a <- a[order(a$padj, decreasing = F),]
+  a$rank.padj <- seq(1:length(rownames(a)))
+a <- a[order(abs(a$log2FoldChange), decreasing = T),]
+  a$rank.l2FC <- seq(1:length(rownames(a)))
+a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
+  a <- a[order(a$rank.sum),]
+a
+}
+```
+
+```{r f_vulcano}
+
+Vulcano_SK <- function(res,
+                       n=NULL,
+                       list1=NULL,
+                       list2=NULL,
+                       list1.col=NULL,
+                       list2.col=NULL,
+                       list1.n=NULL,
+                       list2.n=NULL,
+                       xlim=NULL,
+                       ylim=NULL,
+                       ...)
+{
+# Name
+n <- names(res)
+# unlist results
+if(is.list(res)){
+  res <- res[[1]]
+}
+# define list 1
+if(is.null(list1) ) {
+  list1 <- topgenes_f(res,p=0.05,bM = 10,l2FC = 1)
+}
+# define list 2
+if(is.null(list2) ) {
+  list2 <- topgenes_f(res,p=0.005,bM = 100,l2FC = 2)
+}
+# list 1 color
+if(is.null(list1.col) ) {
+  list1.col <- "royalblue1"
+}
+# list 2 color
+if(is.null(list2.col) ) {
+  list2.col <- "royalblue4"
+}
+# name list 1
+if(is.null(list1.n) ) {
+  list1.n <- "list1"
+}
+# name list 2
+if(is.null(list2.n) ) {
+  list2.n <- "list2"
+}
+
+# xlim
+if(is.null(xlim) ) {
+  xlim <- abs(c(min(res$log2FoldChange, na.rm=TRUE),
+              max(res$log2FoldChange, na.rm=TRUE))) %>% max()
+}
+
+# ylim
+if(is.null(ylim) ) {
+  ylim <- max(-log10(res$padj), na.rm = TRUE) + 5
+}
+
+# colors
+pcol <- "black"
+lcol <- colours[1+1]
+
+# add symbol
+res$symbol <- mcols(dds_list[[i]])$symbol
+
+# define list groups
+res$isl1 <- rownames(res) %in% rownames(list1)
+res$isl2 <- rownames(res) %in% rownames(list2)
+
+# order results (list 1 on top)
+res <- res[order(abs(res$log2FoldChange), decreasing = T),]
+res <- res[order(res$isl1, decreasing = F),]
+res <- res[order(res$isl2, decreasing = F),]
+
+# change outlier shape
+shape <- 1
+if(!is.null(xlim)){
+shape <- ifelse(abs(res$log2FoldChange) > xlim, 6,16)
+}
+if(!is.null(ylim)){
+shape[res$padj < 10^-ylim] <- 6
+}
+# name shape
+names(shape)[shape == 6] <- 'outlier'
+names(shape)[shape == 16] <- 'in range'
+# move outlier to border
+res$log2FoldChange[res$log2FoldChange > xlim] <- xlim
+res$log2FoldChange[res$log2FoldChange < -xlim] <- -xlim
+res$padj[res$padj < 10^-ylim] <- 10^-ylim
+
+# name colors
+keyvals <- ifelse(
+    res$isl2 == TRUE, list2.col,
+    ifelse(
+    res$isl1 == TRUE, list1.col,
+    'grey') )
+
+keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == list1.col] <- list1.n
+  names(keyvals)[keyvals == list2.col] <- list2.n
+  names(keyvals)[keyvals == 'grey'] <- 'other'
+
+ev_f <- EnhancedVolcano(res,
+    x = 'log2FoldChange',
+    y = 'padj',
+    lab = res$symbol,
+    selectLab  = res$symbol[which(names(keyvals) == list2.n)],
+    labSize = 3,
+    drawConnectors = TRUE,
+    boxedLabels = FALSE,
+    widthConnectors = 0.5,
+    colConnectors = lcol,
+    pointSize = 4.0,
+    max.overlaps = 100,
+    colCustom = keyvals,
+    xlim = c(-xlim,xlim),
+    ylim = c(0, ylim),
+    ylab = "Padj (-log10)",
+    title = n,
+    subtitle = paste("list1:",nrow(list1),", list2: ",nrow(list2)),
+    # sub = "SVF",
+    # pCutoff = 10^(-60),
+    # FCcutoff = 1,
+    shapeCustom =shape,
+    # pointSize = c(ifelse(rownames(res_WT_D_vs.WT_BL) %in% rownames(top_WT_BL_vs.pcry_BL), 8, 1)),
+    legendLabels=c('Not sig.','|L2F| > 2','p-adj < 0.05',
+                    'p-adj & L2F'),
+    legendPosition = 'bottom',
+    col=c('grey', pcol, pcol, pcol)
+    )
+ev_f
+
+}
+
+```
+
 # 1. Load aCRY & pCRY
@@
 
+#      [,1]                         [,2]
+# [1,] "Intercept"                  "0" 
+# [2,] "genotype_pcry_vs_WT"        "0" 
+# [3,] "treatment_blue_vs_dark"     "0" 
+# [4,] "treatment_red_vs_dark"      "0" 
+# [5,] "genotypepcry.treatmentblue" "0" 
+# [6,] "genotypepcry.treatmentred"  "0"
+
+# assign(paste("res",n,sep="_"),list())
+
+res_list <- list()
+for (i in 1:2) {
+dds <- dds_list[[i]]
+
+nlist <- names(dds_list[i])
+c <- resultsNames(dds)
+print(nlist)
+res_l <- list()
+
 # res_WT_D_vs.pcry_D <- results(dds, contrast = c("condition","pcry_D","WT_D"))
 
-# Unterschiede zwischen WT und pCRY unter versch. Belichtungen
-res1 <- res_WT_D_vs.pcry_D <- results(dds, contrast=c(0,0,0,1,0,0))
-res1 <- res_WT_D_vs.pcry_D <- results(dds,name = "genotype_pcry_vs_WT")
-
-res2 <- res_WT_BL_vs.pcry_BL <- results(dds, contrast=c(0,1,0,0,1,0))
-
-res3 <- res_WT_R_vs.pcry_R <- results(dds, contrast=c(0,0,1,0,0,1))
-
+# Belichtungseffect für jeden Genotyp
+nres <- "WT_BL.vs.D"
+contrast <- list(c("treatment_blue_vs_dark"))
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- "WT_R.vs.D"
+contrast <- list(c("treatment_red_vs_dark"))
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"_BL.vs.D",sep="")
+contrast <- list(c("treatment_blue_vs_dark",c[5])) # [5,] "genotypepcry.treatmentblue"
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"_R.vs.D",sep="")
+contrast <- list(c("treatment_red_vs_dark",c[6])) # [6,] "genotypepcry.treatmentred"  
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+# Unterschiede zwischen WT und ko unter gleicher Belichtung
+nres <- paste(nlist,"D.vs.WT_D",sep="_")
+contrast <- list(c[2])                # [2,] "genotype_pcry_vs_WT"
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"BL.vs.WT_BL",sep="_")
+contrast <- list(c[2],c[5])           # [5,] "genotypepcry.treatmentblue"
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"R.vs.WT_R",sep="_")
+contrast <- list(c[2],c[6])           # [6,] "genotypepcry.treatmentred"  
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+# Unterschiede im Belichtungseffekt
+nres <- paste(nlist,"BLvD.vs.WT_BLvD",sep="_")
+contrast <- c(0,0,0,0,1,0)
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- paste(nlist,"RvD.vs.WT_RvD",sep="_")
+contrast <- c(0,0,0,0,0,1)
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+## Exotische Kontraste
 # Unterschiede zwischen versch. Belichtungen über beide Genotypen
-# D vs. (BL+R)
-res4 <- res_D_vs._BL_R <-results(dds, contrast=c(0,0,2/6,2/6,1/3,1/3))
-
-# D vs. BL
-res5 <- res_BL_vs._D <- results(dds, contrast=c(0,0,1,0,0.5,0))
-
-# D vs. R
-res6 <- res_D_vs._R <- results(dds, contrast=c(0,0,0,1,0,0.5))
-
-# WT vs. pCRY
-res7 <- res_WT_vs._pcry <- results(dds, contrast=c(0,3,0,0,1,1))
-
-## Make top lists:
-
-topgenes <- function(res){
-a <- subset(res, padj < 0.05 & baseMean > 20 & (log2FoldChange < -0.5 | log2FoldChange > 0.5))
-a$id.symbol <- anno[rownames(a),"id.symbol"]
-a$symbol <- anno[rownames(a),"geneSymbol"]
-a <- a[order(a$baseMean, decreasing = T),]
-  a$rank.bm <- seq(1:length(rownames(a)))
-a <- a[order(a$padj, decreasing = F),]
-  a$rank.padj <- seq(1:length(rownames(a)))
-a <- a[order(abs(a$log2FoldChange), decreasing = T),]
-  a$rank.l2FC <- seq(1:length(rownames(a)))
-a$rank.sum <- a$rank.l2FC+a$rank.bm+a$rank.padj
-  a <- a[order(a$rank.sum),]
-print(a)
-}
-
-top.res1 <- topgenes(res=res1)
-summary(res1)
-top.res2 <- topgenes(res=res2)
-summary(res2)
-top.res3 <- topgenes(res=res3)
-summary(res3)
-top.res4 <- topgenes(res=res4)
-summary(res4)
-top.res5 <- topgenes(res=res5)
-summary(res5)
-top.res6 <- topgenes(res=res6)
-summary(res6)
-top.res7 <- topgenes(res=res7)
-summary(res7)
-
-top.all <- rbind(top.res1,
-         top.res2,
-         top.res3,
-         top.res4,
-         top.res5,
-         top.res6,
-         top.res7)
-top.all <- top.all[order(top.all$rank.sum),]
-head(top.all, n=185) %>% rownames() %>% unique() %>% length()
-top.all <- top.all[unique(rownames(top.all)),]
-
-# top.all <- head(top.all, n=100)
-
-## Shrinkage
-res1_shrink <- lfcShrink(dds, res=res1, type="ashr")
-res2_shrink <- lfcShrink(dds, res=res2, type="ashr")
-res3_shrink <- lfcShrink(dds, res=res3, type="ashr")
-res4_shrink <- lfcShrink(dds, res=res4, type="ashr")
-res5_shrink <- lfcShrink(dds, res=res5, type="ashr")
-res6_shrink <- lfcShrink(dds, res=res6, type="ashr")
-res7_shrink <- lfcShrink(dds, res=res7, type="ashr")
+nres <- "BL+R.vs.D"
+contrast <- c(0,0,2/6,2/6,1/3,1/3)
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- "BL.vs.D"
+contrast <- c(0,0,1,0,0.5,0)
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+nres <- "R.vs.D"
+contrast <- c(0,0,0,1,0,0.5)
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+# WT vs. ko
+nres <- paste(nlist,".vs.WT",sep="")
+contrast <- c(0,1,0,0,1/3,1/3)
+res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+print(length(res_l))
+print(length(res_list))
+print(names(res_list))
+
+res_list[[nlist]] <- res_l
+print(length(res_list))
+print(names(res_list))
+
+    }
+
+## Make deg & top lists:
+deg_list <- list()
+top_list <- list()
+
+for (i in 1:2) {
+ko <- names(res_list[i])
+
+deg_genes_l <- list()
+deg_genes_l <- lapply(res_list[[i]],topgenes_f) %>%  lapply(.,rownames) 
+names(deg_genes_l) <- paste("deg",names(res_list[[i]]),sep="_")
+deg_list[[ko]] <- deg_genes_l
+
+top_genes_l <- list()
+top_genes_l <- lapply(res_list[[i]],topgenes_f,p=0.01, bM=100, l2FC=2) %>%  lapply(.,rownames) 
+names(top_genes_l) <- paste("top",names(res_list[[i]]),sep="_")
+top_list[[ko]] <- top_genes_l
+
+}
+
+top_genes_l <- lapply(res_list,topgenes_f,p=0.01, bM=100, l2FC=2) %>%  lapply(.,rownames)
+
+## Make shrinkage
+res_ashr_list <- list()
+for (i in 1:2) {
+ko <- names(res_list[i])
+
+res_ashr_l <- list()
+l <- length(res_list[[i]])
+for (j in 1:l){
+res_ashr_l[[names(res_list[[i]])[j]]] <- lfcShrink(res=res_list[[i]][[j]], dds=dds_list[[i]], type="ashr")
+}
+res_ashr_list[[ko]] <- res_ashr_l
+}
 
@@
 ### dark
-```{r, eval=FALSE}
+```{r, eval=TRUE}
+
+i <- 2
+
+Vulcano_SK(res_ashr_list[[i]]["pcry_BL.vs.WT_BL"],xlim=5,ylim=50)
 

### 2024-06-14 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
advanced volcanos

@@
 
+library(VennDiagram)
+
 # library(wget)
@@
 
+pub_figures <- paste(gitdir,"pub_figures",sep="/")
+
 # Macbook
@@
 ```{r f_topgenes}
-topgenes_f <- function(res,p=0.5,bM=20,l2FC=1){
+
+topgenes_f <- function(res,p=0.005,bM=10,l2FC=1){
 a <- subset(res, padj < p & baseMean > bM & abs(log2FoldChange) > l2FC)
+if(nrow(a)>0) {
 a <- a[order(a$baseMean, decreasing = T),]
@@
   a <- a[order(a$rank.sum),]
+  }
 a
@@
 ```{r f_vulcano}
+
+# res <- res_ashr_list[[i]][6]
 
@@
                        list2.col=NULL,
+                       ol.col=NULL,
                        list1.n=NULL,
@@
                        ylim=NULL,
+
                        ...)
@@
 # Name
+if(is.null(n))  {
+  if(is.list(res)){
       n <- names(res)
+  } else {
+    n <- "results"
+  }
+  }
+
 # unlist results
@@
 }
+
 # define list 1
 if(is.null(list1) ) {
-  list1 <- topgenes_f(res,p=0.05,bM = 10,l2FC = 1)
+  list1 <- topgenes_f(res=res,p=0.05,bM = 10,l2FC = 1) %>% rownames()
 }
@@
 if(is.null(list2) ) {
-  list2 <- topgenes_f(res,p=0.005,bM = 100,l2FC = 2)
+  list2 <- topgenes_f(res=res, p=0.01, l2FC = 2, bM=10) %>% rownames()
 }
@@
 }
+# list ol color
+if(is.null(ol.col) ) {
+  ol.col <- "yellow"
+}
+
+# calculate overlap
+ol <- calculate.overlap(list(list1,list2))
+
 # name list 1
@@
 # add symbol
-res$symbol <- mcols(dds_list[[i]])$symbol
+res$symbol <- mcols(dds_list[[i]])$geneSymbol
 
 # define list groups
-res$isl1 <- rownames(res) %in% rownames(list1)
-res$isl2 <- rownames(res) %in% rownames(list2)
-
-# order results (list 1 on top)
+res$isl1 <- rownames(res) %in% list1
+res$isl2 <- rownames(res) %in% list2
+res$isol <- rownames(res) %in% ol$a3
+
+# order results (list ol on top)
 res <- res[order(abs(res$log2FoldChange), decreasing = T),]
@@
 res <- res[order(res$isl2, decreasing = F),]
+res <- res[order(res$isol, decreasing = F),]
 
@@
 keyvals <- ifelse(
+  res$isol == TRUE, ol.col,
+    ifelse(
   res$isl2 == TRUE, list2.col,
@@
     res$isl1 == TRUE, list1.col,
-    'grey') )
+    'grey') ) )
 
 keyvals[is.na(keyvals)] <- 'grey'
+  names(keyvals)[keyvals == ol.col] <- "overlap"
   names(keyvals)[keyvals == list1.col] <- list1.n
@@
     lab = res$symbol,
-    selectLab  = res$symbol[which(names(keyvals) == list2.n)],
+    selectLab  = res$symbol[which(names(keyvals) == list2.n | names(keyvals) ==list1.n | names(keyvals) =="overlap")],
     labSize = 3,
@@
     title = n,
-    subtitle = paste("list1:",nrow(list1),", list2: ",nrow(list2)),
+    subtitle = paste("list1:",length(list1),", list2: ",length(list2)),
     # sub = "SVF",
@@
     legendPosition = 'bottom',
-    col=c('grey', pcol, pcol, pcol)
+    col=c('grey', pcol, pcol, pcol),
+    legendLabSize = 8,
+    legendIconSize = 2.0,
+    axisLabSize = 8,
+    titleLabSize = 8,
+    subtitleLabSize = 8,
+    captionLabSize = 8
     )
@@
     d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
-
-gcounts <- ggplot(d, aes(x = treatment, y = count, fill=condition, color=condition)) +
+    colnames(d)[1] <- "counts"
+
+gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
     geom_boxplot(color="black") +
     geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
-  geom_text_repel(aes(label=clientName)) +
+  # geom_text_repel(aes(label=clientName)) +
     scale_fill_manual(values=group.colors) +
@@
     scale_y_continuous(trans = "log2") +
-    labs(title = paste("ROC40"," in Δ",colData(dds)$experiment[1],sep = ""))
+    labs(title = paste("ROC40"," in ",colData(dds)$experiment[1],sep = "")) +
+ theme(axis.text.x = element_text(size = 10, face = "bold"),
+        axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
+        axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))
+
 gcounts %>% print()
@@
 
-# ggsave(paste(figures,"2023_12_counts_phots_aio_log2.pdf",sep="/"), plot = gcounts,
-#   width = 12,
-#   height = 8)
+# ggsave(paste(pub_figures,"ROC40.pdf",sep="/"), plot = gcounts,
+#   width = 5,
+#   height = 4)
 
@@
 ## Make results
-```{r, eval=FALSE, include=FALSE, echo=FALSE}
+```{r, eval=TRUE, include=FALSE, echo=FALSE}
 
@@
 
-top_genes_l <- lapply(res_list,topgenes_f,p=0.01, bM=100, l2FC=2) %>%  lapply(.,rownames)
-
 ## Make shrinkage
@@
 ## Volcanos
+
+### Volcanos 2.0
+#### top genes
+```{r top_volcano2}
+
+i <- 1
+ev1_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
+                    xlim=5,ylim=10,
+                    list2.col = "black",list1.col = "grey35")
+ev1_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
+                    xlim=5,ylim=10,
+                    list2.col = "royalblue4",list1.col = "royalblue1")
+ev1_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
+                    xlim=5,ylim=10,
+                    list2.col = "firebrick4",list1.col = "firebrick1")
+
+i <- 2
+ev2_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
+                    xlim=5,ylim=50,
+                    list2.col = "black",list1.col = "grey35")
+ev2_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
+                    xlim=5,ylim=50,
+                    list2.col = "royalblue4",list1.col = "royalblue1")
+ev2_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
+                    xlim=5,ylim=50,
+                    list2.col = "firebrick4",list1.col = "firebrick1")
+
+ev1_d+ev1_b+ev1_r  +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+ev2_d+ev2_b+ev2_r  +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+# ga <- ggarrange(evb,
+#                 evd+rremove("ylab"),
+#                 evr+rremove("ylab"),
+#                 ncol = 3, nrow = 2,common.legend = TRUE, legend = "bottom",
+#                 widths = c(1.05,1,1))
+# 
+# 
+# ggsave(ga, file="figures/Vulcano_all.pdf",
+#        width = 12,
+#        height = 8)
+
+```
+
+#### interaction in D-D, Bl-Bl,R-R
+```{r vsvs_volcano2}
+
+i <- 1
+ev1_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=10)
+ev1_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=10)
+ev1_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=10)
+
+i <- 2
+ev2_d <- Vulcano_SK(res=res_ashr_list[[i]][5],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=50)
+ev2_b <- Vulcano_SK(res=res_ashr_list[[i]][6],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=50)
+ev2_r <- Vulcano_SK(res=res_ashr_list[[i]][7],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=50)
+
+ev1_d+ev1_b+ev1_r  +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+ev2_d+ev2_b+ev2_r  +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+```
+
+#### interaction in BL-D, R-D
+```{r vsvs2_volcano2}
+
+deg_list[[1]][8:9] %>% names()
+
+i <- 1
+ev1.1 <- Vulcano_SK(res=res_ashr_list[[i]][1],
+                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5, ylim=50)
+ev1.2 <- Vulcano_SK(res=res_ashr_list[[i]][2],
+                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=50)
+ev1.3 <- Vulcano_SK(res=res_ashr_list[[i]][3],
+                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=50)
+ev1.4 <- Vulcano_SK(res=res_ashr_list[[i]][4],
+                    list1=top_list[[i]][[8]], list2=top_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=50)
+
+i <- 2
+ev2.1 <- Vulcano_SK(res=res_ashr_list[[i]][1],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=100)
+ev2.2 <- Vulcano_SK(res=res_ashr_list[[i]][2],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=100)
+ev2.3 <- Vulcano_SK(res=res_ashr_list[[i]][3],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=100)
+ev2.4 <- Vulcano_SK(res=res_ashr_list[[i]][4],
+                    list1=deg_list[[i]][[8]], list2=deg_list[[i]][[9]],
+                    list1.n="blue", list2.n="red",
+                    list1.col = "royalblue1",list2.col = "firebrick1",ol.col="orchid",
+                    xlim=5,ylim=100)
+
+ev1.1 + ev1.2 + ev1.3 + ev1.4 +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+ev2.1 + ev2.2 + ev2.3 + ev2.4 +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+ev1.1 + ev1.2 + ev2.1 + ev2.2 +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'bottom', axis.title=element_text(size=8))
+
+```
+
 ### dark
-```{r, eval=TRUE}
-
-i <- 2
-
-Vulcano_SK(res_ashr_list[[i]]["pcry_BL.vs.WT_BL"],xlim=5,ylim=50)
+```{r, eval=FALSE}
 
 # dark
-res <- res1_shrink
+res <- res_ashr_list[[i]][[5]]
 n <- 'WT vs. pCRY in dark'
 nf <- 'figures/Volcano_dark.png'
-l <- nrow(top.res1)
+l <- nrow(top_list[[i]][['pCRY_D.vs.WT_D']])
 pcol <- "black"
@@
 
-rownames(res) <- mcols(dds)$id.symbol
+rownames(res) <- mcols(dds_list[[i]])$id.symbol
 res <- res[!is.na(res$padj),]

### 2024-10-17 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
restart
manuscript figures, new structure, PCA, Counts, Voclanos

@@
 editor_options: 
-  chunk_output_type: console
+  chunk_output_type: inline
 knit: (function(input_file, encoding) {
@@
 ```{r r_libraries, echo=FALSE}
+library(ggExtra)
 library(patchwork)
@@
 
+### -change wrong aCRY sample
+```{r correct_acry, eval=TRUE}
+
+gse_acry
+mcols(gse_acry)
+colData(gse_acry)
+
+colData(gse_acry)[c("aCRYdark1","WTdark1"),]
+
+colData(gse_acry)$Samplename_new <- colData(gse_acry)$clientName
+colData(gse_acry)$Samplename_new[colData(gse_acry)$clientName == "aCRYdark1"] <- "WTdark1"
+colData(gse_acry)$group[colData(gse_acry)$clientName == "aCRYdark1"] <- "WTdark"
+colData(gse_acry)$genotype[colData(gse_acry)$clientName == "aCRYdark1"] <- "WT"
+colData(gse_acry)$treatment[colData(gse_acry)$clientName == "aCRYdark1"] <- "dark"
+colData(gse_acry)$condition[colData(gse_acry)$clientName == "aCRYdark1"] <- "WT_dark"
+
+colData(gse_acry)$Samplename_new[colData(gse_acry)$clientName == "WTdark1"] <- "aCRYdark1"
+colData(gse_acry)$group[colData(gse_acry)$clientName == "WTdark1"] <- "aCRYdark"
+colData(gse_acry)$genotype[colData(gse_acry)$clientName == "WTdark1"] <- "acry"
+colData(gse_acry)$treatment[colData(gse_acry)$clientName == "WTdark1"] <- "dark"
+colData(gse_acry)$condition[colData(gse_acry)$clientName == "WTdark1"] <- "acry_dark"
+
+colData(gse_acry)[c("aCRYdark1","WTdark1"),]
+
+rownames(colData(gse_acry)) <- colData(gse_acry)$Samplename_new
+
+```
+
 ## Run Deseq2
@@
 ## Make results
-```{r, eval=TRUE, include=FALSE, echo=FALSE}
+```{r results, eval=TRUE, include=FALSE, echo=FALSE}
 

### 2024-10-17 pub_figures/manuscript figures.Rmd
manuscript figures, new structure, PCA, Counts, Voclanos

---
title: "figures_manuscript"
output:
  github_document:
    html_preview: false
    toc: true
    toc_depth: 2
knit: (function(input_file, encoding) {
    rmarkdown::render(input_file,output_file= 'README.md')
        })
date: "2024-10-17"
editor_options: 
  chunk_output_type: inline
---

```{r knit, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      error=FALSE,
                      warning=FALSE,
                      message=FALSE,
                      dpi=300)

setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

```

BiocManager::install()

# run prepare_data scripts
```{r run_start_up, include=FALSE, results='hide'}
library(knitr)
source(purl("/mnt/s/AG/AG-Scholz-NGS/Daten/Simon/Chlamy_RNASeq_aCRY/git_Chlamy_RNAseq_aCRY/2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd", output = tempfile()), local = knitr::knit_global())

futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")

```

# Fig 1: PCA
```{r 1_pca, fig.show='hold', out.width="50%", dpi=300, results='hide'}

fig <- "Fig1"

# group
for (dds in dds_list) {
vsd <- vst(dds, blind=FALSE)
# colData(dds)
### PCA with top 500 genes with highest row variance 
pcaData <- plotPCA(vsd, intgroup=c("treatment", "genotype","condition"), returnData=TRUE,)
percentVar <- round(100 * attr(pcaData, "percentVar"))

g1 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(size=4, stroke = 1) +
  scale_shape_manual(values = c(16,21)) + 
  labs(title = "PCA plot") +
  ylab(paste0("PC1: ",percentVar[1],"% variance")) +
  xlab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment) +
  # coord_fixed() +
  scale_x_reverse() +
  theme_bw() +
  removeGrid(x=T, y=T)
}

# pCRY
dds <- dds_list[["pcry"]]
vsd <- vst(dds, blind=FALSE)

pcaData <- plotPCA(vsd, intgroup=c("treatment", "genotype","condition"), returnData=TRUE,)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g1 <- ggplot(pcaData, aes(PC2, PC1, color=treatment, shape=genotype)) +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(size=4, stroke = 1) +
  scale_shape_manual(values = c(16,21)) + 
  labs(title = "PCA plot: pCRY") +
  ylab(paste0("PC1: ",percentVar[1],"% variance")) +
  xlab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment) +
  # coord_fixed() +
  scale_x_reverse() +
  scale_y_reverse() +
  theme_bw() +
  removeGrid(x=T, y=T)
g1

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_PCA.pdf",sep=""), plot = g1,
width = 6,
height = 6)

# aCRY
dds <- dds_list[["acry"]]
vsd <- vst(dds, blind=FALSE)
pcaData <- plotPCA(vsd, intgroup=c("treatment", "genotype","condition"), returnData=TRUE,)
percentVar <- round(100 * attr(pcaData, "percentVar"))
g2 <- ggplot(pcaData, aes(PC1, PC2, color=treatment, shape=genotype)) +
  geom_hline(yintercept = 0, linewidth = 0.1) + 
  geom_vline(xintercept = 0, linewidth = 0.1) +
  geom_point(size=4, stroke = 1) +
  scale_shape_manual(values = c(16,21)) + 
  labs(title = "PCA plot: aCRY") +
  ylab(paste0("PC1: ",percentVar[1],"% variance")) +
  xlab(paste0("PC2: ",percentVar[2],"% variance")) + 
  scale_color_manual(values = anno_colors$treatment) +
  # coord_fixed() +
  scale_x_reverse() +
  theme_bw() +
  removeGrid(x=T, y=T)
g2

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_PCA.pdf",sep=""), plot = g2,
width = 6,
height = 6)

```

# Fig 2: Counts
```{r Counts, echo=FALSE, fig.width=10}

fig <- "Fig2"

dds <- dds_list[[1]]
for (dds in dds_list){
# TF genes

goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
anno_tf <- subset(anno,geneSymbol %in% goi_tf | gene_id %in% c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"))
anno_tf[,c("gene_id", "geneSymbol","id.symbol")]
summary(anno_tf$gene_id %in% rownames(dds))
anno_tf[!anno_tf$gene_id %in% rownames(dds),]
anno_tf <- anno_tf[anno_tf$gene_id %in% rownames(dds),]

# Combined counts table TFs

goi <- anno_tf
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
levels(all_counts$Gene)

max_val <- 1.0*max(all_counts$count)

# Plot
gcounts_tf1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  labs(title = "ROC genes") + 
  theme_bw() +
  removeGrid(x=T, y=T) +
    geom_vline(xintercept=seq(1,length(levels(all_counts$Gene))-1,1)+.5,color="grey") +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_tf1 %>% print()

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_ROCs.pdf",sep=""), plot = gcounts_tf1,
width = 12,
height = 8)

## Phot genes all-in-one
goi_phot <- c("CHR1", "CHR2", "PCRY1", "ACRY1", "DCRY1", "PHOT1", "UVR8", "HKR1") #"HKR2" "DCRY2",

anno_phot <- subset(anno,geneSymbol %in% goi_phot, drop = FALSE)
rownames(anno_phot) <- anno_phot$geneSymbol
anno_phot <- anno_phot[goi_phot,]

# Combined counts table TFs

goi <- anno_phot
l <- nrow(goi)
all_counts <- {}
for (i in 1:l){
  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
  d$Gene <- rep(goi[i,"geneSymbol"],length(rownames(d)))
  d$sample <- rownames(d)
  rownames(d) <- {}
  all_counts <- bind_rows(all_counts,d)
  }

all_counts$Gene
levels(all_counts$condition)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene)
levels(all_counts$Gene)

all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
levels(all_counts$Gene)

# Plot
max_val <- 1.0*max(all_counts$count)
gcounts_phot1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
  geom_boxplot(fatten = 1) +
  scale_fill_manual(values = group.colors) +
  labs(title = "Photoreceptor genes") + 
  theme_bw() +
  removeGrid(x=T, y=T) +
  geom_vline(xintercept=seq(1,length(levels(all_counts$Gene))-1,1)+.5,color="grey") +
  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
gcounts_phot1 %>% print()

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_Phots.pdf",sep=""), plot = gcounts_phot1,
width = 12,
height = 8)

# print(gcounts_tf1 / gcounts_phot1 & plot_annotation(title = colData(dds)$experiment[1]))

# single gene
    d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
    colnames(d)[1] <- "counts"

gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
    geom_boxplot(color="black") +
    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
  # geom_text_repel(aes(label=clientName)) +
    scale_fill_manual(values=group.colors) +
    scale_color_manual(values=group.colors) +
    scale_y_continuous(trans = "log2") +
    theme_bw() +
  removeGrid(x=T, y=T) +
    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +

    labs(title = paste("ROC40"," in ",colData(dds)$experiment[1],sep = "")) +
 theme(axis.text.x = element_text(size = 10, face = "bold"),
        axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
        axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))

gcounts %>% print()

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_ROC40.pdf",sep=""), plot = gcounts,
width = 6,
height = 5)
}

# ggsave(paste(pub_figures,"ROC40.pdf",sep="/"), plot = gcounts,
#   width = 5,
#   height = 4)

```

# Fig 3: Volcanos
```{r volcanos, fig.height=8 }

fig <- "Fig3"

dds <- dds_list[["pcry"]]

res_ashr_list %>% names()
res_ashr_list[[1]] %>% names()

res <- res_ashr_list$pcry$pcry_R.vs.WT_R
res_n <- res_list$pcry$pcry_R.vs.WT_R

# of shrinked results
total <- subset(res, padj< 0.05) %>% nrow()
up <- subset(res, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# of "true" results
total <- subset(res_n, padj< 0.05) %>% nrow()
up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# points outside the grid
subset(res, padj < 10^-50 | log2FoldChange > 6 | log2FoldChange < -6)
#                 baseMean log2FoldChange     lfcSE      pvalue        padj
#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
# Cre16.g681750 2907.4406       -3.51641   0.22216 2.99151e-57 4.79270e-53
# Cre17.g802135   46.8587       36.40049   7.35075 1.00156e-08 7.64096e-06

res["Cre16.g681750","padj"] <- 10^-50
res["Cre17.g802135","log2FoldChange"] <- 6

mcols(dds_list[["pcry"]]) %>% nrow()
res %>% nrow()

volcano_red <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[5]),
    title = "red-light, pCRY vs. WT",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 40,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)
# volcano_red

# blue

res_ashr_list %>% names()
res_ashr_list[[1]] %>% names()

res <- res_ashr_list$pcry$pcry_BL.vs.WT_BL
res_n <- res_list$pcry$pcry_BL.vs.WT_BL

# of "true" results
total <- subset(res_n, padj< 0.05) %>% nrow()
up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# points outside the grid

pmax <- 10^-50
l2FCmax <- 6
subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
# res[res$padj < pmax,]$padj <- pmax

subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

mcols(dds_list[["pcry"]]) %>% nrow()
res %>% nrow()

volcano_blue <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[3]),
    title = "blue-light, pCRY vs. WT",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 40,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)
# volcano_blue

# dark

res_ashr_list %>% names()
res_ashr_list[[2]] %>% names()

res <- res_ashr_list$pcry$pcry_D.vs.WT_D
res_n <- res_list$pcry$pcry_D.vs.WT_D

# of "true" results
total <- subset(res_n, padj< 0.05) %>% nrow()
up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()

# points outside the grid

pmax <- 10^-50
l2FCmax <- 6
subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
res[subset(res, padj < pmax) %>% rownames(),]$padj <- pmax

subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)

mcols(dds_list[["pcry"]]) %>% nrow()
res %>% nrow()

volcano_dark <- EnhancedVolcano(res,
    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
    x = 'log2FoldChange',
    y = 'padj',
    col=c("grey","grey","grey",group.colors[1]),
    title = "dark, pCRY vs. WT",
    titleLabSize = 12,
    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
#    subtitle = {},
    subtitleLabSize = 10,
    caption = NULL,
    xlim = c(-7,7),
    ylim = c(0,50),
    pCutoff = 0.05,
    FCcutoff = 1,
    maxoverlapsConnectors = 40,
    drawConnectors = TRUE,
    widthConnectors = 0.5,
    colConnectors = "grey70",
    legendLabels=c('ns','ns','ns',
      'padj < 0.05 & Log2FC > 1'),
    labSize = 4,
    axisLabSize = 12,
    legendLabSize = 12,
    legendIconSize = 4,
    gridlines.major = FALSE,
    gridlines.minor = FALSE,
    pointSize = 3
)
# volcano_dark

volcanos_all <- volcano_dark + volcano_blue + volcano_red +
  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
  theme(legend.position = 'none', axis.title=element_text(size=12))
volcanos_all

ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
width = 8,
height = 6)

```

# Fig 4: Overlaps
```{r overlap, eval=FALSE}
fig <- "Fig4"

res_list[["pcry"]] %>% names()
#  "pcry_D.vs.WT_D"       "pcry_BL.vs.WT_BL"     "pcry_R.vs.WT_R"      
DEGs <- list(dark = res_list$pcry$pcry_D.vs.WT_D %>% subset(padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )),
             blue = res_list$pcry$pcry_BL.vs.WT_BL %>% subset(padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )),
             red =res_list$pcry$pcry_R.vs.WT_R %>% subset(padj < 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )))
DEGs <- lapply(DEGs,data.frame)

venn.ol <- calculate.overlap(DEGs)
# names_res_MO_XX_padj05_abs05 %>% unique()

summary(names_res_MO_XX_padj05_abs05)

venn.ol %>% lapply(length)

input_list <- list(XX=names_res_MO_XX_padj05_abs05,XY=names_res_MO_XY_padj05_abs05)

plt <- venn.diagram(
    x = input_list,
    inverted=TRUE,
    # total.population = TRUE,
    filename = NULL,
    fontfamily ="Arial",
    lwd = 2,
    lty = 'blank',
    fill = cols2[c(1,2)],
    category.names = paste0(names(input_list),"\n(",lapply(input_list,length),")"),
    cat.col=cols2d[c(1,2)],
    cat.fontface = "bold",
    cat.fontfamily = "arial",
    cat.pos = c(+30,-30),
    cat.dist = c(0.12, 0.12),
    disable.logging = TRUE
)

wrap_elements(plt) + plot_annotation(caption = paste0("Total differentially expressed genes: ",deg_table %>% nrow()))

```

### 2024-10-17 pub_figures/manuscript figures.Rmd
volcanos

@@
 # Fig 4: Overlaps
+## Fig 4a: Venns
 ```{r overlap, eval=FALSE}
@@
 DEGs <- lapply(DEGs,data.frame)
+DEGs_genes <- lapply(DEGs,rownames)
 
-venn.ol <- calculate.overlap(DEGs)
-# names_res_MO_XX_padj05_abs05 %>% unique()
-
-summary(names_res_MO_XX_padj05_abs05)
-
+venn.ol <- calculate.overlap(DEGs_genes)
 venn.ol %>% lapply(length)
 
-input_list <- list(XX=names_res_MO_XX_padj05_abs05,XY=names_res_MO_XY_padj05_abs05)
+input_list <- DEGs_genes
 
@@
     lty = 'blank',
-    fill = cols2[c(1,2)],
+    fill = group.colors[c(2,4,6)],
     category.names = paste0(names(input_list),"\n(",lapply(input_list,length),")"),
-    cat.col=cols2d[c(1,2)],
+#    cat.col=cols2d[c(1,2)],
     cat.fontface = "bold",
     cat.fontfamily = "arial",
-    cat.pos = c(+30,-30),
-    cat.dist = c(0.12, 0.12),
+#    cat.pos = c(+30,-30),
+#    cat.dist = c(0.12, 0.12),
     disable.logging = TRUE
@@
 
-wrap_elements(plt) + plot_annotation(caption = paste0("Total differentially expressed genes: ",deg_table %>% nrow()))
+wrap_elements(plt) + plot_annotation(caption = paste0("pCRY"))
 
 ```
+
+## Fig 4b: groups
+```{r groups}
+
+res_
+DEGs_comb <- bind_cols(DEGs[[1]],DEGs[[2]],DEGs[[3]])
+
+```
+
+# Fig. 3: Grouping
+```{r 3_groups, eval=FALSE, fig.show='hold', fig.width=6, fig.height=4, dpi=300, out.width="50%"}
+
+p_new2 <- ggplot(deg_table, aes(x=XX.log2FC, y=XY.log2FC,color=group, fill=group)) + 
+  geom_polygon(data = xyup, aes(x=x,y=y),color=cols2l[2],fill=cols2l[2], alpha=0.2) +
+  geom_polygon(data = xydo, aes(x=x,y=y),color=cols2l[2],fill=cols2l[2], alpha=0.8) +
+  geom_polygon(data = xxdo, aes(x=x,y=y),color=cols2l[1],fill=cols2l[1], alpha=0.8) +
+  geom_polygon(data = xxup, aes(x=x,y=y),color=cols2l[1],fill=cols2l[1], alpha=0.2) +
+  annotate("text", x= text$x,y= text$y,label = paste(text$label,text$size,sep="\n"), color=text$color,size=3 , fontface="bold") +
+  geom_hline(yintercept = 0, linewidth = 0.1) + 
+  geom_vline(xintercept = 0, linewidth = 0.1) +
+  geom_point(shape=21) + 
+  scale_color_manual(values=alpha(c(cols2d[1],cols2d[1], cols2d[2],cols2d[2],rep(colsp,2),rep(colsap,2)),0.8)) +
+  scale_fill_manual(values=alpha(c(cols2d[1],cols2d[1], cols2d[2],cols2d[2],rep(colsp,2),rep(colsap,2)),0.3)) +
+#  scale_color_manual(values = text$color) +
+  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
+  theme_bw() +
+  removeGrid(x=T, y=T)
+
+p_new2 + scale_fill_manual(values=c(cols2d[1],cols2d[1], cols2d[2],cols2d[2],alpha(c(rep(colsp,2),rep(colsap,2)),0.3)))
+
+```

### 2024-10-17 pub_figures/manuscript-figures.Rmd
fix volcano ratio

--- 2024-10-17 pub_figures/manuscript figures.Rmd
@@
 # Fig 3: Volcanos
-```{r volcanos, fig.height=8 }
+```{r volcanos, fig.height=6, fig.width=10}
 
@@
 ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
-width = 8,
+width = 10,
 height = 6)
@@
 
-## Fig 4b: groups
-```{r groups}
+## Fig 4b: Groups
+```{r groups, eval=FALSE}
 
-res_
-DEGs_comb <- bind_cols(DEGs[[1]],DEGs[[2]],DEGs[[3]])
+res_core <- list(dark = res_list$pcry$pcry_D.vs.WT_D,
+             blue = res_list$pcry$pcry_BL.vs.WT_BL,
+             red =res_list$pcry$pcry_R.vs.WT_R) %>% lapply(data.frame)
+
+DEGs_comb <- bind_cols(res_core[[1]][],res_core[[2]],res_core[[3]])
 

### 2024-10-19 pub_figures/manuscript-figures.Rmd
Venn + Grouping

@@
 ## Fig 4a: Venns
-```{r overlap, eval=FALSE}
+```{r overlap, eval=TRUE}
 fig <- "Fig4"
@@
 
-wrap_elements(plt) + plot_annotation(caption = paste0("pCRY"))
+wrap_elements(plt) # + plot_annotation(caption = paste0("pCRY"))
 
@@
 
-DEGs_comb <- bind_cols(res_core[[1]][],res_core[[2]],res_core[[3]])
+DEGs_comb <- bind_cols(res_core[[1]],res_core[[2]][,c(2,6)],res_core[[3]][,c(2,6)])
+colnames(DEGs_comb)[c(2,6,8,9,10,11)] <- c("l2FC.dark","padj.dark","l2FC.blue","padj.blue","l2FC.red","padj.red")
+
+DEGs_genes["dark"]
+blue_red_genes <- unlist(DEGs_genes[c("blue","red")]) %>% unique()
+
+non-sig <- data.frame(x = c(-9, 0, 9), y = c(20, 0.5, 20))
+
+DEGs_comb$sig
+
+gg_all <- ggplot(DEGs_comb, aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
+  geom_point(shape=21) +
+  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
+  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
+  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
+  geom_text_repel() +
+  theme_bw() +
+  removeGrid(x=T, y=T)
+gg_all
+
+gg_br <- ggplot(DEGs_comb[blue_red_genes,], aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
+  geom_point(shape=21) +
+  geom_text_repel() +
+  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
+  theme_bw() +
+  removeGrid(x=T, y=T)
+gg_br
+
+red_specific <- DEGs_comb %>% subset(l2FC.blue < 1 & l2FC.blue > -1 & (l2FC.red > 0.5 | l2FC.red < -0.5) & padj.red < 0.05) %>% arrange(l2FC.red)
+
+goi <- "Cre03.g192050"
+
+# single gene
+d <- plotCounts(dds, gene=goi, intgroup=c("condition","experiment","genotype","treatment"), main=s,returnData=TRUE)
+colnames(d)[1] <- "counts"
+
+gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
+    geom_boxplot(color="black") +
+    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
+  # geom_text_repel(aes(label=clientName)) +
+    scale_fill_manual(values=group.colors) +
+    scale_color_manual(values=group.colors) +
+    scale_y_continuous(trans = "log2") +
+    theme_bw() +
+    removeGrid(x=T, y=T) +
+    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +
+    labs(title = paste(mcols(dds)[goi,]$symbol," in ",colData(dds)$experiment[1],sep = "")) +
+    theme(axis.text.x = element_text(size = 10, face = "bold"),
+    axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
+    axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
+    axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))
+gcounts
 
@@
 
-# Fig. 3: Grouping
+# Fig. X
 ```{r 3_groups, eval=FALSE, fig.show='hold', fig.width=6, fig.height=4, dpi=300, out.width="50%"}

### 2024-10-21 pub_figures/manuscript-figures.Rmd
grouping

@@
 ## Fig 4b: Groups
-```{r groups, eval=FALSE}
+```{r groups, eval=TRUE, out.width="50%", fig.show='hold'}
 
@@
 
-non-sig <- data.frame(x = c(-9, 0, 9), y = c(20, 0.5, 20))
+non_sig <- data.frame(x = c(-9, 0, 9), y = c(20, 0.5, 20))
 
@@
   coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
-  geom_text_repel() +
+  geom_text_repel(size=2, max.overlaps = 20) +
   theme_bw() +
@@
   geom_point(shape=21) +
-  geom_text_repel() +
+  geom_text_repel(size=3) +
   coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +

### 2025-01-02 2_A_dive_aCRY_pCRY (separately)/2_A_dive_aCRY_pCRY_sep.Rmd
results fix

@@
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 knit: (function(input_file, encoding) {
@@
 
+dds <- dds_list$pcry
+
 design(dds)
@@
 
-# Belichtungseffect für jeden Genotyp
+# Belichtungseffekt für jeden Genotyp
 nres <- "WT_BL.vs.D"
@@
 nres <- paste(nlist,"D.vs.WT_D",sep="_")
-contrast <- list(c[2])                # [2,] "genotype_pcry_vs_WT"
+contrast <- list(c(c[2]))                # [2,] "genotype_pcry_vs_WT"
 res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
 
+# !! Fehler?!
+# 31.1.24 fixed: wrong terminus for list!
 nres <- paste(nlist,"BL.vs.WT_BL",sep="_")
-contrast <- list(c[2],c[5])           # [5,] "genotypepcry.treatmentblue"
+contrast <- list(c(c[2],c[5]))           # [5,] "genotypepcry.treatmentblue"
 res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
@@
 nres <- paste(nlist,"R.vs.WT_R",sep="_")
-contrast <- list(c[2],c[6])           # [6,] "genotypepcry.treatmentred"  
+contrast <- list(c(c[2],c[6]))           # [6,] "genotypepcry.treatmentred"  
 res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+
+# # New -> this is correct
+# nres <- paste(nlist,"BL.vs.WT_BL_2",sep="_")
+# contrast <- c(0,1,0,0,1,0)         # [5,] "genotypepcry.treatmentblue"
+# res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
+# 
+# nres <- paste(nlist,"R.vs.WT_R_2",sep="_")
+# contrast <- c(0,1,0,0,0,1)         # [6,] "genotypepcry.treatmentred"  
+# res_l[nres] <- getresults_SK(contrast = contrast, nres = nres)
 
@@
 }
+
+```
+
+## Control results
+```{r ctrl_results, fig.show='hold', out.width='25%'}
+
+colData(gse_pcry)
+design <- ~condition
+dds_ctrl <- DESeqDataSet(gse_pcry, design=design)
+keep.sn <- rowSums(counts(dds_ctrl) >= 5) >= 3
+keep.sn %>% summary()
+dds_ctrl <- dds_ctrl[keep.sn,]
+
+dds_ctrl <- DESeq(dds_ctrl)
+resultsNames(dds_ctrl)
+colData(dds_ctrl)
+res_ctrl_D_pvw <- results(dds_ctrl,contrast=c("condition","pcry_dark","WT_dark"))
+res_ctrl_R_pvw <- results(dds_ctrl,contrast=c("condition","pcry_red","WT_red"))
+res_ctrl_BL_pvw <- results(dds_ctrl,contrast=c("condition","pcry_blue","WT_blue"))
+res_ctrl_WT_rvd <- results(dds_ctrl,contrast=c("condition","WT_red","WT_dark"))
+res_ctrl_WT_blvd <- results(dds_ctrl,contrast=c("condition","WT_blue","WT_dark"))
+res_ctrl_pcry_rvd <- results(dds_ctrl,contrast=c("condition","pcry_red","pcry_dark"))
+res_ctrl_pcry_blvd <- results(dds_ctrl,contrast=c("condition","pcry_blue","pcry_dark"))
+
+((res_ctrl_D_pvw %>% rownames()) == (res_list$pcry$pcry_R.vs.WT_R %>% rownames())) %>% summary()
+
+res_ctrl_D_pvw %>% nrow()
+res_list$pcry$pcry_R.vs.WT_R %>% nrow()
+res_list$pcry %>% names()
+plot(res_ctrl_D_pvw$log2FoldChange ~ res_list$pcry$pcry_D.vs.WT_D$log2FoldChange)
+plot(res_ctrl_R_pvw$log2FoldChange ~ res_list$pcry$pcry_R.vs.WT_R$log2FoldChange)
+# plot(res_ctrl_R_pvw$log2FoldChange ~ res_list$pcry$pcry_R.vs.WT_R_2$log2FoldChange)
+plot(res_ctrl_BL_pvw$log2FoldChange ~ res_list$pcry$pcry_BL.vs.WT_BL$log2FoldChange)
+# plot(res_ctrl_BL_pvw$log2FoldChange ~ res_list$pcry$pcry_BL.vs.WT_BL_2$log2FoldChange)
+plot(res_ctrl_WT_rvd$log2FoldChange ~ res_list$pcry$WT_R.vs.D$log2FoldChange)
+plot(res_ctrl_WT_blvd$log2FoldChange ~ res_list$pcry$WT_BL.vs.D$log2FoldChange)
+plot(res_ctrl_pcry_rvd$log2FoldChange ~ res_list$pcry$pcry_R.vs.D$log2FoldChange)
+plot(res_ctrl_pcry_blvd$log2FoldChange ~ res_list$pcry$pcry_BL.vs.D$log2FoldChange)
+
+res_ctrl_WT_rvd["Cre06.g275350",]
+
+plotCounts(dds, gene="Cre06.g275350")
+plotCounts(dds_ctrl, gene="Cre06.g275350")
+
+# New results
+dds_pcry <- dds_list$pcry
+
+resultsNames(dds_pcry)
+res_pcry_R_WT_pcry <- results(dds_pcry,contrast=c(0,1,0,0,0,1))
+res_pcry_BL_WT_pcry <- results(dds_pcry,contrast=c(0,1,0,0,1,0))
+
+plot(res_ctrl_R_pvw$log2FoldChange ~ res_pcry_R_WT_pcry$log2FoldChange)
+plot(res_ctrl_BL_pvw$log2FoldChange ~ res_pcry_BL_WT_pcry$log2FoldChange)
 

### 2025-01-02 pub_figures/manuscript-figures.Rmd
NMR plot restart
results fix

@@
 editor_options: 
-  chunk_output_type: inline
+  chunk_output_type: console
 ---
@@
 
-goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC114", "ROC55", "CON1", "CRB1")
+goi_tf <- c("CCM1","LCR1", "HY5", "QER7", "QER4", "QER6", "ROC15", "ROC40", "ROC66", "ROC75", "ROC59", "ROC114", "ROC55", "CON1", "CRB1")
 anno[c("Cre17.g745697","Cre13.g567250","Cre01.g043550","Cre02.g079550"),"geneSymbol"] <- c("QER4","QER6","QER7","ROC110")
@@
 
-all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
+all_counts$Gene <- factor(all_counts$Gene, levels = c("CON1", "ROC40", "CRB1" ,  "QER4" , "ROC110",  "QER7" ,"ROC114" , "ROC75", "ROC59","CCM1" , "ROC55" , "LCR1",  "QER6" , "ROC66",    "ROC15")  )
 levels(all_counts$Gene)
@@
 # Plot
+all_counts$Gene
 gcounts_tf1 <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
@@
 # single gene
+
     d <- plotCounts(dds, gene=ROC40, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
@@
 
-gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
+gcounts_roc40 <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
     geom_boxplot(color="black") +
@@
 
-gcounts %>% print()
+gcounts_roc40 %>% print()
 
-ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_ROC40.pdf",sep=""), plot = gcounts,
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_ROC40.pdf",sep=""), plot = gcounts_roc40,
 width = 6,
@@
 }
+
+gcounts_roc40
 
@@
 
+## individual counts
+```{r plot_counts2}
+ROC59 <- "Cre10.g425050"
+goi <- ROC59
+
+    d <- plotCounts(dds, gene=goi, intgroup=c("condition","experiment","genotype","treatment","clientName"), main=s,returnData=TRUE)
+    colnames(d)[1] <- "counts"
+
+gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
+    geom_boxplot(color="black") +
+    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
+  # geom_text_repel(aes(label=clientName)) +
+    scale_fill_manual(values=group.colors) +
+    scale_color_manual(values=group.colors) +
+    scale_y_continuous(trans = "log2") +
+    theme_bw() +
+  removeGrid(x=T, y=T) +
+    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +
+
+    labs(title = paste(goi," (",mcols(dds)[goi,"id.symbol"], ")", " in ",colData(dds)$experiment[1],sep = "")) +
+ theme(axis.text.x = element_text(size = 10, face = "bold"),
+        axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
+        axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
+        axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))
+
+gcounts
+
+gcounts_roc59 <- gcounts
+
+```
+
+# Fig 2x: Explain data
+```{r explain_data, eval=TRUE}
+
+c_graphic <- png::readPNG("Contrasts.png", native = TRUE)
+patchwork::wrap_elements((c_graphic))
+
+res_core <- list(dark = res_list$pcry$pcry_D.vs.WT_D,
+             blue = res_list$pcry$pcry_BL.vs.WT_BL,
+             red =res_list$pcry$pcry_R.vs.WT_R) %>% lapply(data.frame)
+
+DEGs_comb <- bind_cols(res_core[[1]],res_core[[2]][,c(2,6)],res_core[[3]][,c(2,6)])
+colnames(DEGs_comb)[c(2,6,8,9,10,11)] <- c("l2FC.dark","padj.dark","l2FC.blue","padj.blue","l2FC.red","padj.red")
+
+res_list[["pcry"]] %>% names()
+
+res_list[["pcry"]][["WT_R.vs.D"]][ROC40,c(2,6)]
+res_list[["pcry"]][["pCRY_R.vs.D"]][ROC40,c(2,6)]
+
+# anno_tf
+ROC59 <- "Cre10.g425050"
+goi  <- ROC40  # ROC40 ROC59
+
+results <- lapply(res_list[["pcry"]],data.frame)
+l2fc <- sapply(results,"[[",goi,c(2)) %>% round(digits=2)
+names(l2fc) <- paste0("l2fc.",names(l2fc))
+p <- sapply(results,"[[",goi,c(6)) %>% round(digits=3)
+
+goi_lp <- list("l2fc"=l2fc,
+               "p"=p)
+
+l2fc
+# gcounts_roc40
+gcounts_roc40 + coord_cartesian(ylim = c(1,10000)) + 
+  # (1.)
+  geom_text(aes(x = 0.5,y = 120),
+            label="1.", color="red2") +
+  # "pcry_D.vs.WT_D" 
+  geom_segment(
+        aes(x = 0.7,y = 90, xend = 1.3,yend = 90),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="red2") +
+  geom_text(aes(x = 1,y = 120),
+            label=goi_lp$l2fc["l2fc.pcry_D.vs.WT_D"], color="red2") +
+
+    # "pcry_BL.vs.WT_BL" 
+  geom_segment(
+        aes(x = 1.7,y = 90, xend = 2.3,yend = 90),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="red2") +
+  geom_text(aes(x = 2,y = 120),
+            label=goi_lp$l2fc["l2fc.pcry_BL.vs.WT_BL"], color="red2") +
+
+    # "pcry_R.vs.WT_R" 
+  geom_segment(
+        aes(x = 2.7,y = 75, xend = 3.3,yend = 75),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="red2") +
+  geom_text(aes(x = 3,y = 100),
+            label=goi_lp$l2fc["l2fc.pcry_R.vs.WT_R"], color="red2") +
+
+  # (2.)  
+  geom_text(aes(x = 0.5,y = 50),
+            label="2.", color="green4") +
+  # ""WT_BL.vs.D"" 
+    geom_segment(
+        aes(x = 0.7,y = 40, xend = 1.7,yend = 40),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="green4") +
+  geom_text(aes(x = 1,y = 55),
+            label=goi_lp$l2fc["l2fc.WT_BL.vs.D"], color="green4") +
+  # 3.
+  geom_text(aes(x = 0.5,y = 20),
+            label="3.", color="purple3") +
+    # "WT_R.vs.D" 
+    geom_segment(
+        aes(x = 1.3,y = 35, xend = 1.7,yend = 12),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="purple3") +
+  geom_text(aes(x = 1.7,y = 25),
+            label=goi_lp$l2fc["l2fc.pcry_BLvD.vs.WT_BLvD"], color="purple3") +
+  # "pcry_BL.vs.D"
+    geom_segment(
+        aes(x = 1.2,y = 9, xend = 2.3,yend = 9),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="green3") +
+  geom_text(aes(x = 2,y = 12),
+            label=goi_lp$l2fc["l2fc.pcry_BL.vs.D"], color="green3") +
+
+    # "WT_R.vs.D" 
+    geom_segment(
+        aes(x = 0.7,y = 6, xend = 2.7,yend = 6),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="green4") +
+  geom_text(aes(x = 0.9,y = 8),
+            label=goi_lp$l2fc["l2fc.WT_R.vs.D"], color="green4") +
+  # "pcry_BL.vs.D" 
+    geom_segment(
+        aes(x = 1.2,y = 1, xend = 3.3,yend = 1),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="green3") +
+  geom_text(aes(x = 3,y = 1.3),
+            label=goi_lp$l2fc["l2fc.pcry_R.vs.D"], color="green3") +
+
+  # "pcry_BL.vs.D" 
+    geom_segment(
+        aes(x = 1.7,y = 5, xend = 2.3,yend = 1.3),
+        arrow = arrow(length = unit(0.03,units = "npc")),color ="purple1") +
+  geom_text(aes(x = 2.3,y = 2.7),
+            label=goi_lp$l2fc["l2fc.pcry_RvD.vs.WT_RvD"], color="purple1") #+
+
+DEGs_comb[ROC40,]
+
+```
+
 # Fig 3: Volcanos
-```{r volcanos, fig.height=6, fig.width=10}
+## a) Genotype (1)
+```{r volcanos1, fig.height=6, fig.width=10}
 
@@
 
-res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
+# res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
 res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
-# res[res$padj < pmax,]$padj <- pmax
 
+x <- res$padj
+res[(x < pmax) & (!is.na(x)),]$padj <- pmax
 subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
@@
 height = 6)
+
+```
+
+## b): light effect (2)
+```{r volcanos2, fig.height=6, fig.width=10}
+
+fig <- "Fig3"
+
+dds <- dds_list[["pcry"]]
+
+res_ashr_list %>% names()
+res_ashr_list[[2]] %>% names()
+
+res <- res_ashr_list$pcry$WT_R.vs.D
+res_n <- res_list$pcry$WT_R.vs.D
+
+# of shrinked results
+total <- subset(res, padj< 0.05) %>% nrow()
+up <- subset(res, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# of "true" results
+total <- subset(res_n, padj< 0.05) %>% nrow()
+up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# points outside the grid
+subset(res, padj < 10^-50 | log2FoldChange > 6 | log2FoldChange < -6)
+#                 baseMean log2FoldChange     lfcSE      pvalue        padj
+#               <numeric>      <numeric> <numeric>   <numeric>   <numeric>
+# Cre16.g681750 2907.4406       -3.51641   0.22216 2.99151e-57 4.79270e-53
+# Cre17.g802135   46.8587       36.40049   7.35075 1.00156e-08 7.64096e-06
+
+res["Cre16.g681750","padj"] <- 10^-50
+res["Cre17.g802135","log2FoldChange"] <- 6
+
+mcols(dds_list[["pcry"]]) %>% nrow()
+res %>% nrow()
+
+volcano_WT_red <- EnhancedVolcano(res,
+    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
+    x = 'log2FoldChange',
+    y = 'padj',
+    col=c("grey","grey","grey",group.colors[5]),
+    title = "WT_R.vs.D",
+    titleLabSize = 12,
+    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
+#    subtitle = {},
+    subtitleLabSize = 10,
+    caption = NULL,
+    xlim = c(-7,7),
+    ylim = c(0,100),
+    pCutoff = 0.05,
+    FCcutoff = 1,
+    maxoverlapsConnectors = 40,
+    drawConnectors = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = "grey70",
+    legendLabels=c('ns','ns','ns',
+      'padj < 0.05 & Log2FC > 1'),
+    labSize = 4,
+    axisLabSize = 12,
+    legendLabSize = 12,
+    legendIconSize = 4,
+    gridlines.major = FALSE,
+    gridlines.minor = FALSE,
+    pointSize = 3
+)
+# volcano_red
+
+# blue
+
+res_ashr_list %>% names()
+res_ashr_list[[1]] %>% names()
+
+res <- res_ashr_list$pcry$pcry_R.vs.D
+res_n <- res_list$pcry$pcry_R.vs.D
+
+# of "true" results
+total <- subset(res_n, padj< 0.05) %>% nrow()
+up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# points outside the grid
+
+pmax <- 10^-100
+l2FCmax <- 6
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+# res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
+res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
+x <- res$padj
+res[(x < pmax) & (!is.na(x)),]$padj <- pmax
+
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+mcols(dds_list[["pcry"]]) %>% nrow()
+res %>% nrow()
+
+volcano_pcry_R <- EnhancedVolcano(res,
+    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
+    x = 'log2FoldChange',
+    y = 'padj',
+    col=c("grey","grey","grey",group.colors[6]),
+    title = "pcry_R.vs.D",
+    titleLabSize = 12,
+    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
+#    subtitle = {},
+    subtitleLabSize = 10,
+    caption = NULL,
+    xlim = c(-7,7),
+    ylim = c(0,100),
+    pCutoff = 0.05,
+    FCcutoff = 1,
+    maxoverlapsConnectors = 40,
+    drawConnectors = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = "grey70",
+    legendLabels=c('ns','ns','ns',
+      'padj < 0.05 & Log2FC > 1'),
+    labSize = 4,
+    axisLabSize = 12,
+    legendLabSize = 12,
+    legendIconSize = 4,
+    gridlines.major = FALSE,
+    gridlines.minor = FALSE,
+    pointSize = 3
+)
+# volcano_blue
+
+# WT BLUE
+
+res_ashr_list %>% names()
+res_ashr_list[[2]] %>% names()
+
+res <- res_ashr_list$pcry$WT_BL.vs.D
+res_n <- res_list$pcry$WT_BL.vs.D
+
+# of "true" results
+total <- subset(res_n, padj< 0.05) %>% nrow()
+up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# points outside the grid
+
+pmax <- 10^-100
+l2FCmax <- 6
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
+# res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
+res[subset(res, padj < pmax) %>% rownames(),]$padj <- pmax
+
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+mcols(dds_list[["pcry"]]) %>% nrow()
+res %>% nrow()
+
+volcano_WT_BL <- EnhancedVolcano(res,
+    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
+    x = 'log2FoldChange',
+    y = 'padj',
+    col=c("grey","grey","grey",group.colors[3]),
+    title = "WT_BL.vs.D",
+    titleLabSize = 12,
+    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
+#    subtitle = {},
+    subtitleLabSize = 10,
+    caption = NULL,
+    xlim = c(-7,7),
+    ylim = c(0,100),
+    pCutoff = 0.05,
+    FCcutoff = 1,
+    maxoverlapsConnectors = 40,
+    drawConnectors = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = "grey70",
+    legendLabels=c('ns','ns','ns',
+      'padj < 0.05 & Log2FC > 1'),
+    labSize = 4,
+    axisLabSize = 12,
+    legendLabSize = 12,
+    legendIconSize = 4,
+    gridlines.major = FALSE,
+    gridlines.minor = FALSE,
+    pointSize = 3
+)
+
+# pcry BLUE
+
+res_ashr_list %>% names()
+res_ashr_list[[2]] %>% names()
+
+res <- res_ashr_list$pcry$pcry_BL.vs.D
+res_n <- res_list$pcry$pcry_BL.vs.D
+
+# of "true" results
+total <- subset(res_n, padj< 0.05) %>% nrow()
+up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# points outside the grid
+
+pmax <- 10^-100
+l2FCmax <- 6
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
+res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
+res[subset(res, padj < pmax) %>% rownames(),]$padj <- pmax
+
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+mcols(dds_list[["pcry"]]) %>% nrow()
+res %>% nrow()
+
+volcano_pcry_BL <- EnhancedVolcano(res,
+    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
+    x = 'log2FoldChange',
+    y = 'padj',
+    col=c("grey","grey","grey",group.colors[4]),
+    title = "pcry_BL.vs.D",
+    titleLabSize = 12,
+    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
+#    subtitle = {},
+    subtitleLabSize = 10,
+    caption = NULL,
+    xlim = c(-7,7),
+    ylim = c(0,100),
+    pCutoff = 0.05,
+    FCcutoff = 1,
+    maxoverlapsConnectors = 40,
+    drawConnectors = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = "grey70",
+    legendLabels=c('ns','ns','ns',
+      'padj < 0.05 & Log2FC > 1'),
+    labSize = 4,
+    axisLabSize = 12,
+    legendLabSize = 12,
+    legendIconSize = 4,
+    gridlines.major = FALSE,
+    gridlines.minor = FALSE,
+    pointSize = 3
+)
+
+volcanos_2 <- (volcano_WT_BL + volcano_WT_red) /  (volcano_pcry_BL + volcano_pcry_R) +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'none', axis.title=element_text(size=12))
+volcanos_2
+
+# ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
+# width = 10,
+# height = 6)
 

### 2025-01-04 pub_figures/manuscript-figures.Rmd
NMR plot v1

@@
 editor_options: 
-  chunk_output_type: console
+  chunk_output_type: inline
 ---
@@
 # Fig 3: Volcanos
-## a) Genotype (1)
+## a) Δgenotype per condition (1)
 ```{r volcanos1, fig.height=6, fig.width=10}
@@
 
-ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos1.pdf",sep=""), plot = volcanos_all,
 width = 10,
@@
 
-## b): light effect (2)
-```{r volcanos2, fig.height=6, fig.width=10}
+## b): Δlight = light effect (2)
+```{r volcanos2, fig.height=10, fig.width=10}
 
@@
 
-# ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos.pdf",sep=""), plot = volcanos_all,
-# width = 10,
-# height = 6)
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos2.pdf",sep=""), plot = volcanos_2,
+width = 10,
+height = 12)
+
+```
+
+## c): ΔgenotypeΔlight = difference in light effect (interaction term) (3)
+```{r volcanos3, fig.height=6, fig.width=10}
+
+fig <- "Fig3"
+
+dds <- dds_list[["pcry"]]
+
+res_ashr_list %>% names()
+res_ashr_list[[2]] %>% names()
+
+res <- res_ashr_list$pcry$pcry_BLvD.vs.WT_BLvD
+res_n <- res_list$pcry$pcry_BLvD.vs.WT_BLvD
+
+# of shrinked results
+total <- subset(res, padj< 0.05) %>% nrow()
+up <- subset(res, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# of "true" results
+total <- subset(res_n, padj< 0.05) %>% nrow()
+up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# points outside the grid
+
+pmax <- 10^-50
+l2FCmax <- 6
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+# res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
+res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
+
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+mcols(dds_list[["pcry"]]) %>% nrow()
+res %>% nrow()
+
+volcano3_blue <- EnhancedVolcano(res,
+    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
+    x = 'log2FoldChange',
+    y = 'padj',
+    col=c("grey","grey","grey",group.colors[3]),
+    title = "pcry_BLvD.vs.WT_BLvD",
+    titleLabSize = 12,
+    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
+#    subtitle = {},
+    subtitleLabSize = 10,
+    caption = NULL,
+    xlim = c(-7,7),
+    ylim = c(0,50),
+    pCutoff = 0.05,
+    FCcutoff = 1,
+    maxoverlapsConnectors = 50,
+    drawConnectors = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = "grey70",
+    legendLabels=c('ns','ns','ns',
+      'padj < 0.05 & Log2FC > 1'),
+    labSize = 4,
+    axisLabSize = 12,
+    legendLabSize = 12,
+    legendIconSize = 3,
+    gridlines.major = FALSE,
+    gridlines.minor = FALSE,
+    pointSize = 3
+)
+
+# volcano_red
+res_ashr_list %>% names()
+res_ashr_list[[2]] %>% names()
+
+res <- res_ashr_list$pcry$pcry_RvD.vs.WT_RvD
+res_n <- res_list$pcry$pcry_RvD.vs.WT_RvD
+
+# of "true" results
+total <- subset(res_n, padj< 0.05) %>% nrow()
+up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
+down <- subset(res_n, padj< 0.05 & log2FoldChange < -1) %>% nrow()
+
+# points outside the grid
+
+pmax <- 10^-50
+l2FCmax <- 6
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+res[res$log2FoldChange > l2FCmax,]$log2FoldChange <- l2FCmax
+res[res$log2FoldChange < -l2FCmax,]$log2FoldChange <- -l2FCmax
+x <- res$padj
+res[(x < pmax) & (!is.na(x)),]$padj <- pmax
+
+subset(res, padj < pmax | log2FoldChange > l2FCmax | log2FoldChange < -l2FCmax)
+
+mcols(dds_list[["pcry"]]) %>% nrow()
+res %>% nrow()
+
+volcano3_red <- EnhancedVolcano(res,
+    lab = mcols(dds_list[["pcry"]])[["geneSymbol"]],
+    x = 'log2FoldChange',
+    y = 'padj',
+    col=c("grey","grey","grey",group.colors[5]),
+    title = "pcry_RvD.vs.WT_RvD",
+    titleLabSize = 12,
+    subtitle = paste0("upregulated: ",up,", downregulated: ",down,"\n(total: ",total,")"),
+#    subtitle = {},
+    subtitleLabSize = 10,
+    caption = NULL,
+    xlim = c(-7,7),
+    ylim = c(0,50),
+    pCutoff = 0.05,
+    FCcutoff = 1,
+    maxoverlapsConnectors = 50,
+    drawConnectors = TRUE,
+    widthConnectors = 0.5,
+    colConnectors = "grey70",
+    legendLabels=c('ns','ns','ns',
+      'padj < 0.05 & Log2FC > 1'),
+    labSize = 4,
+    axisLabSize = 12,
+    legendLabSize = 12,
+    legendIconSize = 4,
+    gridlines.major = FALSE,
+    gridlines.minor = FALSE,
+    pointSize = 3
+)
+
+volcanos_3 <- (volcano3_blue + volcano3_red) +
+  plot_layout(guides = "collect", axes="collect", axis_titles="collect") & 
+  theme(legend.position = 'none', axis.title=element_text(size=12))
+volcanos_3
+
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Volcanos3.pdf",sep=""), plot = volcanos_3,
+width = 10,
+height = 6)
+
+colnames(anno)
+anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,]
+anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,c("geneSymbol","previousIdentifiers","Description","previousIdentifiers","Description","Comments","TMHMM_transmembrane","TargetP","Predalgo","Flagellar_Proteome")] %>% kable()
 
@@
 
-res_core <- list(dark = res_list$pcry$pcry_D.vs.WT_D,
-             blue = res_list$pcry$pcry_BL.vs.WT_BL,
-             red =res_list$pcry$pcry_R.vs.WT_R) %>% lapply(data.frame)
-
-DEGs_comb <- bind_cols(res_core[[1]],res_core[[2]][,c(2,6)],res_core[[3]][,c(2,6)])
-colnames(DEGs_comb)[c(2,6,8,9,10,11)] <- c("l2FC.dark","padj.dark","l2FC.blue","padj.blue","l2FC.red","padj.red")
-
-DEGs_genes["dark"]
-blue_red_genes <- unlist(DEGs_genes[c("blue","red")]) %>% unique()
-
-non_sig <- data.frame(x = c(-9, 0, 9), y = c(20, 0.5, 20))
-
-DEGs_comb$sig
-
-gg_all <- ggplot(DEGs_comb, aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
-  geom_point(shape=21) +
+goi_tf
+
+res_core <- list(WT_blue = res_list$pcry$WT_BL.vs.D,
+             WT_red = res_list$pcry$WT_R.vs.D,
+             pcry_blue = res_list$pcry$pcry_BL.vs.D,
+             pcry_red = res_list$pcry$pcry_R.vs.D,
+             dd_blue = res_list$pcry$pcry_BLvD.vs.WT_BLvD,
+             dd_red = res_list$pcry$pcry_RvD.vs.WT_RvD) %>% lapply(data.frame)
+
+# plot(res_list$pcry$pcry_BL.vs.D$log2FoldChange~res_list$pcry$pcry_R.vs.D$log2FoldChange)
+#res_list$pcry$WT_BL.vs.D$symbol == anno[res_list$pcry$WT_BL.vs.D %>% rownames(),"geneSymbol"]
+
+res_comb <- bind_cols(symbol = anno[res_list$pcry$WT_BL.vs.D %>% rownames(),"geneSymbol"],
+                       res_core[[1]][,c(1,2,6)],
+                       res_core[[2]][,c(2,6)],
+                       res_core[[3]][,c(2,6)],
+                       res_core[[4]][,c(2,6)],
+                       res_core[[5]][,c(2,6)],
+                       res_core[[6]][,c(2,6)])
+
+colnames(res_comb)[c(3,4)] <- c("l2FC.WT_blue","padj.WT_blue")
+colnames(res_comb)[c(5,6)] <- c("l2FC.WT_red","padj.WT_red")
+colnames(res_comb)[c(7,8)] <- c("l2FC.pcry_blue","padj.pcry_blue")
+colnames(res_comb)[c(9,10)] <- c("l2FC.pcry_red","padj.pcry_red")
+colnames(res_comb)[c(11,12)] <- c("l2FC.dd_blue","padj.dd_blue")
+colnames(res_comb)[c(13,14)] <- c("l2FC.dd_red","padj.dd_red")
+
+res_comb$symbol2 <- anno[res_comb %>% rownames(),"geneSymbol"]
+  blue_red_genes <- c(deg_list$pcry$deg_pcry_BLvD.vs.WT_BLvD,
+                    deg_list$pcry$deg_pcry_RvD.vs.WT_RvD) %>%
+  unique()
+
+non_sig <- data.frame(x = c(-20, 0, 20), y = c(20, 0.5, 20))
+
+df <- res_comb[anno_tf$gene_id,]
+df <- res_comb[anno_phot$gene_id,]
+
+gg_phots <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=symbol)) + # color=group, fill=group 
+    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
+        arrow = arrow(length = unit(0.01,units = "npc"))) +
+  scale_color_gradient2(mid="grey") +
+  geom_point(shape=21, fill="grey", col="grey40") +
+  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
   geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
   geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
-  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
-  geom_text_repel(size=2, max.overlaps = 20) +
+  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
+  # coord_cartesian(xlim=c(-5,5),ylim = c(-5,5)) +
+  coord_fixed() +
+  geom_text_repel(size=4, max.overlaps = 20) +
   theme_bw() +
   removeGrid(x=T, y=T)
-gg_all
-
-gg_br <- ggplot(DEGs_comb[blue_red_genes,], aes(x=l2FC.blue, y=l2FC.red, label=symbol)) + # color=group, fill=group 
-  geom_point(shape=21) +
-  geom_text_repel(size=3) +
-  coord_cartesian(xlim=c(-10,10),ylim = c(-10,10)) +
+gg_phots
+
+# ROCs
+df <- res_comb[anno_tf$gene_id,]
+gg_rocs <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=symbol)) + # color=group, fill=group 
+    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
+        arrow = arrow(length = unit(0.01,units = "npc"))) +
+  scale_color_gradient2(mid="grey") +
+  geom_point(shape=21, fill="grey", col="grey40") +
+  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
+  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
+  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
+  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
+  # coord_cartesian(xlim=c(-5,5),ylim = c(-5,5)) +
+  coord_fixed() +
+  geom_text_repel(size=4, max.overlaps = 20) +
   theme_bw() +
   removeGrid(x=T, y=T)
-gg_br
-
-red_specific <- DEGs_comb %>% subset(l2FC.blue < 1 & l2FC.blue > -1 & (l2FC.red > 0.5 | l2FC.red < -0.5) & padj.red < 0.05) %>% arrange(l2FC.red)
-
-goi <- "Cre03.g192050"
-
-# single gene
-d <- plotCounts(dds, gene=goi, intgroup=c("condition","experiment","genotype","treatment"), main=s,returnData=TRUE)
-colnames(d)[1] <- "counts"
-
-gcounts <- ggplot(d, aes(x = treatment, y = counts, fill=condition, color=condition)) +
-    geom_boxplot(color="black") +
-    geom_point(shape=21,color="black",aes(fill=condition),position=position_dodge(width=0.75), alpha=1) +
-  # geom_text_repel(aes(label=clientName)) +
-    scale_fill_manual(values=group.colors) +
-    scale_color_manual(values=group.colors) +
-    scale_y_continuous(trans = "log2") +
+gg_rocs
+
+# TOP
+df <- res_comb[blue_red_genes,]
+
+gg_top <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=symbol)) + # color=group, fill=group 
+    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
+        arrow = arrow(length = unit(0.01,units = "npc"))) +
+  scale_color_gradient2(mid="grey") +
+  geom_point(shape=21, fill="grey", col="grey40") +
+  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
+  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
+  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
+  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
+  coord_cartesian(xlim=c(-5,5),ylim = c(-5,5)) +
+  # coord_fixed() +
+  geom_text_repel(size=4, max.overlaps = 20) +
   theme_bw() +
-    removeGrid(x=T, y=T) +
-    geom_vline(xintercept=seq(1,length(levels(all_counts$treatment))-1,1)+.5,color="grey") +
-    labs(title = paste(mcols(dds)[goi,]$symbol," in ",colData(dds)$experiment[1],sep = "")) +
-    theme(axis.text.x = element_text(size = 10, face = "bold"),
-    axis.text.y = element_text(color = "grey20", size = 8, angle = 0, hjust = 1, vjust = 0, face = "plain"),
-    axis.title.x = element_text(color = "white", size = 12, angle = 0, hjust = .5, vjust = 0, face = "plain"),
-    axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "bold"))
-gcounts
+  removeGrid(x=T, y=T)
+gg_top
 

### 2025-07-17 pub_figures/manuscript-figures.Rmd
annotation xls

@@
 anno_phot <- anno_phot[goi_phot,]
-
-# Combined counts table TFs
 
@@
 # of shrinked results
-total <- subset(res, padj< 0.05) %>% nrow()
+total <- subset(res, padj< 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )) %>% nrow()
 up <- subset(res, padj< 0.05 & log2FoldChange > 1) %>% nrow()
@@
 # of "true" results
-total <- subset(res_n, padj< 0.05) %>% nrow()
+total <- subset(res_n, padj< 0.05 & (log2FoldChange > 1 | log2FoldChange < -1 )) %>% nrow()
 up <- subset(res_n, padj< 0.05 & log2FoldChange > 1) %>% nrow()
@@
 colnames(anno)
-anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,]
+# anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,]
 anno[deg_list$pcry$deg_pcry_RvD.vs.WT_RvD,c("geneSymbol","previousIdentifiers","Description","previousIdentifiers","Description","Comments","TMHMM_transmembrane","TargetP","Predalgo","Flagellar_Proteome")] %>% kable()
@@
   geom_point(shape=21, fill="grey", col="grey40") +
-  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red),shape=21, fill="green3", col="grey40") +
+  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red, fill = "pcry"),shape=21, col="grey40") +
+  scale_fill_manual(name="Genotype",values="green3") +
   geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
@@
 df <- res_comb[blue_red_genes,]
+df %>% kable()
+
+plot(df$l2FC.dd_red ~ df$l2FC.dd_blue)
 
@@
 
+  scale_colour_manual(name="Error Bars",values=cols)
+
 ```

### 2025-07-17 1_data_processing/1_data_processing.Rmd
Rstudio fix
manuscript figures, new structure, PCA, Counts, Voclanos
interaction counts pdf

@@
 coldata_Cre <- filter(coldata, species == "Cre")
-se <- tximeta(coldata_Cre, useHub=T)
+se <- tximeta(coldata_Cre, useHub=F)
 se
@@
 
+colnames(anno)
+
 # anno[1:10,1:10]
@@
 # load(paste(dir,"anno.RDS",sep="/"))
+
+anno_xls <- subset(anno, select = -previousIdentifiers_list)
+write_xlsx(data.frame(anno_xls),
+           paste(gitdir,"Chlamy_master_annotation.xlsx",sep="/"))
 

### 2025-07-17 pub_figures/manuscript-figures_bu.Rmd
interaction counts pdf

--- 2025-07-17 pub_figures/manuscript-figures.Rmd
@@
 
-# Fig. X
+# Fig. 5: pCRY interaction partner
+## load csv
+```{r load_interaction}
+top20 <- read.csv2("../UniProt-and-Cre-ID's.csv")
+top20_id <- top20$'Cre.ID' %>% str_remove_all(pattern=" ")
+summary(top20_id %in% anno$gene_id)
+
+```
+
+## Plot interaction counts
+```{r plot_interaction}
+## Interaction genes all-in-one
+fig <- "Fig5"
+goi_interaction <- top20_id
+anno_interaction <- anno[top20_id,]
+
+goi <- anno_interaction
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"id.symbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene)
+all_counts$Gene <- fct_reorder(all_counts$Gene, all_counts$count, .fun = median, .desc = TRUE)
+levels(all_counts$Gene)
+
+# all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
+# levels(all_counts$Gene)
+
+# Plot
+max_val <- 1.0*max(all_counts$count)
+gcounts_interaction <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  labs(title = "Interaction partners") + 
+  theme_bw() +
+  removeGrid(x=T, y=T) +
+  geom_vline(xintercept=seq(1,length(levels(all_counts$Gene))-1,1)+.5,color="grey") +
+  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
+gcounts_interaction %>% print()
+
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_Interaction.pdf",sep=""), plot = gcounts_interaction,
+width = 30,
+height = 8)
+
+```
+
+# Fig. Y
 ```{r 3_groups, eval=FALSE, fig.show='hold', fig.width=6, fig.height=4, dpi=300, out.width="50%"}

### 2025-07-17 pub_figures/manuscript-figures.Rmd
scatter pdf

@@
 res_comb$symbol2 <- anno[res_comb %>% rownames(),"geneSymbol"]
+res_comb$id.symbol <- anno[res_comb %>% rownames(),"id.symbol"]
+
   blue_red_genes <- c(deg_list$pcry$deg_pcry_BLvD.vs.WT_BLvD,
@@
 
-# Fig. X
+# Fig. 5: pCRY interaction partner
+## load csv
+```{r load_interaction}
+top20 <- read.csv2("../UniProt-and-Cre-ID's.csv")
+top20_id <- top20$'Cre.ID' %>% str_remove_all(pattern=" ")
+summary(top20_id %in% anno$gene_id)
+
+```
+
+## Plot interaction counts
+```{r plot_interaction}
+## Interaction genes all-in-one
+fig <- "Fig5"
+goi_interaction <- top20_id
+anno_interaction <- anno[top20_id,]
+
+goi <- anno_interaction
+l <- nrow(goi)
+all_counts <- {}
+for (i in 1:l){
+  d <-  plotCounts(dds, gene=goi[i,"gene_id"], intgroup=c("condition","experiment","treatment","genotype"), col=col,main=res$SYMBOL[i],returnData=TRUE)
+  d$Gene <- rep(goi[i,"id.symbol"],length(rownames(d)))
+  d$sample <- rownames(d)
+  rownames(d) <- {}
+  all_counts <- bind_rows(all_counts,d)
+  }
+
+all_counts$Gene
+levels(all_counts$condition)
+levels(all_counts$Gene)
+
+all_counts$Gene <- factor(all_counts$Gene)
+all_counts$Gene <- fct_reorder(all_counts$Gene, all_counts$count, .fun = median, .desc = TRUE)
+levels(all_counts$Gene)
+
+# all_counts$Gene <- factor(all_counts$Gene, levels = c("PHOT1","CHR1","CHR2","HKR1","UVR8","DCRY1","PCRY1","ACRY1")) # "DCRY2", 
+# levels(all_counts$Gene)
+
+# Plot
+max_val <- 1.0*max(all_counts$count)
+gcounts_interaction <- ggplot(all_counts, aes(x = Gene, y = count, fill=condition)) +
+  geom_boxplot(fatten = 1) +
+  scale_fill_manual(values = group.colors) +
+  labs(title = "Interaction partners") + 
+  theme_bw() +
+  removeGrid(x=T, y=T) +
+  geom_vline(xintercept=seq(1,length(levels(all_counts$Gene))-1,1)+.5,color="grey") +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
+  scale_y_continuous(trans = "log2") & plot_annotation(title = colData(dds)$experiment[1])
+gcounts_interaction %>% print()
+
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Counts_Interaction.pdf",sep=""), plot = gcounts_interaction,
+width = 15,
+height = 8)
+
+```
+
+## scatter plot
+```{r interaction_scatter}
+# Interaction
+df <- res_comb[anno_interaction$gene_id,]
+
+gg_interaction <- ggplot(df, aes(x=l2FC.WT_blue, y=l2FC.WT_red, label=id.symbol)) + # color=group, fill=group 
+    geom_segment(aes(x = l2FC.WT_blue, y= l2FC.WT_red, xend = l2FC.pcry_blue,yend = l2FC.pcry_red, color = l2FC.dd_red),
+        arrow = arrow(length = unit(0.01,units = "npc"))) +
+  scale_color_gradient2(mid="grey") +
+  geom_point(aes(x=l2FC.WT_blue, y=l2FC.WT_red, fill="WT"),shape=21, col="grey40") +
+  geom_point(aes(x=l2FC.pcry_blue, y=l2FC.pcry_red, fill = "pcry"),shape=21, col="grey40") +
+  scale_fill_manual(name="Genotype",values=c("green3","grey")) + 
+  geom_hline(yintercept = c(-1,1), linewidth = 0.1) + 
+  geom_vline(xintercept = c(-1,1), linewidth = 0.1) +
+  geom_abline(slope=c(1), intercept = 0, linewidth = 0.1) +
+  # coord_cartesian(xlim=c(-2,2),ylim = c(-2,2)) +
+  coord_fixed(xlim=c(-2,2),ylim = c(-2,2)) +
+  geom_text_repel(size=3, max.overlaps = 20) +
+  theme_bw() +
+  removeGrid(x=T, y=T)
+gg_interaction
+
+ggsave(paste(fig,"_",colData(dds)$experiment[1],"_Scatter_Interaction.pdf",sep=""), plot = gg_interaction,
+width = 10,
+height = 10)
+
+write_xlsx(data.frame(df),"Interaction_results.xlsx")
+
+```
+
+# Fig. Y
 ```{r 3_groups, eval=FALSE, fig.show='hold', fig.width=6, fig.height=4, dpi=300, out.width="50%"}
